---
title: "Hands-on_Ex04_FunnelPlots"
format: html
author: "Persis Evangeline"
date: "May 7, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Getting started

```{r}
pacman::p_load(tidyverse, FunnelPlotR, plotly, knitr)
```

# Importing Excel file

```{r}
covid19 <- read_csv("COVID-19_DKI_Jakarta.csv") %>%
  mutate_if(is.character, as.factor)

```

# FunnelPlotR methods

## Basic Plot

::: panel-tabset
## The plot

```{r}
#| echo: false

funnel_plot(
  .data = covid19,
  numerator = Positive,
  denominator = Death,
  group = `Sub-district`
)
```

## The code

```{r}
#| eval: false

funnel_plot(
  .data = covid19,
  numerator = Positive,
  denominator = Death,
  group = `Sub-district`)

```
:::

## Makeover 1

::: panel-tabset
## The plot

```{r}
#| echo: false

funnel_plot(
  .data = covid19,
  numerator = Death,
  denominator = Positive,
  group = `Sub-district`,
  data_type = "PR",     #<<
  xrange = c(0, 6500),  #<<
  yrange = c(0, 0.05)   #<<
)
```

## The code

```{r}
#| eval: false

funnel_plot(
  .data = covid19,
  numerator = Death,
  denominator = Positive,
  group = `Sub-district`,
  data_type = "PR",     
  xrange = c(0, 6500),  
  yrange = c(0, 0.05)   
)
```
:::

## Makeover 2

::: panel-tabset
## The plot

```{r}
#| echo: false

funnel_plot(
  .data = covid19,
  numerator = Death,
  denominator = Positive,
  group = `Sub-district`,
  data_type = "PR",   
  xrange = c(0, 6500),  
  yrange = c(0, 0.05),
  label = NA,
  title = "Cumulative COVID-19 Fatality Rate by Cumulative Total Number of COVID-19 Positive Cases", #<<           
  x_label = "Cumulative COVID-19 Positive Cases", #<<
  y_label = "Cumulative Fatality Rate"  #<<
)
```

## The code

```{r}
#| eval: false

funnel_plot(
  .data = covid19,
  numerator = Death,
  denominator = Positive,
  group = `Sub-district`,
  data_type = "PR",   
  xrange = c(0, 6500),  
  yrange = c(0, 0.05),
  label = NA,
  title = "Cumulative COVID-19 Fatality Rate by Cumulative Total Number of COVID-19 Positive Cases", #<<           
  x_label = "Cumulative COVID-19 Positive Cases", #<<
  y_label = "Cumulative Fatality Rate"  #<<
)
```
:::

# Funnel Plot for fair visual comparison: ggplot2 methods

::: panel-tabset
## The plot

```{r}
#| echo: false

df <- covid19 %>%
  mutate(rate = Death / Positive) %>%
  mutate(rate.se = sqrt((rate*(1-rate)) / (Positive))) %>%
  filter(rate > 0)
fit.mean <- weighted.mean(df$rate, 1/df$rate.se^2)

number.seq <- seq(1, max(df$Positive), 1)
number.ll95 <- fit.mean - 1.96 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ul95 <- fit.mean + 1.96 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ll999 <- fit.mean - 3.29 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ul999 <- fit.mean + 3.29 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
dfCI <- data.frame(number.ll95, number.ul95, number.ll999, 
                   number.ul999, number.seq, fit.mean)


p <- ggplot(df, aes(x = Positive, y = rate)) +
  geom_point(aes(label=`Sub-district`), 
             alpha=0.4) +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ll95), 
            size = 0.4, 
            colour = "grey40", 
            linetype = "dashed") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ul95), 
            size = 0.4, 
            colour = "grey40", 
            linetype = "dashed") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ll999), 
            size = 0.4, 
            colour = "grey40") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ul999), 
            size = 0.4, 
            colour = "grey40") +
  geom_hline(data = dfCI, 
             aes(yintercept = fit.mean), 
             size = 0.4, 
             colour = "grey40") +
  coord_cartesian(ylim=c(0,0.05)) +
  annotate("text", x = 1, y = -0.13, label = "95%", size = 3, colour = "grey40") + 
  annotate("text", x = 4.5, y = -0.18, label = "99%", size = 3, colour = "grey40") + 
  ggtitle("Cumulative Fatality Rate by Cumulative Number of COVID-19 Cases") +
  xlab("Cumulative Number of COVID-19 Cases") + 
  ylab("Cumulative Fatality Rate") +
  theme_light() +
  theme(plot.title = element_text(size=12),
        legend.position = c(0.91,0.85), 
        legend.title = element_text(size=7),
        legend.text = element_text(size=7),
        legend.background = element_rect(colour = "grey60", linetype = "dotted"),
        legend.key.height = unit(0.3, "cm"))
p


```

## The code

```{r}
#| eval: false

df <- covid19 %>%
  mutate(rate = Death / Positive) %>%
  mutate(rate.se = sqrt((rate*(1-rate)) / (Positive))) %>%
  filter(rate > 0)
fit.mean <- weighted.mean(df$rate, 1/df$rate.se^2)

number.seq <- seq(1, max(df$Positive), 1)
number.ll95 <- fit.mean - 1.96 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ul95 <- fit.mean + 1.96 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ll999 <- fit.mean - 3.29 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ul999 <- fit.mean + 3.29 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
dfCI <- data.frame(number.ll95, number.ul95, number.ll999, 
                   number.ul999, number.seq, fit.mean)

p <- ggplot(df, aes(x = Positive, y = rate)) +
  geom_point(aes(label=`Sub-district`), 
             alpha=0.4) +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ll95), 
            size = 0.4, 
            colour = "grey40", 
            linetype = "dashed") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ul95), 
            size = 0.4, 
            colour = "grey40", 
            linetype = "dashed") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ll999), 
            size = 0.4, 
            colour = "grey40") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ul999), 
            size = 0.4, 
            colour = "grey40") +
  geom_hline(data = dfCI, 
             aes(yintercept = fit.mean), 
             size = 0.4, 
             colour = "grey40") +
  coord_cartesian(ylim=c(0,0.05)) +
  annotate("text", x = 1, y = -0.13, label = "95%", size = 3, colour = "grey40") + 
  annotate("text", x = 4.5, y = -0.18, label = "99%", size = 3, colour = "grey40") + 
  ggtitle("Cumulative Fatality Rate by Cumulative Number of COVID-19 Cases") +
  xlab("Cumulative Number of COVID-19 Cases") + 
  ylab("Cumulative Fatality Rate") +
  theme_light() +
  theme(plot.title = element_text(size=12),
        legend.position = c(0.91,0.85), 
        legend.title = element_text(size=7),
        legend.text = element_text(size=7),
        legend.background = element_rect(colour = "grey60", linetype = "dotted"),
        legend.key.height = unit(0.3, "cm"))
p 

```
:::

## Interactive Funnel Plot: plotly +ggplot2

::: panel-tabset
## The plot

```{r}
#| echo: false

fp_ggplotly <- ggplotly(p,
  tooltip = c("label", 
              "x", 
              "y"))
fp_ggplotly
```

## The code

```{r}
#| eval: false

fp_ggplotly <- ggplotly(p,
  tooltip = c("label", 
              "x", 
              "y"))
fp_ggplotly

```
:::


# Alternatives

## Funnel Plot with highlighted outliers
This classified each sub district as above expected, below expected and within expected range.

::: panel-tabset
## The plot

```{r}
#| echo: false

df <- covid19 %>%
  mutate(rate = Death / Positive) %>%
  filter(Positive > 0, rate > 0)

overall_rate <- sum(df$Death) / sum(df$Positive)

df <- df %>%
  mutate(
    se = sqrt((overall_rate * (1 - overall_rate)) / Positive),
    lower_95 = overall_rate - 1.96 * se,
    upper_95 = overall_rate + 1.96 * se,
    status = case_when(
      rate > upper_95 ~ "Above expected",
      rate < lower_95 ~ "Below expected",
      TRUE ~ "Within expected range"
    )
  )

ggplot(df, aes(x = Positive, y = rate, colour = status)) +
  geom_point(size = 2, alpha = 0.8) +
  geom_hline(yintercept = overall_rate, linetype = "solid") +
  geom_line(aes(y = lower_95), colour = "grey40", linetype = "dashed") +
  geom_line(aes(y = upper_95), colour = "grey40", linetype = "dashed") +
  labs(
    title = "Funnel Plot with Highlighted Outliers",
    subtitle = "Sub-districts outside the 95% control limits are highlighted",
    x = "Cumulative Positive Cases",
    y = "Fatality Rate",
    colour = "Status"
  ) +
  theme_minimal()
```

## The code

```{r}
#| eval: false

df <- covid19 %>%
  mutate(rate = Death / Positive) %>%
  filter(Positive > 0, rate > 0)

overall_rate <- sum(df$Death) / sum(df$Positive)

df <- df %>%
  mutate(
    se = sqrt((overall_rate * (1 - overall_rate)) / Positive),
    lower_95 = overall_rate - 1.96 * se,
    upper_95 = overall_rate + 1.96 * se,
    status = case_when(
      rate > upper_95 ~ "Above expected",
      rate < lower_95 ~ "Below expected",
      TRUE ~ "Within expected range"
    )
  )

ggplot(df, aes(x = Positive, y = rate, colour = status)) +
  geom_point(size = 2, alpha = 0.8) +
  geom_hline(yintercept = overall_rate, linetype = "solid") +
  geom_line(aes(y = lower_95), colour = "grey40", linetype = "dashed") +
  geom_line(aes(y = upper_95), colour = "grey40", linetype = "dashed") +
  labs(
    title = "Funnel Plot with Highlighted Outliers",
    subtitle = "Sub-districts outside the 95% control limits are highlighted",
    x = "Cumulative Positive Cases",
    y = "Fatality Rate",
    colour = "Status"
  ) +
  theme_minimal()

```
:::

## Interactive labelled funnel plot


::: panel-tabset
## The plot

```{r}
#| echo: false

df <- covid19 %>%
  mutate(rate = Death / Positive) %>%
  filter(Positive > 0, rate > 0)

overall_rate <- sum(df$Death) / sum(df$Positive)

df <- df %>%
  mutate(
    se = sqrt((overall_rate * (1 - overall_rate)) / Positive),
    lower_95 = overall_rate - 1.96 * se,
    upper_95 = overall_rate + 1.96 * se,
    tooltip = paste(
      "Sub-district:", `Sub-district`,
      "<br>Positive cases:", Positive,
      "<br>Deaths:", Death,
      "<br>Fatality rate:", round(rate, 4)
    )
  )

p <- ggplot(df, aes(x = Positive, y = rate, text = tooltip)) +
  geom_point(alpha = 0.7) +
  geom_hline(yintercept = overall_rate, colour = "grey30") +
  geom_line(aes(y = lower_95), colour = "grey40", linetype = "dashed") +
  geom_line(aes(y = upper_95), colour = "grey40", linetype = "dashed") +
  labs(
    title = "Interactive Funnel Plot",
    subtitle = "Hover over each point to view sub-district details",
    x = "Cumulative Positive Cases",
    y = "Fatality Rate"
  ) +
  theme_minimal()

ggplotly(p, tooltip = "text")
```

## The code

```{r}
#| eval: false

df <- covid19 %>%
  mutate(rate = Death / Positive) %>%
  filter(Positive > 0, rate > 0)

overall_rate <- sum(df$Death) / sum(df$Positive)

df <- df %>%
  mutate(
    se = sqrt((overall_rate * (1 - overall_rate)) / Positive),
    lower_95 = overall_rate - 1.96 * se,
    upper_95 = overall_rate + 1.96 * se,
    tooltip = paste(
      "Sub-district:", `Sub-district`,
      "<br>Positive cases:", Positive,
      "<br>Deaths:", Death,
      "<br>Fatality rate:", round(rate, 4)
    )
  )

p <- ggplot(df, aes(x = Positive, y = rate, text = tooltip)) +
  geom_point(alpha = 0.7) +
  geom_hline(yintercept = overall_rate, colour = "grey30") +
  geom_line(aes(y = lower_95), colour = "grey40", linetype = "dashed") +
  geom_line(aes(y = upper_95), colour = "grey40", linetype = "dashed") +
  labs(
    title = "Interactive Funnel Plot",
    subtitle = "Hover over each point to view sub-district details",
    x = "Cumulative Positive Cases",
    y = "Fatality Rate"
  ) +
  theme_minimal()

ggplotly(p, tooltip = "text")
```
:::

## Simple funnel plot using FunnelPlotR

::: panel-tabset
## The plot

```{r}
#| echo: false

funnel_plot(
  .data = covid19,
  numerator = Death,
  denominator = Positive,
  group = `Sub-district`,
  data_type = "PR",
  label = NA,
  title = "COVID-19 Fatality Rate by Sub-district",
  x_label = "Positive Cases",
  y_label = "Fatality Rate"
)
```

## The code

```{r}
#| eval: false

funnel_plot(
  .data = covid19,
  numerator = Death,
  denominator = Positive,
  group = `Sub-district`,
  data_type = "PR",
  label = NA,
  title = "COVID-19 Fatality Rate by Sub-district",
  x_label = "Positive Cases",
  y_label = "Fatality Rate"
)

```
:::
