---
title: "Hands-on_Ex04_Visualising_Distribution"
format: html
author: "Persis Evangeline"
date: "May 7, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Getting started

```{r}
pacman::p_load(ggstatsplot, tidyverse)
```

# Data import

```{r}
exam <- read.csv("Exam_data.csv")

```

# 1 sample test

1 sample test is used when we want to determine whether the mean/ distribution of a sample differs significantly from a known or hypothesised value. Instead of relying on visual inspection, the test provides a statistical method to evaluate whether the difference is meaningful. 

::: {.panel-tabset}

## The plot


```{r}
#| echo: false

library(ggstatsplot)

set.seed(1234)

gghistostats(
  data = exam,
  x = ENGLISH,
  type = "bayes",
  test.value = 60,
  xlab = "English scores"
)
```

## The code

```{r}
#| eval: false
library(ggstatsplot)
set.seed(1234)

gghistostats(
  data = exam,
  x = ENGLISH,
  type = "bayes",
  test.value = 60,
  xlab = "English scores"
)
```

:::

# Bayes factor

Bayes factor is a statistical measure used in Bayesian hypothesis testing to compare how well 2 competing hypotheses explain the observed data. There is usually the null and alternative hypothesis. The Bayes Factor is commonly written as the ratio of the likelihood of the data under the alternative hypothesis to the likelihood of the data under the null hypothesis. Bayes factor tells us which hypothesis explains the data better. The Schwarz Criterion is a simple method used to approximate the Bayes Factor. 


# 2 sample mean test: ggbetweenstats()

A two-sample mean test is a statistical test used to determine whether the means of two groups are significantly different from each other.A two-sample mean test evaluates whether the observed difference between these means is large enough to be considered statistically significant, rather than due to random variation.There are 2 common types of 2 sample mean test: independent 2 sample t-test and paired sample t-test.

::: {.panel-tabset}

## The plot


```{r}
#| echo: false
ggbetweenstats(
  data = exam,
  x = GENDER, 
  y = MATHS,
  type = "np",
  messages = FALSE
)

```

## The code

```{r}
#| eval: false

ggbetweenstats(
  data = exam,
  x = GENDER, 
  y = MATHS,
  type = "np",
  messages = FALSE
)

```

:::

# 1 way ANOVA test: ggbetweenstats() method

A one-way ANOVA (Analysis of Variance) is a statistical test used to determine whether the means of three or more groups are significantly different from each other.
It extends the idea of the two-sample t-test.Instead of comparing only 2 groups, ANOVA compares multiple groups simultaneously. If you were to do multiple t-tests instead of ANOVA, it increases the probability of Type I error(false positives). 

::: {.panel-tabset}

## The plot


```{r}
#| echo: false
ggbetweenstats(
  data = exam,
  x = RACE, 
  y = ENGLISH,
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

## The code

```{r}
#| eval: false

ggbetweenstats(
  data = exam,
  x = RACE, 
  y = ENGLISH,
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

:::


# Correlation Analysis
The goal is to determine whether 2 numeric variables are related, how strong the relationship is and whether the relationship is statistically significant. 


::: {.panel-tabset}

## The plot


```{r}
#| echo: false
ggscatterstats(
  data = exam,
  x = MATHS,
  y = ENGLISH,
  marginal = FALSE,
  )
```

## The code

```{r}
#| eval: false

ggscatterstats(
  data = exam,
  x = MATHS,
  y = ENGLISH,
  marginal = FALSE,
  )
```

:::


# Siginicant test of association (dependence), ggbarstats() method

::: {.panel-tabset}

## The plot


```{r}
#| echo: false

exam1 <- exam %>% 
  mutate(MATHS_bins = 
           cut(MATHS, 
               breaks = c(0,60,75,85,100))
)

ggbarstats(exam1, 
           x = MATHS_bins, 
           y = GENDER)
```

## The code

```{r}
#| eval: false

exam1 <- exam %>% 
  mutate(MATHS_bins = 
           cut(MATHS, 
               breaks = c(0,60,75,85,100))
)

ggbarstats(exam1, 
           x = MATHS_bins, 
           y = GENDER)
```

:::

# Alternatives

## Comparing English scores across Classes
This analysis compares English performance across multiple classes

::: {.panel-tabset}

## The plot


```{r}
#| echo: false

ggbetweenstats(
  data = exam,
  x = CLASS,
  y = ENGLISH,
  type = "np",
  pairwise.comparisons = TRUE,
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

## The code

```{r}
#| eval: false

ggbetweenstats(
  data = exam,
  x = CLASS,
  y = ENGLISH,
  type = "np",
  pairwise.comparisons = TRUE,
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

:::

## Correlation between Math and Science scores
This checks if students who perform well in Science also tend to perform well in Maths.

::: {.panel-tabset}

## The plot


```{r}
#| echo: false

ggscatterstats(
  data = exam,
  x = SCIENCE,
  y = MATHS,
  marginal = TRUE,
  type = "parametric",
  conf.level = 0.95,
  messages = FALSE
)
```

## The code

```{r}
#| eval: false

ggscatterstats(
  data = exam,
  x = SCIENCE,
  y = MATHS,
  marginal = TRUE,
  type = "parametric",
  conf.level = 0.95,
  messages = FALSE
)
```

:::

## Association between Race and Gender
This tests whether there is a significant association between Race and Gender. 
It checks whether 2 categorical variables are related. 

::: {.panel-tabset}

## The plot


```{r}
#| echo: false

ggbarstats(
  data = exam,
  x = RACE,
  y = GENDER,
  type = "parametric",
  results.subtitle = TRUE,
  label = "both",
  messages = FALSE
)
```

## The code

```{r}
#| eval: false

ggbarstats(
  data = exam,
  x = RACE,
  y = GENDER,
  type = "parametric",
  results.subtitle = TRUE,
  label = "both",
  messages = FALSE
)
```

:::
