---
title: "Hands-on_Ex04_Visual_Uncertainty"
format: html
author: "Persis Evangeline"
date: "May 7, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Getting started

```{r}
pacman::p_load(plotly, crosstalk, DT, 
               ggdist, ggridges, colorspace,
               gganimate, tidyverse)

```

# Data import

```{r}
exam <- read_csv("Exam_data.csv")

```

# Visualising the uncertainty of point estimates: ggplot2 methods

```{r}
my_sum <- exam %>%
  group_by(RACE) %>%
  summarise(
    n=n(),
    mean=mean(MATHS),
    sd=sd(MATHS)
    ) %>%
  mutate(se=sd/sqrt(n-1))

```

# Displaying table

::: {.panel-tabset}

## The plot


```{r}
#| echo: false

knitr::kable(head(my_sum), format = 'html')
```

## The code

```{r}
#| eval: false

knitr::kable(head(my_sum), format = 'html')
```

:::


# Plotting standard error bars of point estimates

::: {.panel-tabset}

## The plot


```{r}
#| echo: false

ggplot(my_sum) +
  geom_errorbar(
    aes(x=RACE, 
        ymin=mean-se, 
        ymax=mean+se), 
    width=0.2, 
    colour="black", 
    alpha=0.9, 
    linewidth=0.5) +
  geom_point(aes
           (x=RACE, 
            y=mean), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha=1) +
  ggtitle("Standard error of mean maths score by rac")

```

## The code

```{r}
#| eval: false

ggplot(my_sum) +
  geom_errorbar(
    aes(x=RACE, 
        ymin=mean-se, 
        ymax=mean+se), 
    width=0.2, 
    colour="black", 
    alpha=0.9, 
    linewidth=0.5) +
  geom_point(aes
           (x=RACE, 
            y=mean), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha=1) +
  ggtitle("Standard error of mean maths score by rac")

```

:::

# Plotting confidence interval of point estimates

::: {.panel-tabset}

## The plot


```{r}
#| echo: false

ggplot(my_sum) +
  geom_errorbar(
    aes(x=reorder(RACE, -mean), 
        ymin=mean-1.96*se, 
        ymax=mean+1.96*se), 
    width=0.2, 
    colour="black", 
    alpha=0.9, 
    linewidth=0.5) +
  geom_point(aes
           (x=RACE, 
            y=mean), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha=1) +
  labs(x = "Maths score",
       title = "95% confidence interval of mean maths score by race")

```

## The code

```{r}
#| eval: false

ggplot(my_sum) +
  geom_errorbar(
    aes(x=reorder(RACE, -mean), 
        ymin=mean-1.96*se, 
        ymax=mean+1.96*se), 
    width=0.2, 
    colour="black", 
    alpha=0.9, 
    linewidth=0.5) +
  geom_point(aes
           (x=RACE, 
            y=mean), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha=1) +
  labs(x = "Maths score",
       title = "95% confidence interval of mean maths score by race")

```

:::

# Visualising the uncertainty of point estimates with interactive error bars

::: {.panel-tabset}

## The plot


```{r}
#| echo: false

shared_df = SharedData$new(my_sum)

bscols(widths = c(4,8),
       ggplotly((ggplot(shared_df) +
                   geom_errorbar(aes(
                     x=reorder(RACE, -mean),
                     ymin=mean-2.58*se, 
                     ymax=mean+2.58*se), 
                     width=0.2, 
                     colour="black", 
                     alpha=0.9, 
                     size=0.5) +
                   geom_point(aes(
                     x=RACE, 
                     y=mean, 
                     text = paste("Race:", `RACE`, 
                                  "<br>N:", `n`,
                                  "<br>Avg. Scores:", round(mean, digits = 2),
                                  "<br>95% CI:[", 
                                  round((mean-2.58*se), digits = 2), ",",
                                  round((mean+2.58*se), digits = 2),"]")),
                     stat="identity", 
                     color="red", 
                     size = 1.5, 
                     alpha=1) + 
                   xlab("Race") + 
                   ylab("Average Scores") + 
                   theme_minimal() + 
                   theme(axis.text.x = element_text(
                     angle = 45, vjust = 0.5, hjust=1)) +
                   ggtitle("99% Confidence interval of average /<br>maths scores by race")), 
                tooltip = "text"), 
       DT::datatable(shared_df, 
                     rownames = FALSE, 
                     class="compact", 
                     width="100%", 
                     options = list(pageLength = 10,
                                    scrollX=T), 
                     colnames = c("No. of pupils", 
                                  "Avg Scores",
                                  "Std Dev",
                                  "Std Error")) %>%
         formatRound(columns=c('mean', 'sd', 'se'),
                     digits=2))

```

## The code

```{r}
#| eval: false

shared_df = SharedData$new(my_sum)

bscols(widths = c(4,8),
       ggplotly((ggplot(shared_df) +
                   geom_errorbar(aes(
                     x=reorder(RACE, -mean),
                     ymin=mean-2.58*se, 
                     ymax=mean+2.58*se), 
                     width=0.2, 
                     colour="black", 
                     alpha=0.9, 
                     size=0.5) +
                   geom_point(aes(
                     x=RACE, 
                     y=mean, 
                     text = paste("Race:", `RACE`, 
                                  "<br>N:", `n`,
                                  "<br>Avg. Scores:", round(mean, digits = 2),
                                  "<br>95% CI:[", 
                                  round((mean-2.58*se), digits = 2), ",",
                                  round((mean+2.58*se), digits = 2),"]")),
                     stat="identity", 
                     color="red", 
                     size = 1.5, 
                     alpha=1) + 
                   xlab("Race") + 
                   ylab("Average Scores") + 
                   theme_minimal() + 
                   theme(axis.text.x = element_text(
                     angle = 45, vjust = 0.5, hjust=1)) +
                   ggtitle("99% Confidence interval of average /<br>maths scores by race")), 
                tooltip = "text"), 
       DT::datatable(shared_df, 
                     rownames = FALSE, 
                     class="compact", 
                     width="100%", 
                     options = list(pageLength = 10,
                                    scrollX=T), 
                     colnames = c("No. of pupils", 
                                  "Avg Scores",
                                  "Std Dev",
                                  "Std Error")) %>%
         formatRound(columns=c('mean', 'sd', 'se'),
                     digits=2))

```

:::

# Visualising Uncertainty: ggdist package

## stat_pointinterval()

::: {.panel-tabset}

## The plot


```{r}
#| echo: false

exam %>%
  ggplot(aes(x = RACE, 
             y = MATHS)) +
  stat_pointinterval() +
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Mean Point + Multiple-interval plot")

```

## The code

```{r}
#| eval: false

exam %>%
  ggplot(aes(x = RACE, 
             y = MATHS)) +
  stat_pointinterval() +
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Mean Point + Multiple-interval plot")

```

:::

## adding in arguments to above code

::: {.panel-tabset}

## The plot


```{r}
#| echo: false

exam %>%
  ggplot(aes(x = RACE, y = MATHS)) +
  stat_pointinterval(.width = 0.95,
  .point = median,
  .interval = qi) +
  labs(
    title = "Visualising confidence intervals of median math score",
    subtitle = "Median Point + Multiple-interval plot")

```

## The code

```{r}
#| eval: false

exam %>%
  ggplot(aes(x = RACE, y = MATHS)) +
  stat_pointinterval(.width = 0.95,
  .point = median,
  .interval = qi) +
  labs(
    title = "Visualising confidence intervals of median math score",
    subtitle = "Median Point + Multiple-interval plot")
```

:::

## using stat_grradientinterval()

::: {.panel-tabset}

## The plot


```{r}
#| echo: false

exam %>%
  ggplot(aes(x = RACE, 
             y = MATHS)) +
  stat_gradientinterval(   
    fill = "skyblue",      
    show.legend = TRUE     
  ) +                        
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Gradient + interval plot")

```

## The code

```{r}
#| eval: false

exam %>%
  ggplot(aes(x = RACE, 
             y = MATHS)) +
  stat_gradientinterval(   
    fill = "skyblue",      
    show.legend = TRUE     
  ) +                        
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Gradient + interval plot")
```

:::


# Visualising Uncertainty with hypothetical outcome plots (HOPs)

::: {.panel-tabset}

## The plot


```{r}
#| echo: false

devtools::install_github("wilkelab/ungeviz")
library(ungeviz)
ggplot(data = exam, 
       (aes(x = factor(RACE), 
            y = MATHS))) +
  geom_point(position = position_jitter(
    height = 0.3, 
    width = 0.05), 
    size = 0.4, 
    color = "#0072B2", 
    alpha = 1/2) +
  geom_hpline(data = sampler(25, 
                             group = RACE), 
              height = 0.6, 
              color = "#D55E00") +
  theme_bw() + 
  transition_states(.draw, 1, 3)

```

## The code

```{r}
#| eval: false

devtools::install_github("wilkelab/ungeviz")
library(ungeviz)
ggplot(data = exam, 
       (aes(x = factor(RACE), 
            y = MATHS))) +
  geom_point(position = position_jitter(
    height = 0.3, 
    width = 0.05), 
    size = 0.4, 
    color = "#0072B2", 
    alpha = 1/2) +
  geom_hpline(data = sampler(25, 
                             group = RACE), 
              height = 0.6, 
              color = "#D55E00") +
  theme_bw() + 
  transition_states(.draw, 1, 3)
```

:::

# Alternatives

## Violin plot with boxplot overlay
Unlike error bars which only show interval estimates, violin plots reveal: skewness, multimodal distributions, clustering, spread uncertainty. 

::: {.panel-tabset}

## The plot


```{r}
#| echo: false

ggplot(exam,
       aes(x = RACE,
           y = MATHS,
           fill = RACE)) +

  geom_violin(alpha = 0.6,
              trim = FALSE) +

  geom_boxplot(width = 0.15,
               outlier.color = "red",
               alpha = 0.7) +

  labs(
    title = "Distribution and uncertainty of Maths scores by race",
    subtitle = "Violin plot with boxplot overlay"
  ) +

  theme_minimal()

```

## The code

```{r}
#| eval: false

ggplot(exam,
       aes(x = RACE,
           y = MATHS,
           fill = RACE)) +

  geom_violin(alpha = 0.6,
              trim = FALSE) +

  geom_boxplot(width = 0.15,
               outlier.color = "red",
               alpha = 0.7) +

  labs(
    title = "Distribution and uncertainty of Maths scores by race",
    subtitle = "Violin plot with boxplot overlay"
  ) +

  theme_minimal()
```

:::


## Bootstrapped mean confidence intervals

::: {.panel-tabset}

## The plot


```{r}
#| echo: false

ggplot(exam,
       aes(x = RACE,
           y = MATHS,
           color = RACE)) +

  stat_summary(
    fun.data = mean_cl_boot,
    geom = "pointrange"
  ) +

  labs(
    title = "Bootstrapped confidence intervals of Maths scores",
    subtitle = "Mean with bootstrap confidence intervals"
  ) +

  theme_minimal()

```

## The code

```{r}
#| eval: false

ggplot(exam,
       aes(x = RACE,
           y = MATHS,
           color = RACE)) +

  stat_summary(
    fun.data = mean_cl_boot,
    geom = "pointrange"
  ) +

  labs(
    title = "Bootstrapped confidence intervals of Maths scores",
    subtitle = "Mean with bootstrap confidence intervals"
  ) +

  theme_minimal()
```

:::

In the plot above, it appears blank as the points/error bars are probably very small or hidden due to grouping/ aesthetic issues. 
Another improved code will be to split geom from pointrange into point and errorbar. This makes the interval and mean point much more visible.

::: {.panel-tabset}

## The plot


```{r}
#| echo: false

ggplot(exam,
       aes(x = RACE,
           y = MATHS)) +
  stat_summary(
    fun.data = mean_cl_boot,
    geom = "errorbar",
    width = 0.2,
    linewidth = 0.8
  ) +
  stat_summary(
    fun = mean,
    geom = "point",
    size = 3
  ) +
  labs(
    title = "Bootstrapped mean confidence intervals",
    subtitle = "Mean Maths score with bootstrap confidence intervals",
    x = "Race",
    y = "Maths score"
  ) +
  theme_minimal()

```

## The code

```{r}
#| eval: false

ggplot(exam,
       aes(x = RACE,
           y = MATHS)) +
  stat_summary(
    fun.data = mean_cl_boot,
    geom = "errorbar",
    width = 0.2,
    linewidth = 0.8
  ) +
  stat_summary(
    fun = mean,
    geom = "point",
    size = 3
  ) +
  labs(
    title = "Bootstrapped mean confidence intervals",
    subtitle = "Mean Maths score with bootstrap confidence intervals",
    x = "Race",
    y = "Maths score"
  ) +
  theme_minimal()
```

:::

## Half eye plot (ggdist)
The code below creates a half-density + interval plot. It simultaneously shows full distribution, median, credible confidence intervals and uncertainty spread. Compared to traditional error bars, half-eye plots show the full uncertainty distribution, reveal skewness, show density concentration and communicate uncertainty more naturally. 

::: {.panel-tabset}

## The plot


```{r}
#| echo: false

library(ggdist)

exam %>%
  ggplot(aes(x = MATHS,
             y = RACE,
             fill = RACE)) +

  stat_halfeye(
    .width = c(0.50, 0.95),
    point_interval = median_qi
  ) +

  labs(
    title = "Half-eye plot of Maths score uncertainty",
    subtitle = "Density + interval uncertainty visualisation"
  ) +

  theme_minimal()

```

## The code

```{r}
#| eval: false

library(ggdist)

exam %>%
  ggplot(aes(x = MATHS,
             y = RACE,
             fill = RACE)) +

  stat_halfeye(
    .width = c(0.50, 0.95),
    point_interval = median_qi
  ) +

  labs(
    title = "Half-eye plot of Maths score uncertainty",
    subtitle = "Density + interval uncertainty visualisation"
  ) +

  theme_minimal()
```

:::

The differences between the 3 alternatives would be:
1. Violin + boxplot is best for detecting clustering/skewness and its main focus is distribution shape + spread. 
2. Bootstrapped CI is best for non-normal data and its main focus is robust interval estimation. 
3. Half-eye plot is best for Bayesian style uncertainty visualisation and its main focus is density + uncertainty together. 