---
title: "Hands-on Exercise 1"
author: "Persis Evangeline"
date: "April 18, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Getting started

## Install and launching R packages

The code chunk below uses p_load() of pacman package to check if tidyverse packages are installed in the computer. If they are, then they will be launched into R.

```{r}
pacman::p_load(tidyverse)
```

## Importing the data

```{r}
exam_data <- read_csv("Exam_data.csv")
```

# Plotting a simple bar chart

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS))
```

The below code is to plot a bar chart of the distribution by race.

```{r}
ggplot(data=exam_data,
       aes(x=RACE))+
  geom_bar()
```

Plotting histogram using ggplot2.
```{r}
ggplot(data=exam_data, aes(x = MATHS)) +
  geom_histogram(bins=10, 
                 boundary = 100,
                 color="black", 
                 fill="grey") +
  ggtitle("Distribution of Maths scores")
```

Another alternative is to use dotplot. In a dot plot, each dot represents one observation and the dots are stacked. However, the y scale is misleading.

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot(dotsize = 0.5)
```

So, we remove the y axis. 
```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot(binwidth=2.5,         
               dotsize = 0.5) +      
  scale_y_continuous(NULL,           
                     breaks = NULL)  
```

Looking at histograms again:
We can modify a geometric object by changing aes.
the fill() is what differentiates the color.
```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           fill = GENDER)) +
  geom_histogram(bins=20, 
                 color="grey30")
```

When using continuous data, an alternative to histogram would be geom-density().
```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_density()           
```
To plot more than one line:
```{r}
ggplot(data=exam_data, 
       aes(x = MATHS, 
           colour = GENDER)) +
  geom_density()

```

```{r}
ggplot(data=exam_data, 
       aes(x = ENGLISH, 
           colour = CLASS)) +
  geom_density()
```
Boxplot:
```{r}
ggplot(data=exam_data, 
       aes(y = MATHS,       
           x= GENDER)) +    
  geom_boxplot()            
```

Notches in boxplots help to visually determine if the medians of distributions differ. If the notches do not overlap, then the medians
are different. 
```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_boxplot(notch=TRUE)
```

In the above plot,
1. The notches for female and male overlap. This suggests that there is no
statistically significant difference in median Maths scores between genders.
2. Next, by comparing the median lines inside the boxes, we can see that both medians are around the same (mid 70s). This means that central tendency is very similar for both groups. 
3. Look at IQR 
4. Look at whiskers & outliers. 


