---
title: "Hands-on_Ex10"
format: html
date: "11 June, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Getting started

We will be using the following packages in this exercise

```{r}
#| message: false
pacman::p_load(
  lubridate, ggthemes, reactable, reactablefmtr,
  gt, gtExtras, tidyverse
)


```

- **tidyverse** – Collection of R packages for data import, cleaning, transformation, and visualisation (e.g., `dplyr`, `tidyr`, `ggplot2`, `readr`).
- **lubridate** – Simplifies working with dates and times.
- **ggthemes** – Provides additional themes for enhancing `ggplot2` visualisations.
- **gtExtras** – Adds helper functions for creating and styling attractive `gt` tables.
- **reactable** – Creates interactive data tables using the React Table library.
- **reactablefmtr** – Provides easy-to-use formatting and styling tools for `reactable` tables.

# Importing Microsoft access database

## Dataset

For the purpose of this study, a personal database in Microsoft Access mdb format called Coffee Chain will be used.

## Importing database into R

In the code chunk below, odbcConnectAccess() of RODBC package is used used to import a database query table into R.

```{r}
library(RODBC)
con <- odbcConnectAccess2007('Coffee Chain.mdb')
coffeechain <- sqlFetch(con, 'CoffeeChain Query')
write_rds(coffeechain, "CoffeeChain.rds")
odbcClose(con)

```


## Data preparation

The code chunk below is used to import CoffeeChain.rds into R.

```{r}
coffeechain <- read_rds("CoffeeChain.rds")

```

```{r}

product <- coffeechain %>%
  group_by(`Product`) %>%
  summarise(`target` = sum(`Budget Sales`),
            `current` = sum(`Sales`)) %>%
  ungroup()

```


## Bullet Chart

::: panel-tabset
## The plot

```{r}
#| echo: false

ggplot(product, aes(Product, current)) + 
  geom_col(aes(Product, max(target) * 1.01),
           fill="grey85", width=0.85) +
  geom_col(aes(Product, target * 0.75),
           fill="grey60", width=0.85) +
  geom_col(aes(Product, target * 0.5),
           fill="grey50", width=0.85) +
  geom_col(aes(Product, current), 
           width=0.35,
           fill = "black") + 
  geom_errorbar(aes(y = target,
                    x = Product, 
                    ymin = target,
                    ymax= target), 
                width = .4,
                colour = "red",
                size = 1) +
  coord_flip()
```

## The code

```{r}
#| eval: false
ggplot(product, aes(Product, current)) + 
  geom_col(aes(Product, max(target) * 1.01),
           fill="grey85", width=0.85) +
  geom_col(aes(Product, target * 0.75),
           fill="grey60", width=0.85) +
  geom_col(aes(Product, target * 0.5),
           fill="grey50", width=0.85) +
  geom_col(aes(Product, current), 
           width=0.35,
           fill = "black") + 
  geom_errorbar(aes(y = target,
                    x = Product, 
                    ymin = target,
                    ymax= target), 
                width = .4,
                colour = "red",
                size = 1) +
  coord_flip()
```
:::

This is a bullet chart, which is designed to compare actual performance (current) against a target, while providing qualitative performance bands.
The black bar refers to the current value while the red value refers to target marker. The longer the black bar, the higher the current performance. The grey background refer to performance ranges(50%, 75%).the shaded gray regions are just background reference bands to help you compare the current value against the target.

Interpretation:

1. Some of the products which current performance exceed targets are Mint, Lemon, Decaf Espresso.

2. There is only 1 product which current performance meets its target is Caffe Mocha.

3. Some of the products which current performance do not meet its targets are Colombian, Decaf Irish Cream.

# Plotting sparklines using ggplot2

In this section, you will learn how to plot sparklines by using ggplot2.

## Preparing data

```{r}
sales_report <- coffeechain %>%
  filter(Date >= "2013-01-01") %>%
  mutate(Month = month(Date)) %>%
  group_by(Month, Product) %>%
  summarise(Sales = sum(Sales)) %>%
  ungroup() %>%
  select(Month, Product, Sales)

```

The code chunk below is used to compute the minimum, maximum and end othe the month sales.

```{r}
mins <- group_by(sales_report, Product) %>% 
  slice(which.min(Sales))
maxs <- group_by(sales_report, Product) %>% 
  slice(which.max(Sales))
ends <- group_by(sales_report, Product) %>% 
  filter(Month == max(Month))


```

The code chunk below is used to compute the 25 and 75 quantiles.

```{r}
quarts <- sales_report %>%
  group_by(Product) %>%
  summarise(quart1 = quantile(Sales, 
                              0.25),
            quart2 = quantile(Sales, 
                              0.75)) %>%
  right_join(sales_report)

```

## sparklines in ggplot2


::: panel-tabset
## The plot

```{r}
#| echo: false

ggplot(sales_report, aes(x=Month, y=Sales)) + 
  facet_grid(Product ~ ., scales = "free_y") + 
  geom_ribbon(data = quarts, aes(ymin = quart1, max = quart2), 
              fill = 'grey90') +
  geom_line(size=0.3) +
  geom_point(data = mins, col = 'red') +
  geom_point(data = maxs, col = 'blue') +
  geom_text(data = mins, aes(label = Sales), vjust = -1) +
  geom_text(data = maxs, aes(label = Sales), vjust = 2.5) +
  geom_text(data = ends, aes(label = Sales), hjust = 0, nudge_x = 0.5) +
  geom_text(data = ends, aes(label = Product), hjust = 0, nudge_x = 1.0) +
  expand_limits(x = max(sales_report$Month) + 
                  (0.25 * (max(sales_report$Month) - min(sales_report$Month)))) +
  scale_x_continuous(breaks = seq(1, 12, 1)) +
  scale_y_continuous(expand = c(0.1, 0)) +
  theme_tufte(base_size = 3, base_family = "Helvetica") +
  theme(axis.title=element_blank(), axis.text.y = element_blank(), 
        axis.ticks = element_blank(), strip.text = element_blank())
```

## The code

```{r}
#| eval: false
ggplot(sales_report, aes(x=Month, y=Sales)) + 
  facet_grid(Product ~ ., scales = "free_y") + 
  geom_ribbon(data = quarts, aes(ymin = quart1, max = quart2), 
              fill = 'grey90') +
  geom_line(size=0.3) +
  geom_point(data = mins, col = 'red') +
  geom_point(data = maxs, col = 'blue') +
  geom_text(data = mins, aes(label = Sales), vjust = -1) +
  geom_text(data = maxs, aes(label = Sales), vjust = 2.5) +
  geom_text(data = ends, aes(label = Sales), hjust = 0, nudge_x = 0.5) +
  geom_text(data = ends, aes(label = Product), hjust = 0, nudge_x = 1.0) +
  expand_limits(x = max(sales_report$Month) + 
                  (0.25 * (max(sales_report$Month) - min(sales_report$Month)))) +
  scale_x_continuous(breaks = seq(1, 12, 1)) +
  scale_y_continuous(expand = c(0.1, 0)) +
  theme_tufte(base_size = 3, base_family = "Helvetica") +
  theme(axis.title=element_blank(), axis.text.y = element_blank(), 
        axis.ticks = element_blank(), strip.text = element_blank())
```
:::


# Static Information Dashboard Design: gt and gtExtras methods

This section covers the creation of static information dashboards with gt and gtExtras. A basic understanding of their functions from the package documentation is recommended.

## Plotting a simply bullet chart

::: panel-tabset
## The plot

```{r}
#| echo: false
product %>%
  gt::gt() %>%
  gt_plt_bullet(column = current, 
              target = target, 
              width = 60,
              palette = c("lightblue", 
                          "black")) %>%
  gt_theme_538()
```

## The code

```{r}
#| eval: false
product %>%
  gt::gt() %>%
  gt_plt_bullet(column = current, 
              target = target, 
              width = 60,
              palette = c("lightblue", 
                          "black")) %>%
  gt_theme_538()
```
:::

This creates a bullet chart for each row of the table.
The black is the target line, while the blue refers to the bar colour. Some products exceed target such as Chamomile, Darjeeling. Other products such as Regular Espresso and Amaretto meet targets. Some products such as Colombian or Decaf Irish Cream do not meet target.


# sparklines: gtExtras method

```{r}
report <- coffeechain %>%
  mutate(Year = year(Date)) %>%
  filter(Year == "2013") %>%
  mutate (Month = month(Date, 
                        label = TRUE, 
                        abbr = TRUE)) %>%
  group_by(Product, Month) %>%
  summarise(Sales = sum(Sales)) %>%
  ungroup()

```


It is important to note that one of the requirement of gtExtras functions is that almost exclusively they require you to pass data.frame with list columns. In view of this, code chunk below will be used to convert the report data.frame into list columns.


::: panel-tabset
## The plot

```{r}
#| echo: false
report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop")
```

## The code

```{r}
#| eval: false
report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop")
```
:::

## Plotting coffeechain sales report

::: panel-tabset
## The plot

```{r}
#| echo: false

report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop") %>%
   gt() %>%
   gt_plt_sparkline('Monthly Sales',
                    same_limit = FALSE)

report %>% 
  group_by(Product) %>% 
  summarise("Min" = min(Sales, na.rm = T),
            "Max" = max(Sales, na.rm = T),
            "Average" = mean(Sales, na.rm = T)
            ) %>%
  gt() %>%
  fmt_number(columns = 4,
    decimals = 2)
```

## The code

```{r}
#| eval: false

report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop") %>%
   gt() %>%
   gt_plt_sparkline('Monthly Sales',
                    same_limit = FALSE)

report %>% 
  group_by(Product) %>% 
  summarise("Min" = min(Sales, na.rm = T),
            "Max" = max(Sales, na.rm = T),
            "Average" = mean(Sales, na.rm = T)
            ) %>%
  gt() %>%
  fmt_number(columns = 4,
    decimals = 2)
```
:::

This code creates a table where each product gets a tiny sparkline showing how Sales changed over time, and the second block summarizes each product with minimum, maximum, and average sales.

## Combining the dataframe

```{r}

spark <- report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop")


sales <- report %>% 
  group_by(Product) %>% 
  summarise("Min" = min(Sales, na.rm = T),
            "Max" = max(Sales, na.rm = T),
            "Average" = mean(Sales, na.rm = T)
            )


sales_data = left_join(sales, spark)

```

## Plotting the updated data.table


::: panel-tabset
## The plot

```{r}
#| echo: false

sales_data %>%
  gt() %>%
  gt_plt_sparkline('Monthly Sales',
                   same_limit = FALSE)
```

## The code

```{r}
#| eval: false

sales_data %>%
  gt() %>%
  gt_plt_sparkline('Monthly Sales',
                   same_limit = FALSE)
```
:::

## Combining bullet chart and sparklines

::: panel-tabset
## The plot

```{r}
#| echo: false

bullet <- coffeechain %>%
  filter(Date >= "2013-01-01") %>%
  group_by(`Product`) %>%
  summarise(`Target` = sum(`Budget Sales`),
            `Actual` = sum(`Sales`)) %>%
  ungroup() 

sales_data = sales_data %>%
  left_join(bullet)

sales_data %>%
  gt() %>%
  gt_plt_sparkline('Monthly Sales') %>%
  gt_plt_bullet(column = Actual, 
                target = Target, 
                width = 28,
                palette = c("lightblue", 
                          "black")) %>%
  gt_theme_538()

```

## The code

```{r}
#| eval: false

bullet <- coffeechain %>%
  filter(Date >= "2013-01-01") %>%
  group_by(`Product`) %>%
  summarise(`Target` = sum(`Budget Sales`),
            `Actual` = sum(`Sales`)) %>%
  ungroup() 

sales_data = sales_data %>%
  left_join(bullet)

sales_data %>%
  gt() %>%
  gt_plt_sparkline('Monthly Sales') %>%
  gt_plt_bullet(column = Actual, 
                target = Target, 
                width = 28,
                palette = c("lightblue", 
                          "black")) %>%
  gt_theme_538()
```
:::

# Interactive Information Dashboard Design: reactable and reactablefmtr methods

This section introduces creating interactive information dashboards using the reactable and reactablefmtr packages, and encourages reviewing their documentation for an overview of their features.
It also covers installing the dataui package, which is required to build interactive sparklines.


In order to build an interactive sparklines, we need to install dataui R package by using the code chunk below.

```{r}
remotes::install_github("timelyportfolio/dataui")

```

Next, you all need to load the package onto R environment by using the code chunk below.

```{r}

library(dataui)
```

## Plotting interactive sparklines

Similar to gtExtras, to plot an interactive sparklines by using reactablefmtr package we need to prepare the list field by using the code chunk below.

::: panel-tabset
## The plot

```{r}
#| echo: false

remotes::install_github("timelyportfolio/dataui")
library(dataui)
library(reactablefmtr)

report <- report %>%
  group_by(Product) %>%
  summarize(`Monthly Sales` = list(Sales))

reactable(
  report,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(report)
    )
  )
)

```

## The code

```{r}
#| eval: false

remotes::install_github("timelyportfolio/dataui")
library(dataui)
library(reactablefmtr)

report <- report %>%
  group_by(Product) %>%
  summarize(`Monthly Sales` = list(Sales))


reactable(
  report,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(report)
    )
  )
)
```
:::

## Changing the pagesize

::: panel-tabset
## The plot

```{r}
#| echo: false

reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(report)
    )
  )
)

```

## The code

```{r}
#| eval: false
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(report)
    )
  )
)
```
:::

## Adding points and labels

::: panel-tabset
## The plot

```{r}
#| echo: false

reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        labels = c("first", "last")
        )
    )
  )
)

```

## The code

```{r}
#| eval: false

reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        labels = c("first", "last")
        )
    )
  )
)
```
:::

## Adding reference line

::: panel-tabset
## The plot

```{r}
#| echo: false

reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        statline = "mean"
        )
    )
  )
)

```

## The code

```{r}
#| eval: false

reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        statline = "mean"
        )
    )
  )
)
```
:::

## Adding bandline

::: panel-tabset
## The plot

```{r}
#| echo: false

reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        line_width = 1,
        bandline = "innerquartiles",
        bandline_color = "green"
        )
    )
  )
)

```

## The code

```{r}
#| eval: false

reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        line_width = 1,
        bandline = "innerquartiles",
        bandline_color = "green"
        )
    )
  )
)
```
:::

## changing from sparkline to sparkbar

::: panel-tabset
## The plot

```{r}
#| echo: false

reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkbar(
        report,
        highlight_bars = highlight_bars(
          min = "red", max = "blue"),
        bandline = "innerquartiles",
        statline = "mean")
    )
  )
)
```

## The code

```{r}
#| eval: false

reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkbar(
        report,
        highlight_bars = highlight_bars(
          min = "red", max = "blue"),
        bandline = "innerquartiles",
        statline = "mean")
    )
  )
)

```
:::


# Alternatives

## Alternative 1: Interactive sparkline with labels


::: panel-tabset
## The plot

```{r}
#| echo: false

reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        labels = c("first", "last")
      )
    )
  )
)
```

## The code

```{r}
#| eval: false

reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        labels = c("first", "last")
      )
    )
  )
)

```
:::

Each sparkline summarises a product's monthly sales trend, while the first and last values highlight overall changes throughout the year.


## Alternative 2: Sparkline with min and max points

::: panel-tabset
## The plot

```{r}
#| echo: false

reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red",
          max = "blue"
        )
      )
    )
  )
)
```

## The code

```{r}
#| eval: false

reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red",
          max = "blue"
        )
      )
    )
  )
)

```
:::

The highlighted minimum (red) and maximum (blue) points make it easy to identify each product's lowest and highest monthly sales within the reporting period.


## Alternative 3: Changing line colour

::: panel-tabset
## The plot

```{r}
#| echo: false

reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        line_color = "darkgreen"
      )
    )
  )
)

```

## The code

```{r}
#| eval: false


reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        line_color = "darkgreen"
      )
    )
  )
)
```
:::

## Alternative 4: Row highlighting for readability

::: panel-tabset
## The plot

```{r}
#| echo: false

reactable(
  report,
  defaultPageSize = 13,
  striped = TRUE,
  highlight = TRUE,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report
      )
    )
  )
)
```

## The code

```{r}
#| eval: false

reactable(
  report,
  defaultPageSize = 13,
  striped = TRUE,
  highlight = TRUE,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report
      )
    )
  )
)
```
:::

## Alternative 5: Fill the area beneath the sparkline

::: panel-tabset
## The plot

```{r}
#| echo: false

reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkbar(
        report,
        fill_color = "lightblue"
      )
    )
  )
)

```

## The code

```{r}
#| eval: false

reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkbar(
        report,
        fill_color = "lightblue"
      )
    )
  )
)

```
:::

The sparkbars clearly highlight month-to-month variations in sales while enabling quick comparisons across products.

























































