---
title: "Hands-on_Ex08"
format: html
date: "11 June, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Getting started

```{r}
pacman::p_load(scales, viridis, lubridate, ggthemes,
               gridExtra, readxl, knitr, data.table,
               ggHoriPlot, tidyverse)
```


# Importing data

```{r}

attacks <- read_csv("eventlog.csv")
```

# Examining data

```{r}
kable(head(attacks))
```

# Data preparation

## Deriving weekday and hour of day fields

```{r}

make_hr_wkday <- function(ts, sc, tz) {
  real_times <- ymd_hms(ts, 
                        tz = tz[1], 
                        quiet = TRUE)
  dt <- data.table(source_country = sc,
                   wkday = weekdays(real_times),
                   hour = hour(real_times))
  return(dt)
  }


```


## Deriving the attacks kibble data frame

```{r}
wkday_levels <- c('Saturday', 'Friday', 
                  'Thursday', 'Wednesday', 
                  'Tuesday', 'Monday', 
                  'Sunday')

attacks <- attacks %>%
  group_by(tz) %>%
  do(make_hr_wkday(.$timestamp, 
                   .$source_country, 
                   .$tz)) %>% 
  ungroup() %>% 
  mutate(wkday = factor(
    wkday, levels = wkday_levels),
    hour  = factor(
      hour, levels = 0:23))


```

## display the table after processing

::: panel-tabset
## The plot

```{r}
#| echo: false

kable(head(attacks))
```

## The code

```{r}
#| eval: false

kable(head(attacks))

```
:::

# Building calendar heatmap

::: panel-tabset
## The plot

```{r}
#| echo: false

grouped <- attacks %>% 
  count(wkday, hour) %>% 
  ungroup() %>%
  na.omit()

ggplot(grouped, 
       aes(hour, 
           wkday, 
           fill = n)) + 
geom_tile(color = "white", 
          size = 0.1) + 
theme_tufte(base_family = "Helvetica") + 
coord_equal() +
scale_fill_gradient(name = "# of attacks",
                    low = "sky blue", 
                    high = "dark blue") +
labs(x = NULL, 
     y = NULL, 
     title = "Attacks by weekday and time of day") +
theme(axis.ticks = element_blank(),
      plot.title = element_text(hjust = 0.5),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6) )
```

## The code

```{r}
#| eval: false

grouped <- attacks %>% 
  count(wkday, hour) %>% 
  ungroup() %>%
  na.omit()

ggplot(grouped, 
       aes(hour, 
           wkday, 
           fill = n)) + 
geom_tile(color = "white", 
          size = 0.1) + 
theme_tufte(base_family = "Helvetica") + 
coord_equal() +
scale_fill_gradient(name = "# of attacks",
                    low = "sky blue", 
                    high = "dark blue") +
labs(x = NULL, 
     y = NULL, 
     title = "Attacks by weekday and time of day") +
theme(axis.ticks = element_blank(),
      plot.title = element_text(hjust = 0.5),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6) )

```
:::

# Plotting multiple calendar heatmaps

## Derive attack by country object

```{r}
attacks_by_country <- count(
  attacks, source_country) %>%
  mutate(percent = percent(n/sum(n))) %>%
  arrange(desc(n))
```


## Preparing the data frame

```{r}
top4 <- attacks_by_country$source_country[1:4]
top4_attacks <- attacks %>%
  filter(source_country %in% top4) %>%
  count(source_country, wkday, hour) %>%
  ungroup() %>%
  mutate(source_country = factor(
    source_country, levels = top4)) %>%
  na.omit()

```

## Plotting


::: panel-tabset
## The plot

```{r}
#| echo: false

ggplot(top4_attacks, 
       aes(hour, 
           wkday, 
           fill = n)) + 
  geom_tile(color = "white", 
          size = 0.1) + 
  theme_tufte(base_family = "Helvetica") + 
  coord_equal() +
  scale_fill_gradient(name = "# of attacks",
                    low = "sky blue", 
                    high = "dark blue") +
  facet_wrap(~source_country, ncol = 2) +
  labs(x = NULL, y = NULL, 
     title = "Attacks on top 4 countries by weekday and time of day") +
  theme(axis.ticks = element_blank(),
        axis.text.x = element_text(size = 7),
        plot.title = element_text(hjust = 0.5),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6) )
```

## The code

```{r}
#| eval: false

ggplot(top4_attacks, 
       aes(hour, 
           wkday, 
           fill = n)) + 
  geom_tile(color = "white", 
          size = 0.1) + 
  theme_tufte(base_family = "Helvetica") + 
  coord_equal() +
  scale_fill_gradient(name = "# of attacks",
                    low = "sky blue", 
                    high = "dark blue") +
  facet_wrap(~source_country, ncol = 2) +
  labs(x = NULL, y = NULL, 
     title = "Attacks on top 4 countries by weekday and time of day") +
  theme(axis.ticks = element_blank(),
        axis.text.x = element_text(size = 7),
        plot.title = element_text(hjust = 0.5),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6) )

```
:::

# Plotting cycle plot

## Import data

```{r}
air <- read_excel("arrivals_by_air.xlsx")

```


# Deriving data

## Deriving month and year

```{r}
air$month <- factor(month(air$`Month-Year`), 
                    levels=1:12, 
                    labels=month.abb, 
                    ordered=TRUE) 
air$year <- year(ymd(air$`Month-Year`))

```

## Extracting target country

```{r}
Vietnam <- air %>% 
  select(`Vietnam`, 
         month, 
         year) %>%
  filter(year >= 2010)

```

## Computing year average arrivals by month

```{r}
hline.data <- Vietnam %>% 
  group_by(month) %>%
  summarise(avgvalue = mean(`Vietnam`))

```

## Plot

::: panel-tabset
## The plot

```{r}
#| echo: false

ggplot() + 
  geom_line(data=Vietnam,
            aes(x=year, 
                y=`Vietnam`, 
                group=month), 
            colour="black") +
  geom_hline(aes(yintercept=avgvalue), 
             data=hline.data, 
             linetype=6, 
             colour="red", 
             size=0.5) + 
  facet_grid(~month) +
  labs(axis.text.x = element_blank(),
       title = "Visitor arrivals from Vietnam by air, Jan 2010-Dec 2019") +
  xlab("") +
  ylab("No. of Visitors") +
  theme_tufte(base_family = "Helvetica")
```

## The code

```{r}
#| eval: false

ggplot() + 
  geom_line(data=Vietnam,
            aes(x=year, 
                y=`Vietnam`, 
                group=month), 
            colour="black") +
  geom_hline(aes(yintercept=avgvalue), 
             data=hline.data, 
             linetype=6, 
             colour="red", 
             size=0.5) + 
  facet_grid(~month) +
  labs(axis.text.x = element_blank(),
       title = "Visitor arrivals from Vietnam by air, Jan 2010-Dec 2019") +
  xlab("") +
  ylab("No. of Visitors") +
  theme_tufte(base_family = "Helvetica")

```
:::

As the x axis is hard to read, lets modify the code.

::: panel-tabset
## The plot

```{r}
#| echo: false

ggplot() +
  geom_line(
    data = Vietnam,
    aes(
      x = year,
      y = Vietnam,
      group = month
    ),
    colour = "black",
    linewidth = 0.6
  ) +
  
  geom_hline(
    data = hline.data,
    aes(yintercept = avgvalue),
    linetype = "dashed",
    colour = "red",
    linewidth = 0.5
  ) +
  
  facet_grid(~month) +
  
  labs(
    title = "Visitor arrivals from Vietnam by air, Jan 2010–Dec 2019",
    x = NULL,
    y = "No. of Visitors"
  ) +
  
  theme_tufte(base_family = "Helvetica") +
  
  theme(
    plot.title = element_text(
      size = 16,
      face = "bold",
      hjust = 0.5
    ),
    strip.text = element_text(
      size = 11,
      face = "bold"
    ),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.spacing = unit(0.8, "lines")
  )
```

## The code

```{r}
#| eval: false

ggplot() +
  geom_line(
    data = Vietnam,
    aes(
      x = year,
      y = Vietnam,
      group = month
    ),
    colour = "black",
    linewidth = 0.6
  ) +
  
  geom_hline(
    data = hline.data,
    aes(yintercept = avgvalue),
    linetype = "dashed",
    colour = "red",
    linewidth = 0.5
  ) +
  
  facet_grid(~month) +
  
  labs(
    title = "Visitor arrivals from Vietnam by air, Jan 2010–Dec 2019",
    x = NULL,
    y = "No. of Visitors"
  ) +
  
  theme_tufte(base_family = "Helvetica") +
  
  theme(
    plot.title = element_text(
      size = 16,
      face = "bold",
      hjust = 0.5
    ),
    strip.text = element_text(
      size = 11,
      face = "bold"
    ),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.spacing = unit(0.8, "lines")
  )

```
:::


# Plotting Slopegraph

## Import data

```{r}
rice <- read_csv("rice.csv")

```

## Plot

::: panel-tabset
## The plot

```{r}
#| echo: false

rice_slope <- rice %>% 
  mutate(Year = factor(Year)) %>%
  filter(Year %in% c(1961, 1980))

ggplot(rice_slope,
       aes(x = Year,
           y = Yield,
           group = Country,
           colour = Country)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  geom_text(
    data = rice_slope %>% filter(Year == 1961),
    aes(label = Country),
    hjust = 1.1,
    size = 3
  ) +
  geom_text(
    data = rice_slope %>% filter(Year == 1980),
    aes(label = Country),
    hjust = -0.1,
    size = 3
  ) +
  labs(
    title = "Rice Yield of Top 11 Asian Countries",
    subtitle = "1961-1980",
    x = NULL,
    y = "Yield"
  ) +
  theme_tufte(base_family = "Helvetica") +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5))


```

## The code

```{r}
#| eval: false

rice_slope <- rice %>% 
  mutate(Year = factor(Year)) %>%
  filter(Year %in% c(1961, 1980))

ggplot(rice_slope,
       aes(x = Year,
           y = Yield,
           group = Country,
           colour = Country)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  geom_text(
    data = rice_slope %>% filter(Year == 1961),
    aes(label = Country),
    hjust = 1.1,
    size = 3
  ) +
  geom_text(
    data = rice_slope %>% filter(Year == 1980),
    aes(label = Country),
    hjust = -0.1,
    size = 3
  ) +
  labs(
    title = "Rice Yield of Top 11 Asian Countries",
    subtitle = "1961-1980",
    x = NULL,
    y = "Yield"
  ) +
  theme_tufte(base_family = "Helvetica") +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5))

```
:::


# Alternatives

## loading packages

```{r}

pacman::p_load(plotly, DT, ggridges, ggiraph, patchwork)
```

## Interactive heatmap

::: panel-tabset
## The plot

```{r}
#| echo: false

p1 <- ggplot(grouped, 
             aes(x = hour, y = wkday, fill = n,
                 text = paste("Weekday:", wkday,
                              "<br>Hour:", hour,
                              "<br>Attacks:", n))) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "sky blue", high = "dark blue") +
  labs(title = "Interactive Calendar Heatmap",
       x = "Hour of Day",
       y = "Weekday",
       fill = "# Attacks") +
  theme_minimal()

ggplotly(p1, tooltip = "text")


```

## The code

```{r}
#| eval: false

p1 <- ggplot(grouped, 
             aes(x = hour, y = wkday, fill = n,
                 text = paste("Weekday:", wkday,
                              "<br>Hour:", hour,
                              "<br>Attacks:", n))) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "sky blue", high = "dark blue") +
  labs(title = "Interactive Calendar Heatmap",
       x = "Hour of Day",
       y = "Weekday",
       fill = "# Attacks") +
  theme_minimal()

ggplotly(p1, tooltip = "text")
```
:::

Attack activity is concentrated on Wednesdays and Fridays. 

## top attacking countries

::: panel-tabset
## The plot

```{r}
#| echo: false

attacks_by_country %>%
  slice_max(n, n = 10) %>%
  ggplot(aes(x = reorder(source_country, n), 
             y = n,
             fill = n)) +
  geom_col() +
  coord_flip() +
  scale_fill_viridis_c() +
  labs(title = "Top 10 Source Countries by Number of Attacks",
       x = "Source Country",
       y = "Number of Attacks") +
  theme_minimal()


```

## The code

```{r}
#| eval: false

attacks_by_country %>%
  slice_max(n, n = 10) %>%
  ggplot(aes(x = reorder(source_country, n), 
             y = n,
             fill = n)) +
  geom_col() +
  coord_flip() +
  scale_fill_viridis_c() +
  labs(title = "Top 10 Source Countries by Number of Attacks",
       x = "Source Country",
       y = "Number of Attacks") +
  theme_minimal()
```
:::

Attack activity is highly concentrated in few countries, with China and United States dominating the attack sources. 



## ridge plot of attacks

::: panel-tabset
## The plot

```{r}
#| echo: false

library(ggridges)
library(ggrepel)

ridge_data <- attacks %>%
  count(wkday, hour) %>%
  mutate(
    hour_num = as.numeric(as.character(hour)),
    wkday = factor(wkday,
                   levels = c("Monday", "Tuesday", "Wednesday",
                              "Thursday", "Friday", "Saturday", "Sunday"))
  )

ggplot(ridge_data,
       aes(x = hour_num,
           y = wkday,
           height = n,
           group = wkday,
           fill = wkday)) +
  geom_ridgeline(
    alpha = 0.7,
    scale = 0.08
  ) +
  geom_text(
    data = ridge_data %>% group_by(wkday) %>% slice_min(hour_num, n = 1),
    aes(x = -0.8, y = wkday, label = wkday),
    inherit.aes = FALSE,
    hjust = 1,
    size = 3.5
  ) +
  scale_x_continuous(breaks = 0:23, limits = c(-2, 23)) +
  labs(
    title = "Hourly Attack Pattern by Weekday",
    x = "Hour of Day",
    y = NULL
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )

```

## The code

```{r}
#| eval: false

library(ggridges)
library(ggrepel)

ridge_data <- attacks %>%
  count(wkday, hour) %>%
  mutate(
    hour_num = as.numeric(as.character(hour)),
    wkday = factor(wkday,
                   levels = c("Monday", "Tuesday", "Wednesday",
                              "Thursday", "Friday", "Saturday", "Sunday"))
  )

ggplot(ridge_data,
       aes(x = hour_num,
           y = wkday,
           height = n,
           group = wkday,
           fill = wkday)) +
  geom_ridgeline(
    alpha = 0.7,
    scale = 0.08
  ) +
  geom_text(
    data = ridge_data %>% group_by(wkday) %>% slice_min(hour_num, n = 1),
    aes(x = -0.8, y = wkday, label = wkday),
    inherit.aes = FALSE,
    hjust = 1,
    size = 3.5
  ) +
  scale_x_continuous(breaks = 0:23, limits = c(-2, 23)) +
  labs(
    title = "Hourly Attack Pattern by Weekday",
    x = "Hour of Day",
    y = NULL
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```
:::

Here, the y axis labels are not clear. Reason being the attack counts are too large and ridges overlap.

::: panel-tabset
## The plot

```{r}
#| echo: false

attacks %>%
  count(wkday, hour) %>%
  mutate(
    hour = as.numeric(as.character(hour)),
    wkday = factor(wkday,
                   levels = c("Monday", "Tuesday", "Wednesday",
                              "Thursday", "Friday", "Saturday", "Sunday"))
  ) %>%
  ggplot(aes(x = hour, y = n, colour = wkday, group = wkday)) +
  geom_line(linewidth = 1) +
  geom_point(size = 1.5) +
  scale_x_continuous(breaks = 0:23) +
  labs(
    title = "Hourly Attack Pattern by Weekday",
    x = "Hour of Day",
    y = "Number of Attacks",
    colour = "Weekday"
  ) +
  theme_minimal()

```

## The code

```{r}
#| eval: false

attacks %>%
  count(wkday, hour) %>%
  mutate(
    hour = as.numeric(as.character(hour)),
    wkday = factor(wkday,
                   levels = c("Monday", "Tuesday", "Wednesday",
                              "Thursday", "Friday", "Saturday", "Sunday"))
  ) %>%
  ggplot(aes(x = hour, y = n, colour = wkday, group = wkday)) +
  geom_line(linewidth = 1) +
  geom_point(size = 1.5) +
  scale_x_continuous(breaks = 0:23) +
  labs(
    title = "Hourly Attack Pattern by Weekday",
    x = "Hour of Day",
    y = "Number of Attacks",
    colour = "Weekday"
  ) +
  theme_minimal()
```
:::


Friday and Wednesday consistently experience the highest attack activity, especially during the morning hours.  




## interactive slopegraph

::: panel-tabset
## The plot

```{r}
#| echo: false

rice_slope2 <- rice %>% 
  mutate(Year = factor(Year)) %>%
  filter(Year %in% c(1961, 1980))

p2 <- ggplot(rice_slope2,
             aes(x = Year,
                 y = Yield,
                 group = Country,
                 colour = Country)) +
  geom_line_interactive(
    aes(tooltip = Country),
    linewidth = 0.8
  ) +
  geom_point_interactive(
    aes(tooltip = paste(Country, Year, Yield)),
    size = 2
  ) +
  labs(title = "Interactive Rice Yield Slopegraph",
       subtitle = "1961-1980",
       x = NULL,
       y = "Yield") +
  theme_minimal() +
  theme(legend.position = "none")

girafe(ggobj = p2)

```

## The code

```{r}
#| eval: false

rice_slope2 <- rice %>% 
  mutate(Year = factor(Year)) %>%
  filter(Year %in% c(1961, 1980))

p2 <- ggplot(rice_slope2,
             aes(x = Year,
                 y = Yield,
                 group = Country,
                 colour = Country)) +
  geom_line_interactive(
    aes(tooltip = Country),
    linewidth = 0.8
  ) +
  geom_point_interactive(
    aes(tooltip = paste(Country, Year, Yield)),
    size = 2
  ) +
  labs(title = "Interactive Rice Yield Slopegraph",
       subtitle = "1961-1980",
       x = NULL,
       y = "Yield") +
  theme_minimal() +
  theme(legend.position = "none")

girafe(ggobj = p2)
```
:::

All countries experienced an increase in rice yield from 1961 to 1980. However the amount of increase varies significanlty between countries. 


## table for summary

::: panel-tabset
## The plot

```{r}
#| echo: false

attack_summary <- attacks %>%
  count(source_country, wkday, hour) %>%
  arrange(desc(n))

datatable(
  attack_summary,
  options = list(pageLength = 10),
  caption = "Attack Frequency by Country, Weekday and Hour"
)
```

## The code

```{r}
#| eval: false

attack_summary <- attacks %>%
  count(source_country, wkday, hour) %>%
  arrange(desc(n))

datatable(
  attack_summary,
  options = list(pageLength = 10),
  caption = "Attack Frequency by Country, Weekday and Hour"
)
```
:::
























