---
title: "Hands-on_Ex07B"
format: html
author: "Persis Evangeline"
date: "May 15, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Getting started

```{r}
pacman::p_load(tidytext, widyr, wordcloud, DT, ggwordcloud, textplot, lubridate, hms,
tidyverse, tidygraph, ggraph, igraph)

```

# Importing data

```{r}
news20 <- "20news/20news/"

```

## function to read all files from folder into data frame

```{r}
read_folder <- function(infolder) {
  tibble(file = dir(infolder, 
                    full.names = TRUE)) %>%
    mutate(text = map(file, 
                      read_lines)) %>%
    transmute(id = basename(file), 
              text) %>%
    unnest(text)
}

```

## reading all the files from the 20news folder

```{r}
raw_text <- tibble(folder = 
                     dir(news20, 
                         full.names = TRUE)) %>%
  mutate(folder_out = map(folder, 
                          read_folder)) %>%
  unnest(cols = c(folder_out)) %>%
  transmute(newsgroup = basename(folder), 
            id, text)

dir.create("rds", showWarnings = FALSE)

write_rds(raw_text, "rds/news20.rds")

```

# EDA

::: panel-tabset
## The plot

```{r}
#| echo: false

raw_text %>%
  group_by(newsgroup) %>%
  summarize(messages = n_distinct(id)) %>%
  ggplot(aes(messages, newsgroup)) +
  geom_col(fill = "lightblue") +
  labs(
    title = "Number of Messages by Newsgroup",
    x = "Number of Messages",
    y = NULL
  ) +
  theme_minimal()
```

## The code

```{r}
#| eval: false

raw_text %>%
  group_by(newsgroup) %>%
  summarize(messages = n_distinct(id)) %>%
  ggplot(aes(messages, newsgroup)) +
  geom_col(fill = "lightblue") +
  labs(
    title = "Number of Messages by Newsgroup",
    x = "Number of Messages",
    y = NULL
  ) +
  theme_minimal()

```
:::

# tidytext
tidytext applies tidy data principles to text mining and natural language processing. It allows text data to be manipulated using familair tidyverse packages.

The workflow of tidytext is as such:

1. text data

2. convert to tidy text

3. summarise text

4. visualise results


## removing header & automated email signatures


```{r}


cleaned_text <- raw_text %>%
  group_by(newsgroup, id) %>%
  filter(cumsum(text == "") > 0,
         cumsum(str_detect(
           text, "^--")) == 0) %>%
  ungroup()
```


## removing lines with nested text representing quotes from other users



```{r}

cleaned_text <- cleaned_text %>%
  filter(str_detect(text, "^[^>]+[A-Za-z\\d]")
         | text == "",
         !str_detect(text, 
                     "writes(:|\\.\\.\\.)$"),
         !str_detect(text, 
                     "^In article <")
  )
```



## text data processing


::: panel-tabset
## The plot

```{r}
#| echo: false

usenet_words <- cleaned_text %>%
  unnest_tokens(word, text) %>%
  filter(str_detect(word, "[a-z']$"),
         !word %in% stop_words$word)
usenet_words %>%
  count(word, sort = TRUE)
```

## The code

```{r}
#| eval: false

usenet_words <- cleaned_text %>%
  unnest_tokens(word, text) %>%
  filter(str_detect(word, "[a-z']$"),
         !word %in% stop_words$word)
usenet_words %>%
  count(word, sort = TRUE)
```
:::


## visualising words in newsgroups

::: panel-tabset
## The plot

```{r}
#| echo: false

words_by_newsgroup <- usenet_words %>%
  count(newsgroup, word, sort = TRUE)

wordcloud(
  words = words_by_newsgroup$word,
  freq = words_by_newsgroup$n,
  max.words = 300
)
```

## The code

```{r}
#| eval: false


words_by_newsgroup <- usenet_words %>%
  count(newsgroup, word, sort = TRUE)

wordcloud(
  words = words_by_newsgroup$word,
  freq = words_by_newsgroup$n,
  max.words = 300
)
```
:::


## visualising words in newsgroups

::: panel-tabset
## The plot

```{r}
#| echo: false

set.seed(1234)

words_by_newsgroup %>%
  filter(n > 0) %>%
ggplot(aes(label = word,
           size = n)) +
  geom_text_wordcloud() +
  theme_minimal() +
  facet_wrap(~newsgroup)
```

## The code

```{r}
#| eval: false


set.seed(1234)

words_by_newsgroup %>%
  filter(n > 0) %>%
ggplot(aes(label = word,
           size = n)) +
  geom_text_wordcloud() +
  theme_minimal() +
  facet_wrap(~newsgroup)
```
:::

# TF-IDF

tf-idf stands for term frequency-inverse document frequency, which is a numerical statistic that is intended to reflect how important a word is to a document in a collection of corpus.


```{r}
tf_idf <- words_by_newsgroup %>%
  bind_tf_idf(word, newsgroup, n) %>%
  arrange(desc(tf_idf))

```


## Visualising tf-idf

::: panel-tabset
## The plot

```{r}
#| echo: false

DT::datatable(tf_idf, filter = 'top') %>% 
  formatRound(columns = c('tf', 'idf', 
                          'tf_idf'), 
              digits = 3) %>%
  formatStyle(0, 
              target = 'row', 
              lineHeight='25%')
```

## The code

```{r}
#| eval: false


DT::datatable(tf_idf, filter = 'top') %>% 
  formatRound(columns = c('tf', 'idf', 
                          'tf_idf'), 
              digits = 3) %>%
  formatStyle(0, 
              target = 'row', 
              lineHeight='25%')
```
:::


## visualising tf-idf within newsgroups

::: panel-tabset
## The plot

```{r}
#| echo: false

tf_idf %>%
  filter(str_detect(newsgroup, "^sci\\.")) %>%
  group_by(newsgroup) %>%
  slice_max(tf_idf, 
            n = 12) %>%
  ungroup() %>%
  mutate(word = reorder(word, 
                        tf_idf)) %>%
  ggplot(aes(tf_idf, 
             word, 
             fill = newsgroup)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ newsgroup, 
             scales = "free") +
  labs(x = "tf-idf", 
       y = NULL)
```

## The code

```{r}
#| eval: false


tf_idf %>%
  filter(str_detect(newsgroup, "^sci\\.")) %>%
  group_by(newsgroup) %>%
  slice_max(tf_idf, 
            n = 12) %>%
  ungroup() %>%
  mutate(word = reorder(word, 
                        tf_idf)) %>%
  ggplot(aes(tf_idf, 
             word, 
             fill = newsgroup)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ newsgroup, 
             scales = "free") +
  labs(x = "tf-idf", 
       y = NULL)
```
:::


## Counting and correlating pairs of words

Use widyr when you want word co-occurence networks, topic structure exploration, word association analysis and text similarity between terms. 

```{r}
newsgroup_cors <- words_by_newsgroup %>%
  pairwise_cor(newsgroup, 
               word, 
               n, 
               sort = TRUE)

```


## visualising correlation as network

::: panel-tabset
## The plot

```{r}
#| echo: false

set.seed(2017)

newsgroup_cors %>%
  filter(correlation > .025) %>%
  graph_from_data_frame() %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(alpha = correlation, 
                     width = correlation)) +
  geom_node_point(size = 6, 
                  color = "lightblue") +
  geom_node_text(aes(label = name),
                 color = "red",
                 repel = TRUE) +
  theme_void()
```

## The code

```{r}
#| eval: false


set.seed(2017)

newsgroup_cors %>%
  filter(correlation > .025) %>%
  graph_from_data_frame() %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(alpha = correlation, 
                     width = correlation)) +
  geom_node_point(size = 6, 
                  color = "lightblue") +
  geom_node_text(aes(label = name),
                 color = "red",
                 repel = TRUE) +
  theme_void()
```
:::


## Bigram

A bigram is a sequence of 2 consecutive words in a text. For instance, in a sentence "I like to code", the bigrams are " I like", "like to", "to code". 

::: panel-tabset
## The plot

```{r}
#| echo: false

bigrams <- cleaned_text %>%
  unnest_tokens(bigram, 
                text, 
                token = "ngrams", 
                n = 2)
bigrams

```

## The code

```{r}
#| eval: false



bigrams <- cleaned_text %>%
  unnest_tokens(bigram, 
                text, 
                token = "ngrams", 
                n = 2)
bigrams
```
:::


### Counting bigrams

::: panel-tabset
## The plot

```{r}
#| echo: false

bigrams_count <- bigrams %>%
  filter(bigram != 'NA') %>%
  count(bigram, sort = TRUE)
bigrams

```

## The code

```{r}
#| eval: false



bigrams_count <- bigrams %>%
  filter(bigram != 'NA') %>%
  count(bigram, sort = TRUE)
bigrams
```
:::


### Cleaning bigrams

::: panel-tabset
## The plot

```{r}
#| echo: false

bigrams_separated <- bigrams %>%
  filter(bigram != 'NA') %>%
  separate(bigram, c("word1", "word2"), 
           sep = " ")

bigrams_filtered <- bigrams_separated %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)

bigrams_filtered

```

## The code

```{r}
#| eval: false

bigrams_separated <- bigrams %>%
  filter(bigram != 'NA') %>%
  separate(bigram, c("word1", "word2"), 
           sep = " ")

bigrams_filtered <- bigrams_separated %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)

bigrams_filtered
```
:::


### counting again


```{r}

bigram_counts <- bigrams_filtered %>% 
  count(word1, word2, sort = TRUE)
```


### creating network graph from bigram data frame

::: panel-tabset
## The plot

```{r}
#| echo: false

bigram_graph <- bigram_counts %>%
  filter(n > 3) %>%
  graph_from_data_frame()
bigram_graph

```

## The code

```{r}
#| eval: false

bigram_graph <- bigram_counts %>%
  filter(n > 3) %>%
  graph_from_data_frame()
bigram_graph
```
:::

### visualising network of bigrams

::: panel-tabset
## The plot

```{r}
#| echo: false

set.seed(1234)

ggraph(bigram_graph, layout = "fr") +
  geom_edge_link() +
  geom_node_point() +
  geom_node_text(aes(label = name), 
                 vjust = 1, 
                 hjust = 1)

```

## The code

```{r}
#| eval: false

set.seed(1234)

ggraph(bigram_graph, layout = "fr") +
  geom_edge_link() +
  geom_node_point() +
  geom_node_text(aes(label = name), 
                 vjust = 1, 
                 hjust = 1)
```
:::


### revised version

::: panel-tabset
## The plot

```{r}
#| echo: false

set.seed(1234)

a <- grid::arrow(type = "closed", 
                 length = unit(.15,
                               "inches"))

ggraph(bigram_graph, 
       layout = "fr") +
  geom_edge_link(aes(edge_alpha = n), 
                 show.legend = FALSE,
                 arrow = a, 
                 end_cap = circle(.07,
                                  'inches')) +
  geom_node_point(color = "lightblue", 
                  size = 5) +
  geom_node_text(aes(label = name), 
                 vjust = 1, 
                 hjust = 1) +
  theme_void()

```

## The code

```{r}
#| eval: false

set.seed(1234)

a <- grid::arrow(type = "closed", 
                 length = unit(.15,
                               "inches"))

ggraph(bigram_graph, 
       layout = "fr") +
  geom_edge_link(aes(edge_alpha = n), 
                 show.legend = FALSE,
                 arrow = a, 
                 end_cap = circle(.07,
                                  'inches')) +
  geom_node_point(color = "lightblue", 
                  size = 5) +
  geom_node_text(aes(label = name), 
                 vjust = 1, 
                 hjust = 1) +
  theme_void()
```
:::


# Alternatives

## Lollipop chart of top words

::: panel-tabset
## The plot

```{r}
#| echo: false

top_words <- usenet_words %>%
  count(word, sort = TRUE) %>%
  slice_max(n, n = 20) %>%
  mutate(word = reorder(word, n))

ggplot(top_words,
       aes(x = n,
           y = word)) +
  geom_segment(aes(x = 0,
                   xend = n,
                   y = word,
                   yend = word),
               colour = "grey") +
  geom_point(size = 3,
             colour = "purple") +
  labs(
    title = "Top 20 Most Frequent Words",
    x = "Word Count",
    y = NULL
  ) +
  theme_minimal()

```

## The code

```{r}
#| eval: false

top_words <- usenet_words %>%
  count(word, sort = TRUE) %>%
  slice_max(n, n = 20) %>%
  mutate(word = reorder(word, n))

ggplot(top_words,
       aes(x = n,
           y = word)) +
  geom_segment(aes(x = 0,
                   xend = n,
                   y = word,
                   yend = word),
               colour = "grey") +
  geom_point(size = 3,
             colour = "purple") +
  labs(
    title = "Top 20 Most Frequent Words",
    x = "Word Count",
    y = NULL
  ) +
  theme_minimal()
```
:::

## TF-IDF heatmap

::: panel-tabset
## The plot

```{r}
#| echo: false

tf_idf_heatmap <- tf_idf %>%
  group_by(newsgroup) %>%
  slice_max(tf_idf, n = 5) %>%
  ungroup()

ggplot(tf_idf_heatmap,
       aes(x = newsgroup,
           y = reorder(word, tf_idf),
           fill = tf_idf)) +
  geom_tile() +
  labs(
    title = "Top Distinctive Words by Newsgroup",
    x = "Newsgroup",
    y = "Word",
    fill = "TF-IDF"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))

```

## The code

```{r}
#| eval: false

tf_idf_heatmap <- tf_idf %>%
  group_by(newsgroup) %>%
  slice_max(tf_idf, n = 5) %>%
  ungroup()

ggplot(tf_idf_heatmap,
       aes(x = newsgroup,
           y = reorder(word, tf_idf),
           fill = tf_idf)) +
  geom_tile() +
  labs(
    title = "Top Distinctive Words by Newsgroup",
    x = "Newsgroup",
    y = "Word",
    fill = "TF-IDF"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))
```
:::























