---
title: "Take Home Exercise 1"
format:
  html:
    embed-resources: false
author: "Persis Evangeline"
date: "May 20, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# About the dataset

The dataset being used is the Mendeley Data, a detailed dataset of motor insurance policies with coverage-specific financial information. It is a real-world motor vehicle insurance portfolio dataset from a Spanish non-life insurance company. It is particularly valuable as it is an open insurance dataset with real commercial policy and claims data. This data is in spreadsheet format and contains anonymised microdata on automobile insurance policies and covers the main operations of the non-life portfolio from 2022 up till 2024. The database has 354140 records.

```{r}
#| echo: false
library(tibble)
library(gt)

dataset_overview <- tibble(
  Aspect = c(
    "Dataset Type",
    "Industry",
    "Country",
    "Company Type",
    "Time Period",
    "Number of Records",
    "Number of Variables",
    "Unit of Observation",
    "File Format",
    "License",
    "Main Use Cases"
  ),
  
  Details = c(
    "Real-world insurance portfolio data",
    "Motor / automobile insurance",
    "Spain",
    "Non-life insurer",
    "January 2022 – December 2024",
    "~354,140 rows",
    "47 columns",
    "Each row refers to 1 automobile insurance policy transaction(current state/record) for a specific policy in a given year",
    "Spreadsheet / CSV",
    "CC BY 4.0",
    "Risk modelling, pricing, actuarial analytics, fraud analysis, retention analysis, claims modelling"
  )
)

dataset_overview %>%
  gt() %>%
  tab_header(
    title = "Overview of the Dataset"
  )
```

# Objective and problem statement

There are several limitations arising from traditional methods used to analyse the insurance industry. Old traditional methods mainly included dashboards as there was lack of advanced analytical capabilities.

The main objective is to gain deeper insights into data distributions and identifying key factors influencing the profitability and volatility of the insurance portfolio. Usually, the root causes of poor profitability and high volatility are unknown. The aim is to investigate trends in underlying loss drivers and refine segmentation and underwriting strategy.

# Key tasks

1.  Apply suitable univariate visual analytics techniques to explore the distributions and characteristics of numerical and categorical variables.

2.  Apply suitable bivariate visual analytics techniques to investigate and statistically validate factors influencing the profitability and volatility of the insurance portfolio.

Carry out the analysis using tidyverse packages, ggplot2 and its extensions.

# Getting started

```{r}
#| echo: false

pacman::p_load(
  tidyverse,
  janitor,
  skimr,
  DT
)
```

# Import Data

```{r}
#| echo: false
#| warning: false
#| message: false

library(DT)

dataset <- read.csv("Dataset.csv")

datatable(
  head(dataset, 50),
  options = list(
    scrollX = TRUE,
    scrollY = "300px",
    pageLength = 10
  ),
  caption = "Preview of first 50 rows only"
)
```

Split the data such that each variable is 1 column

```{r}
#| echo: false
#| warning: false
#| message: false

library(tidyverse)
library(DT)

dataset <- read_delim(
  "Dataset.csv",
  delim = ";",
  quote = "",
  show_col_types = FALSE
)

names(dataset) <- names(dataset) %>%
  str_remove_all('"') %>%
  str_remove_all("^X") %>%
  janitor::make_clean_names()


dataset <- dataset %>%
  mutate(across(where(is.character), ~ str_remove_all(.x, '"')))

write_csv(dataset, "Dataset_Splitted.csv")

datatable(
  head(dataset, 50),
  options = list(
    scrollX = TRUE,
    scrollY = "300px",
    pageLength = 10
  ),
  caption = "Preview of first 50 rows only"
)

```

```{r}
dataset <- read.csv("Dataset_Splitted.csv")

```

# Data Wrangling

Inspecting dataset structure:

```{r}
glimpse(dataset)
```

We can see all the variable names and their respective variable types.

Check for missing values:

```{r}
sum(is.na(dataset))
```

There are total of 1804 records with missing values. Out of 354140, 1804 records have missing values. Since it is only a small proportion, remove rows with missing values.

Calculate percentage of missing rows:

```{r}
print(sum(is.na(dataset))/nrow(dataset) * 100)


```

Since there is only about 0.5% of missing values, these records with missing values can be removed. It has insignificant effect on the analysis.

```{r}

library(tidyverse)

dataset <- dataset %>%
  drop_na()


```

Check for duplicate records:

```{r}
sum(duplicated(dataset))

```

There are no duplicated records.

Doing basic descriptive summary:

::: panel-tabset
```{r}
#| echo: false
#| results: asis

library(tidyverse)
library(gt)

walk(names(dataset), function(var_name) {
  
  x <- dataset[[var_name]]
  
  cat("##", var_name, "\n\n")
  
  if (is.numeric(x)) {
    
    summary_table <- tibble(
      Statistic = c("Minimum", "1st Quartile", "Median", "Mean", "3rd Quartile", "Maximum", "Missing Values"),
      Value = c(
        min(x, na.rm = TRUE),
        quantile(x, 0.25, na.rm = TRUE),
        median(x, na.rm = TRUE),
        mean(x, na.rm = TRUE),
        quantile(x, 0.75, na.rm = TRUE),
        max(x, na.rm = TRUE),
        sum(is.na(x))
      )
    )
    
  } else {
    
    summary_table <- dataset %>%
      count(.data[[var_name]], sort = TRUE) %>%
      mutate(
        Percentage = round(n / sum(n) * 100, 2)
      ) %>%
      rename(
        Category = .data[[var_name]],
        Count = n
      )
  }
  
  print(
    summary_table %>%
      gt()
  )
  
  cat("\n\n")
})
```
:::

Before moving on to feature engineering, it is important to understand key concepts and terminologies of insurance policies.

1.  Insurer and Insured. Insurer: company that provides insurance coverage. Insured: person or entity protected by the insurance policy.

2.  Policy: legal contract between insurer and policyholder. It states conditions, claim limits, what is covered and excluded.

3.  Coverage: The types of risks the insurer agrees to pay for. For instance in motor insurance, there would be risks of theft, accident damage and medical expenses.

4.  Claim: A request for compensation after a loss event. For instance, in the event of a car accident, the insurer investigates whether the event is covered, the claim is genuine and if policy conditions were followed.

5.  Deductible/Excess: The amount the customer must pay before insurance starts paying. Its purpose is to reduce small claims, discourage careless behaviour and lower premiums. For example, repair cost= SGD 5000, deductible= SGD 1000, insurer pays = repair cost - deductible

6.  Sum insured/ coverage limit: maximum amount insurer will pay.

7.  Risk pooling: majority group of people fund the minority who will make claims. Eg. 1 million people are insured, but only 1% make major claims.

8.  Underwriting: Underwriting is the process of evaluating risk. The insurer decides whether to insure someone, how much premium to charge and what conditions to apply.

9.  Premium: Price that is paid for insurance.

10. Municipality type (in dataset): This refers to location where policyholder resides or is registered.

11. Circulation area (in dataset): This refers to where the insured vehicle is mainly driven. It describes the vehicle's operating environment and not ncessarily where the owner lives.

Feature Engineering:

1.  Profitability = total_premium - total_incurred

```{r}
dataset <- dataset %>%
  mutate(
    profitability =
      total_premium - total_incurred
  )

```

2.  Loss ratio = Claims incurred/Premium_earned = total_incurred/total_premium

If Loss ratio \< 1, means profitable. If Loss ratio \> 1, means loss-making

```{r}
dataset <- dataset %>%
  mutate(
    loss_ratio =
      total_incurred / total_premium
  )
```

3.  Claim flag.

If the total_incurred \>0, means claim has been made. Else, no claim.

```{r}
dataset <- dataset %>%
  mutate(
    claim_flag =
      if_else(total_incurred > 0, "Claim", "No Claim")
  )
```

4.  Claim severity.

Insurance losses are usually grouped by severity.

```{r}
dataset <- dataset %>%
  mutate(
    claim_severity =
      case_when(
        total_incurred == 0 ~ "No Claim",
        total_incurred < 1000 ~ "Low",
        total_incurred < 5000 ~ "Medium",
        TRUE ~ "High"
      )
  )

```

5.  Claims per exposure. A policy exposed for 12 months should not be compared directly with a 1 month policy. This measures average claim cost per unit exposure, cost intensity and severity adjusted exposure risk. This is a financial metric. A higher value would mean higher insurer losses per exposure. Lower value would mean low insurer losses per exposure.

Claims per exposure= Total Incurred/ Exposure

```{r}
dataset <- dataset %>%
  mutate(
    incurred_per_exposure =
      total_incurred / total_exposure
  )
```

6.  Driver risk group. Taking age into consideration

```{r}
dataset <- dataset %>%
  mutate(
    driver_risk_group =
      case_when(
        driver_age < 25 ~ "Young",
        driver_age < 60 ~ "Middle-aged",
        TRUE ~ "Senior"
      )
  )
```

7.  Driving experience

The driving experience can be derived by finding the difference between the driver age and the age at which the driving licence was obtained.

```{r}
dataset <- dataset %>%
  mutate(
    driving_experience =
      driver_age - age_driving_licence
  )
```

8.  Vehicle age group

Vehicles which are much older perform and behave differently from newer vehicles.

```{r}
dataset <- dataset %>%
  mutate(
    vehicle_age_group =
      case_when(
        vehicle_age <= 5 ~ "New",
        vehicle_age <= 10 ~ "Moderate",
        TRUE ~ "Old"
      )
  )

```

9.  Claim Frequency

Claim frequency measures how often claims occur and NOT how expensive they are. This is a frequency metric. A higher value would mean claims occur more frequently.

```{r}
dataset <- dataset %>%
  mutate(
    claim_frequency =
      total_claims / total_exposure
  )
```

Saving this updated dataset into a new file:

```{r}
write_csv(dataset, "Dataset_Engineered.csv") 
```

```{r}
dataset <- read.csv("Dataset_Engineered.csv")

```

# Univariate visual analytics

## Distribution of Claim Flag

::: panel-tabset
## The plot

```{r}
#| echo: false
ggplot(data = dataset,
       aes(x = claim_flag)) +
  
  geom_bar(
    color = "grey25",
    fill = "grey90"
  ) +
  
  theme_gray() +
  
  ggtitle("Distribution of Claim Flag")
```

## The code

```{r}
#| eval: false
ggplot(data = dataset,
       aes(x = claim_flag)) +
  
  geom_bar(
    color = "grey25",
    fill = "grey90"
  ) +
  
  theme_gray() +
  
  ggtitle("Distribution of Claim Flag")
```
:::

There is a much higher number of transactions that have made no claim as compared to those which have made claims.

## Counts of each policy type

::: panel-tabset
## The plot

```{r}
#| echo: false
policy_counts <- dataset |>
  dplyr::count(policy_type)

ggplot(policy_counts,
       aes(x = reorder(policy_type, n),
           y = n)) +

  geom_segment(aes(xend = policy_type,
                   y = 0,
                   yend = n),
               colour = "purple",
               linewidth = 1) +

  geom_point(
    colour = "purple",
    size = 4
  ) +

  coord_flip() +

  labs(
    title = "Policy Type Counts",
    x = "Policy Type",
    y = "Count"
  ) +

  theme_minimal()
```

## The code

```{r}
#| eval: false
policy_counts <- dataset |>
  dplyr::count(policy_type)

ggplot(policy_counts,
       aes(x = reorder(policy_type, n),
           y = n)) +

  geom_segment(aes(xend = policy_type,
                   y = 0,
                   yend = n),
               colour = "purple",
               linewidth = 1) +

  geom_point(
    colour = "purple",
    size = 4
  ) +

  coord_flip() +

  labs(
    title = "Policy Type Counts",
    x = "Policy Type",
    y = "Count"
  ) +

  theme_minimal()
```
:::

CC Policy takes up majority of the transactions. TP makes up a small minority of the transactions.

## Counts of vehicle brands

::: panel-tabset
## The plot

```{r}
#| echo: false
#| fig-height: 14
#| fig-width: 10

brand_counts <- dataset %>%
  count(vehicle_brand, sort = TRUE)

ggplot(brand_counts,
       aes(x = reorder(vehicle_brand, n),
           y = n)) +

  geom_segment(aes(xend = vehicle_brand,
                   y = 0,
                   yend = n),
               colour = "purple",
               linewidth = 1) +

  geom_point(
    colour = "purple",
    size = 4
  ) +

  coord_flip() +

  labs(
    title = "Vehicle Brand Counts",
    x = "Vehicle Brand",
    y = "Count"
  ) +

  theme_minimal()
```

## The code

```{r}
#| eval: false
#| fig-height: 14
#| fig-width: 10

brand_counts <- dataset %>%
  count(vehicle_brand, sort = TRUE)

ggplot(brand_counts,
       aes(x = reorder(vehicle_brand, n),
           y = n)) +

  geom_segment(aes(xend = vehicle_brand,
                   y = 0,
                   yend = n),
               colour = "purple",
               linewidth = 1) +

  geom_point(
    colour = "purple",
    size = 4
  ) +

  coord_flip() +

  labs(
    title = "Vehicle Brand Counts",
    x = "Vehicle Brand",
    y = "Count"
  ) +

  theme_minimal()
```
:::

This shows the distribution and ranking of vehicle brands. Firstly, there is strong concentration among certain dominant brands. The top brands dominating the portfolio is "RENAULT", "CITROEN", "PEUGEOT", "VOLKSWAGEN", "FORD", "SEAT" and "OPEL". After these major brands, there is a sharp decline in frequency. This portfolio has a long-tail distribution where a few vehicle brands account for majority of policies while numerous other brands appear infrequently. There is portfolio concentration risk. If these dominant brands experience higher claim rates, are expensive to repair, become targets for theft then insurer exposure becomes concentrated. The concentration of policies among a limited number of vehicle brands may increase portfolio exposure to brand-specific behaviours and underwriting risks. Renault, Citroen and Peugeot are 3 major French automotive brands. The dominance of these 3 French brands likely reflect regional automobile market preferences in Spain, where European vehicle manufacturers hold substantial market share.

## Percentage distribution of each vehicle age group

::: panel-tabset
## The plot

```{r}
#| echo: false
dataset %>%
  count(vehicle_age_group) %>%
  mutate(percent = n / sum(n) * 100) %>%
  ggplot(aes(x = vehicle_age_group, y = percent)) +
  geom_col(fill = "purple") +
  geom_text(aes(label = paste0(round(percent, 1), "%")),
            vjust = -0.3) +
  labs(
    title = "Percentage Distribution of Vehicle Age Groups",
    x = "Vehicle Age Group",
    y = "Percentage"
  ) +
  theme_minimal()

```

## The code

```{r}
#| eval: false
dataset %>%
  count(vehicle_age_group) %>%
  mutate(percent = n / sum(n) * 100) %>%
  ggplot(aes(x = vehicle_age_group, y = percent)) +
  geom_col(fill = "purple") +
  geom_text(aes(label = paste0(round(percent, 1), "%")),
            vjust = -0.3) +
  labs(
    title = "Percentage Distribution of Vehicle Age Groups",
    x = "Vehicle Age Group",
    y = "Percentage"
  ) +
  theme_minimal()

```
:::

The vast majority (91%) of the records are from old cars and new cars take up the smallest portion. Most people who take up insurance from this company, own old cars.

## Pareto analysis of driver risk groups

::: panel-tabset
## The plot

```{r}
#| echo: false
library(dplyr)

risk_counts <- dataset |>
  count(driver_risk_group) |>
  arrange(desc(n)) |>
  mutate(
    cum_pct = cumsum(n) / sum(n)
  )

ggplot(risk_counts,
       aes(x = reorder(driver_risk_group, -n),
           y = n)) +
  
  geom_col(fill = "mediumpurple") +
  
  geom_line(
    aes(y = cum_pct * max(n),
        group = 1),
    colour = "red"
  ) +
  
  geom_point(
    aes(y = cum_pct * max(n)),
    colour = "red"
  ) +
  
  scale_y_continuous(
    sec.axis = sec_axis(
      ~ . / max(risk_counts$n),
      labels = scales::percent
    )
  ) +
  
  coord_flip() +
  
  labs(
    title = "Pareto Analysis of Driver Risk Groups",
    x = "Risk Group",
    y = "Count"
  ) +
  
  theme_minimal()

```

## The code

```{r}
#| eval: false
library(dplyr)

risk_counts <- dataset |>
  count(driver_risk_group) |>
  arrange(desc(n)) |>
  mutate(
    cum_pct = cumsum(n) / sum(n)
  )

ggplot(risk_counts,
       aes(x = reorder(driver_risk_group, -n),
           y = n)) +
  
  geom_col(fill = "mediumpurple") +
  
  geom_line(
    aes(y = cum_pct * max(n),
        group = 1),
    colour = "red"
  ) +
  
  geom_point(
    aes(y = cum_pct * max(n)),
    colour = "red"
  ) +
  
  scale_y_continuous(
    sec.axis = sec_axis(
      ~ . / max(risk_counts$n),
      labels = scales::percent
    )
  ) +
  
  coord_flip() +
  
  labs(
    title = "Pareto Analysis of Driver Risk Groups",
    x = "Risk Group",
    y = "Count"
  ) +
  
  theme_minimal()

```
:::

The red line in the Pareto Chart represents the cumulative percentage distribution. It shows the running cumulative proportion of the portfolio contributed by each driver risk group. Based on the chart, we can infer that the portfolio is heavily concentrated in middle-aged drivers. The senior drivers form a much smaller segment and young drivers are very minimally represented.

## Proportion of claim severity among claim records

::: panel-tabset
## The plot

```{r}
#| echo: false
library(tidyverse)
library(scales)

claim_severity_counts <- dataset %>%
  
  # Remove records with no claims
  filter(claim_severity != "No Claim") %>%
  
  # Count each severity category
  count(claim_severity) %>%
  
  # Calculate proportions
  mutate(
    proportion = n / sum(n),
    
    # Arrange severity levels in order
    claim_severity = factor(
      claim_severity,
      levels = c("Low", "Medium", "High")
    )
  )

ggplot(claim_severity_counts,
       aes(x = claim_severity,
           y = proportion,
           fill = claim_severity)) +
  
  # Create bars
  geom_col(width = 0.7) +
  
  # Add percentage labels
  geom_text(
    aes(
      label = percent(proportion, accuracy = 0.1)
    ),
    vjust = -0.5,
    size = 5
  ) +
  
  # Convert y-axis to percentages
  scale_y_continuous(
    labels = percent_format(),
    expand = expansion(mult = c(0, 0.1))
  ) +
  
  labs(
    title = "Proportion of Claim Severity Among Claim Records",
    x = "Claim Severity",
    y = "Proportion"
  ) +
  
  theme_minimal() +
  
  theme(
    legend.position = "none",
    plot.title = element_text(
      face = "bold",
      size = 14
    )
  )


```

## The code

```{r}
#| eval: false
library(tidyverse)
library(scales)

claim_severity_counts <- dataset %>%
  
  # Remove records with no claims
  filter(claim_severity != "No Claim") %>%
  
  # Count each severity category
  count(claim_severity) %>%
  
  # Calculate proportions
  mutate(
    proportion = n / sum(n),
    
    # Arrange severity levels in order
    claim_severity = factor(
      claim_severity,
      levels = c("Low", "Medium", "High")
    )
  )

ggplot(claim_severity_counts,
       aes(x = claim_severity,
           y = proportion,
           fill = claim_severity)) +
  
  # Create bars
  geom_col(width = 0.7) +
  
  # Add percentage labels
  geom_text(
    aes(
      label = percent(proportion, accuracy = 0.1)
    ),
    vjust = -0.5,
    size = 5
  ) +
  
  # Convert y-axis to percentages
  scale_y_continuous(
    labels = percent_format(),
    expand = expansion(mult = c(0, 0.1))
  ) +
  
  labs(
    title = "Proportion of Claim Severity Among Claim Records",
    x = "Claim Severity",
    y = "Proportion"
  ) +
  
  theme_minimal() +
  
  theme(
    legend.position = "none",
    plot.title = element_text(
      face = "bold",
      size = 14
    )
  )


```
:::

After filtering out records where claims have been made, half of the claims made are of low severity. The next largest proportion are of medium severity. The high severity make up the smallest proportion of all claims. This pattern is common in insurance portfolios.

# Bivariate visual analytics

## Trend of total claims over time

::: panel-tabset
## The plot

```{r}
#| echo: false
library(dplyr)
library(ggplot2)

yearly_claims <- dataset %>%
  group_by(year) %>%
  summarise(
    total_claims_year = sum(total_claims)
  )

label_points <- yearly_claims %>%
  slice(c(1, n()))

ggplot(yearly_claims,
       aes(x = year,
           y = total_claims_year)) +
  
  geom_line(
    colour = "purple",
    linewidth = 1.5
  ) +
  
  geom_point(
    colour = "purple",
    size = 4
  ) +
  
  geom_text(
    data = label_points,
    aes(label = total_claims_year),
    vjust = -1,
    colour = "black",
    fontface = "bold",
    size = 4
  ) +
  
   scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  
  labs(
    title = "Trend of Total Claims Over Time",
    x = "Year",
    y = "Total Claims"
  ) +
  
  theme_minimal()


```

## The code

```{r}
#| eval: false
library(dplyr)
library(ggplot2)

yearly_claims <- dataset %>%
  group_by(year) %>%
  summarise(
    total_claims_year = sum(total_claims)
  )

label_points <- yearly_claims %>%
  slice(c(1, n()))

ggplot(yearly_claims,
       aes(x = year,
           y = total_claims_year)) +
  
  geom_line(
    colour = "purple",
    linewidth = 1.5
  ) +
  
  geom_point(
    colour = "purple",
    size = 4
  ) +
  
  geom_text(
    data = label_points,
    aes(label = total_claims_year),
    vjust = -1,
    colour = "black",
    fontface = "bold",
    size = 4
  ) +
  
   scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  
  labs(
    title = "Trend of Total Claims Over Time",
    x = "Year",
    y = "Total Claims"
  ) +
  
  theme_minimal()


```
:::

There is an increasing trend from 2022 to 2024, reaching a total of 38937 total claims in 2024.

## Trend of total liability claims over time

::: panel-tabset
## The plot

```{r}
#| echo: false
library(dplyr)
library(ggplot2)

yearly_claims <- dataset %>%
  group_by(year) %>%
  summarise(
    total_liability_claims_year = sum(liability_claims)
  )

label_points <- yearly_claims %>%
  slice(c(1, n()))

ggplot(yearly_claims,
       aes(x = year,
           y = total_liability_claims_year)) +
  
  geom_line(
    colour = "purple",
    linewidth = 1.5
  ) +
  
  geom_point(
    colour = "purple",
    size = 4
  ) +
  
  geom_text(
    data = label_points,
    aes(label = total_liability_claims_year),
    vjust = -1,
    colour = "black",
    fontface = "bold",
    size = 4
  ) +
  
   scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  
  labs(
    title = "Trend of Total liability Claims Over Time",
    x = "Year",
    y = "Total Claims"
  ) +
  
  theme_minimal()
```

## The code

```{r}
#| eval: false
library(dplyr)
library(ggplot2)

yearly_claims <- dataset %>%
  group_by(year) %>%
  summarise(
    total_liability_claims_year = sum(liability_claims)
  )

label_points <- yearly_claims %>%
  slice(c(1, n()))

ggplot(yearly_claims,
       aes(x = year,
           y = total_liability_claims_year)) +
  
  geom_line(
    colour = "purple",
    linewidth = 1.5
  ) +
  
  geom_point(
    colour = "purple",
    size = 4
  ) +
  
  geom_text(
    data = label_points,
    aes(label = total_liability_claims_year),
    vjust = -1,
    colour = "black",
    fontface = "bold",
    size = 4
  ) +
  
   scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  
  labs(
    title = "Trend of Total liability Claims Over Time",
    x = "Year",
    y = "Total Claims"
  ) +
  
  theme_minimal()
```
:::

There is an increasing trend from 2022 to 2024.

## Trend of total property claims over time

::: panel-tabset
## The plot

```{r}
#| echo: false
library(dplyr)
library(ggplot2)

yearly_claims <- dataset %>%
  group_by(year) %>%
  summarise(
    total_claims_year = sum(liability_property_claims)
  )

label_points <- yearly_claims %>%
  slice(c(1, n()))

ggplot(yearly_claims,
       aes(x = year,
           y = total_claims_year)) +
  
  geom_line(
    colour = "purple",
    linewidth = 1.5
  ) +
  
  geom_point(
    colour = "purple",
    size = 4
  ) +
  
  geom_text(
    data = label_points,
    aes(label = total_claims_year),
    vjust = -1,
    colour = "black",
    fontface = "bold",
    size = 4
  ) +
  
   scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  
  labs(
    title = "Trend of Total liability property Claims Over Time",
    x = "Year",
    y = "Total Claims"
  ) +
  
  theme_minimal()

```

## The code

```{r}
#| eval: false

library(dplyr)
library(ggplot2)

yearly_claims <- dataset %>%
  group_by(year) %>%
  summarise(
    total_claims_year = sum(liability_property_claims)
  )

label_points <- yearly_claims %>%
  slice(c(1, n()))

ggplot(yearly_claims,
       aes(x = year,
           y = total_claims_year)) +
  
  geom_line(
    colour = "purple",
    linewidth = 1.5
  ) +
  
  geom_point(
    colour = "purple",
    size = 4
  ) +
  
  geom_text(
    data = label_points,
    aes(label = total_claims_year),
    vjust = -1,
    colour = "black",
    fontface = "bold",
    size = 4
  ) +
  
   scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  
  labs(
    title = "Trend of Total liability property Claims Over Time",
    x = "Year",
    y = "Total Claims"
  ) +
  
  theme_minimal()

```
:::

There is an increasing trend in total liability property claims over the years from 2022 to 2024.

## Trend of total liability injury claims over time

::: panel-tabset
## The plot

```{r}
#| echo: false
library(dplyr)
library(ggplot2)

yearly_claims <- dataset %>%
  group_by(year) %>%
  summarise(
    total_claims_year = sum(liability_injury_claims)
  )

label_points <- yearly_claims %>%
  slice(c(1, n()))

ggplot(yearly_claims,
       aes(x = year,
           y = total_claims_year)) +
  
  geom_line(
    colour = "purple",
    linewidth = 1.5
  ) +
  
  geom_point(
    colour = "purple",
    size = 4
  ) +
  
  geom_text(
    data = label_points,
    aes(label = total_claims_year),
    vjust = -1,
    colour = "black",
    fontface = "bold",
    size = 4
  ) +
  
   scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  
  labs(
    title = "Trend of Total liability injury Claims Over Time",
    x = "Year",
    y = "Total Claims"
  ) +
  
  theme_minimal()

```

## The code

```{r}
#| eval: false
library(dplyr)
library(ggplot2)

yearly_claims <- dataset %>%
  group_by(year) %>%
  summarise(
    total_claims_year = sum(liability_injury_claims)
  )

label_points <- yearly_claims %>%
  slice(c(1, n()))

ggplot(yearly_claims,
       aes(x = year,
           y = total_claims_year)) +
  
  geom_line(
    colour = "purple",
    linewidth = 1.5
  ) +
  
  geom_point(
    colour = "purple",
    size = 4
  ) +
  
  geom_text(
    data = label_points,
    aes(label = total_claims_year),
    vjust = -1,
    colour = "black",
    fontface = "bold",
    size = 4
  ) +
  
   scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  
  labs(
    title = "Trend of Total liability injury Claims Over Time",
    x = "Year",
    y = "Total Claims"
  ) +
  
  theme_minimal()

```
:::

The total number of liability injury claims from 2022 to 2024 has increased over the years. There is an increasing trend.

## Trend of total property claims over time

::: panel-tabset
## The plot

```{r}
#| echo: false
library(dplyr)
library(ggplot2)

yearly_claims <- dataset %>%
  group_by(year) %>%
  summarise(
    total_claims_year = sum(property_claims)
  )

label_points <- yearly_claims %>%
  slice(c(1, n()))

ggplot(yearly_claims,
       aes(x = year,
           y = total_claims_year)) +
  
  geom_line(
    colour = "purple",
    linewidth = 1.5
  ) +
  
  geom_point(
    colour = "purple",
    size = 4
  ) +
  
  geom_text(
    data = label_points,
    aes(label = total_claims_year),
    vjust = -1,
    colour = "black",
    fontface = "bold",
    size = 4
  ) +
  
   scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  
  labs(
    title = "Trend of Total property Claims Over Time",
    x = "Year",
    y = "Total Claims"
  ) +
  
  theme_minimal()

```

## The code

```{r}
#| eval: false
library(dplyr)
library(ggplot2)

yearly_claims <- dataset %>%
  group_by(year) %>%
  summarise(
    total_claims_year = sum(property_claims)
  )

label_points <- yearly_claims %>%
  slice(c(1, n()))

ggplot(yearly_claims,
       aes(x = year,
           y = total_claims_year)) +
  
  geom_line(
    colour = "purple",
    linewidth = 1.5
  ) +
  
  geom_point(
    colour = "purple",
    size = 4
  ) +
  
  geom_text(
    data = label_points,
    aes(label = total_claims_year),
    vjust = -1,
    colour = "black",
    fontface = "bold",
    size = 4
  ) +
  
   scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  
  labs(
    title = "Trend of Total property Claims Over Time",
    x = "Year",
    y = "Total Claims"
  ) +
  
  theme_minimal()

```
:::

There is an increasing trend, with the total property claims increasing from 2022 to 2024.

## Trend of total theft claims over time

::: panel-tabset
## The plot

```{r}
#| echo: false
library(dplyr)
library(ggplot2)

yearly_claims <- dataset %>%
  group_by(year) %>%
  summarise(
    total_claims_year = sum(theft_claims)
  )

label_points <- yearly_claims %>%
  slice(c(1, n()))

ggplot(yearly_claims,
       aes(x = year,
           y = total_claims_year)) +
  
  geom_line(
    colour = "purple",
    linewidth = 1.5
  ) +
  
  geom_point(
    colour = "purple",
    size = 4
  ) +
  
  geom_text(
    data = label_points,
    aes(label = total_claims_year),
    vjust = -1,
    colour = "black",
    fontface = "bold",
    size = 4
  ) +
  
   scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  
  labs(
    title = "Trend of Total theft Claims Over Time",
    x = "Year",
    y = "Total Claims"
  ) +
  
  theme_minimal()

```

## The code

```{r}
#| eval: false
library(dplyr)
library(ggplot2)

yearly_claims <- dataset %>%
  group_by(year) %>%
  summarise(
    total_claims_year = sum(theft_claims)
  )

label_points <- yearly_claims %>%
  slice(c(1, n()))

ggplot(yearly_claims,
       aes(x = year,
           y = total_claims_year)) +
  
  geom_line(
    colour = "purple",
    linewidth = 1.5
  ) +
  
  geom_point(
    colour = "purple",
    size = 4
  ) +
  
  geom_text(
    data = label_points,
    aes(label = total_claims_year),
    vjust = -1,
    colour = "black",
    fontface = "bold",
    size = 4
  ) +
  
   scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  
  labs(
    title = "Trend of Total theft Claims Over Time",
    x = "Year",
    y = "Total Claims"
  ) +
  
  theme_minimal()

```
:::

There is an increasing trend, with the total theft claims increasing from 2022 to 2024.

## Trend of total fire claims over time

::: panel-tabset
## The plot

```{r}
#| echo: false
library(dplyr)
library(ggplot2)

yearly_claims <- dataset %>%
  group_by(year) %>%
  summarise(
    total_claims_year = sum(fire_claims)
  )

label_points <- yearly_claims %>%
  slice(c(1, n()))

ggplot(yearly_claims,
       aes(x = year,
           y = total_claims_year)) +
  
  geom_line(
    colour = "purple",
    linewidth = 1.5
  ) +
  
  geom_point(
    colour = "purple",
    size = 4
  ) +
  
  geom_text(
    data = label_points,
    aes(label = total_claims_year),
    vjust = -1,
    colour = "black",
    fontface = "bold",
    size = 4
  ) +
  
   scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  
  labs(
    title = "Trend of Total fire Claims Over Time",
    x = "Year",
    y = "Total Claims"
  ) +
  
  theme_minimal()

```

## The code

```{r}
#| eval: false
library(dplyr)
library(ggplot2)

yearly_claims <- dataset %>%
  group_by(year) %>%
  summarise(
    total_claims_year = sum(fire_claims)
  )

label_points <- yearly_claims %>%
  slice(c(1, n()))

ggplot(yearly_claims,
       aes(x = year,
           y = total_claims_year)) +
  
  geom_line(
    colour = "purple",
    linewidth = 1.5
  ) +
  
  geom_point(
    colour = "purple",
    size = 4
  ) +
  
  geom_text(
    data = label_points,
    aes(label = total_claims_year),
    vjust = -1,
    colour = "black",
    fontface = "bold",
    size = 4
  ) +
  
   scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  
  labs(
    title = "Trend of Total fire Claims Over Time",
    x = "Year",
    y = "Total Claims"
  ) +
  
  theme_minimal()

```
:::

There is a generally increasing trend on the total fire claims from 2022 to 2024.

## Trend of total glass claims over time

::: panel-tabset
## The plot

```{r}
#| echo: false

library(dplyr)
library(ggplot2)

yearly_claims <- dataset %>%
  group_by(year) %>%
  summarise(
    total_claims_year = sum(glass_claims)
  )

label_points <- yearly_claims %>%
  slice(c(1, n()))

ggplot(yearly_claims,
       aes(x = year,
           y = total_claims_year)) +
  
  geom_line(
    colour = "purple",
    linewidth = 1.5
  ) +
  
  geom_point(
    colour = "purple",
    size = 4
  ) +
  
  geom_text(
    data = label_points,
    aes(label = total_claims_year),
    vjust = -1,
    colour = "black",
    fontface = "bold",
    size = 4
  ) +
  
   scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  
  labs(
    title = "Trend of Total glass Claims Over Time",
    x = "Year",
    y = "Total Claims"
  ) +
  
  theme_minimal()

```

## The code

```{r}
#| eval: false


library(dplyr)
library(ggplot2)

yearly_claims <- dataset %>%
  group_by(year) %>%
  summarise(
    total_claims_year = sum(glass_claims)
  )

label_points <- yearly_claims %>%
  slice(c(1, n()))

ggplot(yearly_claims,
       aes(x = year,
           y = total_claims_year)) +
  
  geom_line(
    colour = "purple",
    linewidth = 1.5
  ) +
  
  geom_point(
    colour = "purple",
    size = 4
  ) +
  
  geom_text(
    data = label_points,
    aes(label = total_claims_year),
    vjust = -1,
    colour = "black",
    fontface = "bold",
    size = 4
  ) +
  
   scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  
  labs(
    title = "Trend of Total glass Claims Over Time",
    x = "Year",
    y = "Total Claims"
  ) +
  
  theme_minimal()

```
:::

There is an increasing trend, with total glass claims increasing from 2039 to 5718 in 2024.

## Trend of total legal protection claims over time

::: panel-tabset
## The plot

```{r}
#| echo: false

library(dplyr)
library(ggplot2)

yearly_claims <- dataset %>%
  group_by(year) %>%
  summarise(
    total_claims_year = sum(legal_protection_claims)
  )

label_points <- yearly_claims %>%
  slice(c(1, n()))

ggplot(yearly_claims,
       aes(x = year,
           y = total_claims_year)) +
  
  geom_line(
    colour = "purple",
    linewidth = 1.5
  ) +
  
  geom_point(
    colour = "purple",
    size = 4
  ) +
  
  geom_text(
    data = label_points,
    aes(label = total_claims_year),
    vjust = -1,
    colour = "black",
    fontface = "bold",
    size = 4
  ) +
  
   scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  
  labs(
    title = "Trend of Total legal protection Claims Over Time",
    x = "Year",
    y = "Total Claims"
  ) +
  
  theme_minimal()

```

## The code

```{r}
#| eval: false

library(dplyr)
library(ggplot2)

yearly_claims <- dataset %>%
  group_by(year) %>%
  summarise(
    total_claims_year = sum(legal_protection_claims)
  )

label_points <- yearly_claims %>%
  slice(c(1, n()))

ggplot(yearly_claims,
       aes(x = year,
           y = total_claims_year)) +
  
  geom_line(
    colour = "purple",
    linewidth = 1.5
  ) +
  
  geom_point(
    colour = "purple",
    size = 4
  ) +
  
  geom_text(
    data = label_points,
    aes(label = total_claims_year),
    vjust = -1,
    colour = "black",
    fontface = "bold",
    size = 4
  ) +
  
   scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  
  labs(
    title = "Trend of Total legal protection Claims Over Time",
    x = "Year",
    y = "Total Claims"
  ) +
  
  theme_minimal()

```
:::

There is an increasing trend, where the total legal protection claims increased from 666 in 2022 to 1639 in 2024.

## Trend of total occupants claims over time

::: panel-tabset
## The plot

```{r}
#| echo: false
library(dplyr)
library(ggplot2)

yearly_claims <- dataset %>%
  group_by(year) %>%
  summarise(
    total_claims_year = sum(occupants_claims)
  )

label_points <- yearly_claims %>%
  slice(c(1, n()))

ggplot(yearly_claims,
       aes(x = year,
           y = total_claims_year)) +
  
  geom_line(
    colour = "purple",
    linewidth = 1.5
  ) +
  
  geom_point(
    colour = "purple",
    size = 4
  ) +
  
  geom_text(
    data = label_points,
    aes(label = total_claims_year),
    vjust = -1,
    colour = "black",
    fontface = "bold",
    size = 4
  ) +
  
   scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  
  labs(
    title = "Trend of Total occupants Claims Over Time",
    x = "Year",
    y = "Total Claims"
  ) +
  
  theme_minimal()

```

## The code

```{r}
#| eval: false
library(dplyr)
library(ggplot2)

yearly_claims <- dataset %>%
  group_by(year) %>%
  summarise(
    total_claims_year = sum(occupants_claims)
  )

label_points <- yearly_claims %>%
  slice(c(1, n()))

ggplot(yearly_claims,
       aes(x = year,
           y = total_claims_year)) +
  
  geom_line(
    colour = "purple",
    linewidth = 1.5
  ) +
  
  geom_point(
    colour = "purple",
    size = 4
  ) +
  
  geom_text(
    data = label_points,
    aes(label = total_claims_year),
    vjust = -1,
    colour = "black",
    fontface = "bold",
    size = 4
  ) +
  
   scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  
  labs(
    title = "Trend of Total occupants Claims Over Time",
    x = "Year",
    y = "Total Claims"
  ) +
  
  theme_minimal()

```
:::

There is generally an increasing trend from 2022 to 2024, where the total occupants claim increased from 13 to 18. There was a very minimal decline in 2023, where the total occupants claim decreased from 13 to 12.

## Trend of all claims by Claim type

::: panel-tabset
## The plot

```{r}
#| echo: false
library(dplyr)
library(tidyr)
library(ggplot2)

claims_long <- dataset %>%
  select(
    year,
    liability_claims,
    liability_property_claims,
    liability_injury_claims,
    property_claims,
    theft_claims,
    fire_claims,
    glass_claims,
    legal_protection_claims,
    occupants_claims
  ) %>%
  pivot_longer(
    cols = -year,
    names_to = "claim_type",
    values_to = "claims"
  ) %>%
  group_by(year, claim_type) %>%
  summarise(
    total_claims_year = sum(claims, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(claims_long,
       aes(x = year,
           y = total_claims_year,
           group = claim_type)) +
  
  geom_line(
    colour = "purple",
    linewidth = 1
  ) +
  
  geom_point(
    colour = "purple",
    size = 2.5
  ) +
  
  facet_wrap(
    ~ claim_type,
    scales = "free_y"
  ) +
  
  labs(
    title = "Trend of Claims Over Time by Claim Type",
    x = "Year",
    y = "Total Claims"
  ) +
  
  theme_minimal()

```

## The code

```{r}
#| eval: false
library(dplyr)
library(tidyr)
library(ggplot2)

claims_long <- dataset %>%
  select(
    year,
    liability_claims,
    liability_property_claims,
    liability_injury_claims,
    property_claims,
    theft_claims,
    fire_claims,
    glass_claims,
    legal_protection_claims,
    occupants_claims
  ) %>%
  pivot_longer(
    cols = -year,
    names_to = "claim_type",
    values_to = "claims"
  ) %>%
  group_by(year, claim_type) %>%
  summarise(
    total_claims_year = sum(claims, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(claims_long,
       aes(x = year,
           y = total_claims_year,
           group = claim_type)) +
  
  geom_line(
    colour = "purple",
    linewidth = 1
  ) +
  
  geom_point(
    colour = "purple",
    size = 2.5
  ) +
  
  facet_wrap(
    ~ claim_type,
    scales = "free_y"
  ) +
  
  labs(
    title = "Trend of Claims Over Time by Claim Type",
    x = "Year",
    y = "Total Claims"
  ) +
  
  theme_minimal()

```
:::

This is a concise view of the trends for all the different claim types from 2022 to 2024.

## Distribution of loss ratio by policy type

::: panel-tabset
## The plot

```{r}
#| echo: false
ggplot(dataset,
       aes(x = policy_type,
           y = loss_ratio)) +

  geom_boxplot()

```

## The code

```{r}
#| eval: false
ggplot(dataset,
       aes(x = policy_type,
           y = loss_ratio)) +

  geom_boxplot()

```
:::

In the plot above, we can see how loss ratio varies across policy type. From this plot, we can tell that most boxplots are compressed near 0, while there are many extreme points extending upward.The actual box and whiskers become tiny and flattened near 0, making them nearly invisible. This indicates that most policies have low loss ratios and a small number of policies generate very high losses. The distribution of loss ratio is highly right-skewed across all policy types, indicating that most policies are profitable or low-loss, while a small number of policies incur extremely high claim costs. The dots above the whiskers are outliers. COMP_E has the most extreme outlier, with CC and TPG having extreme outliers too. A key limitation of the current plot is that extreme outliers compress the rest of the distribution.

A zoomed in version of the plot:

::: panel-tabset
## The plot

```{r}
#| echo: false
ggplot(dataset,
       aes(x = policy_type,
           y = loss_ratio)) +

  geom_boxplot() +

  scale_y_log10()

```

## The code

```{r}
#| eval: false
ggplot(dataset,
       aes(x = policy_type,
           y = loss_ratio)) +

  geom_boxplot() +

  scale_y_log10()
```
:::

Differences in median is not extremely large. TP appears to have slightly higher median loss ratio.

## Profitability by driver risk group

::: panel-tabset
## The plot

```{r}
#| echo: false
ggplot(dataset,
       aes(x = driver_risk_group,
           y = profitability)) +
  geom_boxplot(outlier.alpha = 0.2) +
  coord_flip() +
  labs(
    title = "Profitability by Driver Risk Group",
    x = "Driver Risk Group",
    y = "Profitability"
  ) +
  theme_minimal()

```

## The code

```{r}
#| eval: false
ggplot(dataset,
       aes(x = driver_risk_group,
           y = profitability)) +
  geom_boxplot(outlier.alpha = 0.2) +
  coord_flip() +
  labs(
    title = "Profitability by Driver Risk Group",
    x = "Driver Risk Group",
    y = "Profitability"
  ) +
  theme_minimal()
```
:::

The above plot is a boxplot comparing the distribution of profitability across different age groups. The profitability values are heavily negative. What can be inferred, is that most profitability values are near 0.There are also many extreme negative values, especially for the senior and middle aged groups. Senior drivers appear to be volatile, which could indicate that senior drivers have larger severe-loss events compared to young drivers.

Zooming in:

::: panel-tabset
## The plot

```{r}
#| echo: false
ggplot(dataset,
       aes(x = driver_risk_group,
           y = profitability)) +

  geom_boxplot(
    outlier.alpha = 0.2
  ) +

  coord_flip() +

  coord_cartesian(ylim = c(-5000, 2000)) +

  labs(
    title = "Profitability by Driver Risk Group",
    x = "Driver Risk Group",
    y = "Profitability"
  ) +

  theme_minimal()

```

## The code

```{r}
#| eval: false
ggplot(dataset,
       aes(x = driver_risk_group,
           y = profitability)) +

  geom_boxplot(
    outlier.alpha = 0.2
  ) +

  coord_flip() +

  coord_cartesian(ylim = c(-5000, 2000)) +

  labs(
    title = "Profitability by Driver Risk Group",
    x = "Driver Risk Group",
    y = "Profitability"
  ) +

  theme_minimal()
```
:::

The median for Young drivers is higher than the middle aged and senior aged drivers. The young age drivers group has much wider Interquartile Range. As there are large range of values ranging downwards and much fewer positive extremes, this shows that the large losses are much more extreme than large profits. Since the boxes themselves are above 0 for all age groups, this shows that mostly, insurers are profitable.

## Distribution of total incurred by claim severity

::: panel-tabset
## The plot

```{r}
#| echo: false
ggplot(dataset %>% filter(claim_severity != "No claim"),
       aes(x = claim_severity,
           y = total_incurred)) +
  geom_boxplot(outlier.alpha = 0.2) +
  labs(
    title = "Total Incurred by Claim Severity",
    x = "Claim Severity",
    y = "Total Incurred"
  ) +
  theme_minimal()

```

## The code

```{r}
#| eval: false
ggplot(dataset %>% filter(claim_severity != "No claim"),
       aes(x = claim_severity,
           y = total_incurred)) +
  geom_boxplot(outlier.alpha = 0.2) +
  labs(
    title = "Total Incurred by Claim Severity",
    x = "Claim Severity",
    y = "Total Incurred"
  ) +
  theme_minimal()
```
:::

The plot shows the "No claim", "Medium", and "Low" have near 0 in total incurred. High claim severity shows much larger number of positive extremes in the value of "Total Incurred".

Zoomed in:

::: panel-tabset
## The plot

```{r}
#| echo: false
ggplot(
  dataset %>%
    filter(claim_severity != "No claim"),

  aes(x = claim_severity,
      y = total_incurred)
) +

  geom_boxplot(
    outlier.shape = NA,
    fill = "mediumpurple",
    alpha = 0.6
  ) +

  coord_cartesian(ylim = c(0, 25000)) +

  labs(
    title = "Total Incurred by Claim Severity",
    x = "Claim Severity",
    y = "Total Incurred"
  ) +

  theme_minimal()

```

## The code

```{r}
#| eval: false
ggplot(
  dataset %>%
    filter(claim_severity != "No claim"),

  aes(x = claim_severity,
      y = total_incurred)
) +

  geom_boxplot(
    outlier.shape = NA,
    fill = "mediumpurple",
    alpha = 0.6
  ) +

  coord_cartesian(ylim = c(0, 25000)) +

  labs(
    title = "Total Incurred by Claim Severity",
    x = "Claim Severity",
    y = "Total Incurred"
  ) +

  theme_minimal()
```
:::

High claim severity has the highest median, while median claim severity has lower median with Low claim having the next lowest and No claim being at 0. High claim severity also has the largest interquartile range.

## Claim frequency by driver risk group

::: panel-tabset
## The plot

```{r}
#| echo: false

library(ggdist)

ggplot(dataset,
       aes(x = driver_risk_group,
           y = claim_frequency)) +

  stat_halfeye(
    adjust = 0.5,
    justification = -0.2,
    .width = 0,
    point_colour = NA
  ) +

  geom_boxplot(
    width = 0.15,
    outlier.shape = NA
  ) +

  coord_flip() +

  coord_cartesian(xlim = c(0, 50)) +

  labs(
    title = "Claim Frequency by Driver Risk Group",
    x = "Driver Risk Group",
    y = "Claim Frequency"
  ) +

  theme_minimal()

```

## The code

```{r}
#| eval: false

library(ggdist)

ggplot(dataset,
       aes(x = driver_risk_group,
           y = claim_frequency)) +

  stat_halfeye(
    adjust = 0.5,
    justification = -0.2,
    .width = 0,
    point_colour = NA
  ) +

  geom_boxplot(
    width = 0.15,
    outlier.shape = NA
  ) +

  coord_flip() +

  coord_cartesian(xlim = c(0, 50)) +

  labs(
    title = "Claim Frequency by Driver Risk Group",
    x = "Driver Risk Group",
    y = "Claim Frequency"
  ) +

  theme_minimal()
```
:::

## Relationship between total premium and loss ratio

::: panel-tabset
## The plot

```{r}
#| echo: false
ggplot(dataset,
       aes(x = total_premium,
           y = loss_ratio)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm",
              se = TRUE,
              linewidth = 0.7) +
  labs(
    title = "Relationship Between Total Premium and Loss Ratio",
    x = "Total Premium",
    y = "Loss Ratio"
  ) +
  theme_minimal()
```

## The code

```{r}
#| eval: false
ggplot(dataset,
       aes(x = total_premium,
           y = loss_ratio)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm",
              se = TRUE,
              linewidth = 0.7) +
  labs(
    title = "Relationship Between Total Premium and Loss Ratio",
    x = "Total Premium",
    y = "Loss Ratio"
  ) +
  theme_minimal()
```
:::

This plot suggests that there is little to no strong linear relationship between total_premium and loss_ratio. Most points are concentrated towards the bottom, with only a few points having loss ratio above 9000 when the total premium is small. At low premium values, loss ratios are spread very widely, with loss ratio values ranging from 0 to 3000, and some extreme values above 9000. At higher total premium values, the loss ratio is concentrated around 0. This means that high premium policies are more stable whiile low premium prices are more volatile.

As outliers are compressing the graph, we will zoom in to see the trend more closely.

::: panel-tabset
## The plot

```{r}
#| echo: false
ggplot(dataset,
       aes(x = total_premium,
           y = loss_ratio)) +
  
  geom_point(
    alpha = 0.2,
    colour = "black"
  ) +
  
  geom_smooth(
    method = "loess",
    se = TRUE,
    colour = "blue",
    linewidth = 1
  ) +
  
  coord_cartesian(
    xlim = c(0, 2000),
    ylim = c(0, 200)
  ) +
  
  labs(
    title = "Relationship Between Total Premium and Loss Ratio",
    x = "Total Premium",
    y = "Loss Ratio"
  ) +
  
  theme_minimal()
```

## The code

```{r}
#| eval: false
ggplot(dataset,
       aes(x = total_premium,
           y = loss_ratio)) +
  
  geom_point(
    alpha = 0.2,
    colour = "black"
  ) +
  
  geom_smooth(
    method = "loess",
    se = TRUE,
    colour = "blue",
    linewidth = 1
  ) +
  
  coord_cartesian(
    xlim = c(0, 2000),
    ylim = c(0, 200)
  ) +
  
  labs(
    title = "Relationship Between Total Premium and Loss Ratio",
    x = "Total Premium",
    y = "Loss Ratio"
  ) +
  
  theme_minimal()
```
:::

## Scatterplot of driver age vs Total incurred claims

::: panel-tabset
## The plot

```{r}
#| echo: false
ggplot(dataset,
       aes(x = driver_age,
           y = total_incurred)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "loess",
              se = TRUE) +
  labs(
    title = "Total Incurred Claims by Driver Age",
    x = "Driver Age",
    y = "Total Incurred"
  ) +
  theme_minimal()
```

## The code

```{r}
#| eval: false
ggplot(dataset,
       aes(x = driver_age,
           y = total_incurred)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "loess",
              se = TRUE) +
  labs(
    title = "Total Incurred Claims by Driver Age",
    x = "Driver Age",
    y = "Total Incurred"
  ) +
  theme_minimal()
```
:::

This scatter plot shows the relationship between driver age and total incurred claims. As points are compressed toward the bottom, most drivers have little to no total incurred claims regardless of age. There are several extreme outliers which represent rare but extremely costly claims. This shows that the insurance losses are driven by small number of severe cases. A small minority of customers can contribute to a large proportion of total payouts by insurer. Generally, the large extreme cases which lead to high incurred values are from drivers aged 35 to 80 years old. Very young (below 20 years old) and very old(above 80 years old) drivers show much less observations. The LOESS trend line suggests that there is no strong linear relationship between driver age and total incurred claims.

Zoomed in version:

::: panel-tabset
## The plot

```{r}
#| echo: false
ggplot(dataset,
       aes(x = driver_age,
           y = total_incurred)) +
  
  geom_point(alpha = 0.2) +
  
  geom_smooth(method = "loess",
              se = TRUE) +
  
  coord_cartesian(ylim = c(0, 2e+05)) +
  
  labs(
    title = "Relationship Between Driver Age and Total Incurred Claims",
    x = "Driver Age",
    y = "Total Incurred"
  ) +
  
  theme_minimal()
```

## The code

```{r}
#| eval: false
ggplot(dataset,
       aes(x = driver_age,
           y = total_incurred)) +
  
  geom_point(alpha = 0.2) +
  
  geom_smooth(method = "loess",
              se = TRUE) +
  
  coord_cartesian(ylim = c(0, 2e+05)) +
  
  labs(
    title = "Relationship Between Driver Age and Total Incurred Claims",
    x = "Driver Age",
    y = "Total Incurred"
  ) +
  
  theme_minimal()
```
:::

## Distribution of profitability by policy type

::: panel-tabset
## The plot

```{r}
#| echo: false
library(ggridges)

ggplot(dataset,
       aes(x = profitability,
           y = policy_type,
           fill = policy_type)) +
  geom_density_ridges(alpha = 0.7,
                      show.legend = FALSE) +
  labs(
    title = "Distribution of Profitability by Policy Type",
    x = "Profitability",
    y = "Policy Type"
  ) +
  theme_ridges()
```

## The code

```{r}
#| eval: false
library(ggridges)

ggplot(dataset,
       aes(x = profitability,
           y = policy_type,
           fill = policy_type)) +
  geom_density_ridges(alpha = 0.7,
                      show.legend = FALSE) +
  labs(
    title = "Distribution of Profitability by Policy Type",
    x = "Profitability",
    y = "Policy Type"
  ) +
  theme_ridges()
```
:::

This is a ridge plot. It shows the distribution of profitability across different policy types. Based on the plot, most profitability values are concentrated near 0. All the different policies have very small gains/losses. Due to the long tails extending left, the distributions are left skewed. This indicates that there are cases which have extremely negative profits. As the different policy types have broadly similar distributions, the policies may not differ much. To analyse further, there is a need to zoom in.

Zoomed in:

::: panel-tabset
## The plot

```{r}
#| echo: false
library(ggplot2)
library(ggridges)

ggplot(dataset,
       aes(x = profitability,
           y = policy_type,
           fill = policy_type)) +
  
  geom_density_ridges(
    alpha = 0.7,
    show.legend = FALSE
  ) +
  
  coord_cartesian(xlim = c(-10000, 10000)) +
  
  labs(
    title = "Distribution of Profitability by Policy Type",
    x = "Profitability",
    y = "Policy Type"
  ) +
  
  theme_ridges()
```

## The code

```{r}
#| eval: false
library(ggplot2)
library(ggridges)

ggplot(dataset,
       aes(x = profitability,
           y = policy_type,
           fill = policy_type)) +
  
  geom_density_ridges(
    alpha = 0.7,
    show.legend = FALSE
  ) +
  
  coord_cartesian(xlim = c(-10000, 10000)) +
  
  labs(
    title = "Distribution of Profitability by Policy Type",
    x = "Profitability",
    y = "Policy Type"
  ) +
  
  theme_ridges()
```
:::

## Heatmap of circulation area by municiality type

::: panel-tabset
## The plot

```{r}
#| echo: false
area_counts <- dataset %>%
  count(municipality_type, circulation_area)

ggplot(area_counts,
       aes(x = circulation_area,
           y = municipality_type,
           fill = n)) +

  geom_tile() +

  geom_text(aes(label = n),
            color = "white",
            size = 4) +

  labs(
    title = "Heatmap of Circulation Area by Municipality Type",
    x = "Circulation Area",
    y = "Municipality Type",
    fill = "Count"
  ) +

  theme_minimal()
```

## The code

```{r}
#| eval: false
library(dplyr)
library(ggplot2)

area_counts <- dataset %>%
  count(municipality_type, circulation_area)

ggplot(area_counts,
       aes(x = circulation_area,
           y = municipality_type,
           fill = n)) +

  geom_tile() +

  geom_text(aes(label = n),
            color = "white",
            size = 4) +

  labs(
    title = "Heatmap of Circulation Area by Municipality Type",
    x = "Circulation Area",
    y = "Municipality Type",
    fill = "Count"
  ) +

  theme_minimal()
```
:::

This heatmap displays the relationship between circulation area and municipality type. The lighter blue indicates higher count. The inland row has the highest counts overall. This suggests that most insurance policies come from inland municipalities. Island municipalities account for the least number of policy counts. Among those that are from coastal areas, cars driven in Urban circulation area dominate.

## Claim frequency distribution by circulation area

::: panel-tabset
## The plot

```{r}
#| echo: false
library(ggplot2)

ggplot(dataset,
       aes(x = circulation_area,
           y = claim_frequency,
           fill = circulation_area)) +

  geom_violin(
    alpha = 0.6,
    trim = TRUE
  ) +

  geom_boxplot(
    width = 0.12,
    outlier.shape = NA,
    alpha = 0.5
  ) +

  coord_cartesian(ylim = c(0, 2)) +

  labs(
    title = "Claim Frequency Distribution by Circulation Area",
    x = "Circulation Area",
    y = "Claim Frequency"
  ) +

  theme_minimal()
```

## The code

```{r}
#| eval: false
library(ggplot2)

ggplot(dataset,
       aes(x = circulation_area,
           y = claim_frequency,
           fill = circulation_area)) +

  geom_violin(
    alpha = 0.6,
    trim = TRUE
  ) +

  geom_boxplot(
    width = 0.12,
    outlier.shape = NA,
    alpha = 0.5
  ) +

  coord_cartesian(ylim = c(0, 2)) +

  labs(
    title = "Claim Frequency Distribution by Circulation Area",
    x = "Circulation Area",
    y = "Claim Frequency"
  ) +

  theme_minimal()
```
:::

This plot shows a violin plot with a boxplot embedded into it. For a violin plot, the width of the violin represents where observations are more concentrated. Most observations are concentrated near 0, meaning most policyholders have very low claim frequency. This is typical and common insurance pattern. As the claim frequency is highly concentrated near 0 for both rural and urban circulation areas, it suggests that people make few claims despite the driving environment. The urban circulation plot appears slightly more spread out while the rural circulation is more concentrated at very low frequencies. The difference in urban driving environments could play a role in this. However, there could be other differentiating factors which are much stronger in explaining claim frequency.

# Insights & Conclusion

-   portfolio is heavily concentrated in CC policies, old vehicles and European car brands.

-   claims activity increased from 2022 to 2024.

-   circulation area alone is not sufficient to predict claims behaviour

-   Most policies are low loss, small number of claims drive huge losses

-   lower premium policies are much more volatile
