---
title: "Hands-on Exercise 3"
author: "Persis Evangeline"
date: "April 30, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Getting started

```{r}
pacman::p_load(ggiraph, plotly, 
               patchwork, DT, tidyverse)
```

# Importing Data
```{r}
exam_data <- read_csv("Exam_data.csv")

```


# Plot an interactive statistical graph 
## using ggiraph package

::: {.panel-tabset}

## The plot

```{r}
#| echo: false

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(
    aes(tooltip = ID),
    stackgroups = TRUE, 
    binwidth = 1, 
    method = "histodot") +
  scale_y_continuous(NULL, 
                     breaks = NULL)
girafe(
  ggobj = p,
  width_svg = 6,
  height_svg = 6*0.618
)
```

## The code

```{r}
#| eval: false
p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(
    aes(tooltip = ID),
    stackgroups = TRUE, 
    binwidth = 1, 
    method = "histodot") +
  scale_y_continuous(NULL, 
                     breaks = NULL)
girafe(
  ggobj = p,
  width_svg = 6,
  height_svg = 6*0.618
)
```

:::

# Displaying more than 1 info on tooltip

::: {.panel-tabset}

## The plot

```{r}
#| echo: false

exam_data$tooltip <- c(paste0(     
  "Name = ", exam_data$ID,         
  "\n Class = ", exam_data$CLASS)) 

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(
    aes(tooltip = exam_data$tooltip), 
    stackgroups = TRUE,
    binwidth = 1,
    method = "histodot") +
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(
  ggobj = p,
  width_svg = 8,
  height_svg = 8*0.618
)
```

## The code

```{r}
#| eval: false
exam_data$tooltip <- c(paste0(     
  "Name = ", exam_data$ID,         
  "\n Class = ", exam_data$CLASS)) 

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(
    aes(tooltip = exam_data$tooltip), 
    stackgroups = TRUE,
    binwidth = 1,
    method = "histodot") +
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(
  ggobj = p,
  width_svg = 8,
  height_svg = 8*0.618
)
```

:::

# Customising tooltip style

## Style 1

::: {.panel-tabset}

## The plot

```{r}
#| echo: false

tooltip_css <- "background-color:white; #<<
font-style:bold; color:black;" #<<

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(tooltip = ID),                   
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(    #<<
    opts_tooltip(    #<<
      css = tooltip_css)) #<<
)                                        
```

## The code

```{r}
#| eval: false
tooltip_css <- "background-color:white; #<<
font-style:bold; color:black;" #<<

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(tooltip = ID),                   
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(    #<<
    opts_tooltip(    #<<
      css = tooltip_css)) #<<
)                                        
```

:::

## Style 2

::: {.panel-tabset}

## The plot

```{r}
#| echo: false
tooltip_css <- "
background-color: #1f2937;
color: #ffffff;
padding: 8px 12px;
border-radius: 6px;
font-size: 14px;
font-weight: bold;
box-shadow: 0px 2px 8px rgba(0,0,0,0.3);
" #<<

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(tooltip = ID),                   
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(    #<<
    opts_tooltip(    #<<
      css = tooltip_css)) #<<
)                                        
```

## The code

```{r}
#| eval: false
tooltip_css <- "
background-color: #1f2937;
color: #ffffff;
padding: 8px 12px;
border-radius: 6px;
font-size: 14px;
font-weight: bold;
box-shadow: 0px 2px 8px rgba(0,0,0,0.3);
" #<<

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(tooltip = ID),                   
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(    #<<
    opts_tooltip(    #<<
      css = tooltip_css)) #<<
)                                        
```

:::


## Style 3

::: {.panel-tabset}

## The plot

```{r}
#| echo: false
tooltip_css <- "
background: linear-gradient(135deg, #ff7e5f, #feb47b);
color: black;
padding: 12px 18px;
border-radius: 999px;
font-size: 16px;
font-weight: 600;
border: 2px solid #ffffff;
box-shadow: 0px 4px 12px rgba(0,0,0,0.25);
text-align: center;
" #<<

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(tooltip = ID),                   
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(    #<<
    opts_tooltip(    #<<
      css = tooltip_css)) #<<
)                                        
```

## The code

```{r}
#| eval: false
tooltip_css <- "
background: linear-gradient(135deg, #ff7e5f, #feb47b);
color: black;
padding: 12px 18px;
border-radius: 999px;
font-size: 16px;
font-weight: 600;
border: 2px solid #ffffff;
box-shadow: 0px 4px 12px rgba(0,0,0,0.25);
text-align: center;
" #<<

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(tooltip = ID),                   
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(    #<<
    opts_tooltip(    #<<
      css = tooltip_css)) #<<
)                                        
```

:::

# Displaying statistics on tooltip

::: {.panel-tabset}

## The plot

```{r}
#| echo: false
tooltip <- function(y, ymax, accuracy = .01) {
  mean <- scales::number(y, accuracy = accuracy)
  sem <- scales::number(ymax - y, accuracy = accuracy)
  paste("Mean maths scores:", mean, "+/-", sem)
}

gg_point <- ggplot(data=exam_data, 
                   aes(x = RACE),
) +
  stat_summary(aes(y = MATHS, 
                   tooltip = after_stat(  
                     tooltip(y, ymax))),  
    fun.data = "mean_se", 
    geom = GeomInteractiveCol,  
    fill = "light blue"
  ) +
  stat_summary(aes(y = MATHS),
    fun.data = mean_se,
    geom = "errorbar", width = 0.2, size = 0.2
  )

girafe(ggobj = gg_point,
       width_svg = 8,
       height_svg = 8*0.618)                                     
```

## The code

```{r}
#| eval: false
tooltip <- function(y, ymax, accuracy = .01) {
  mean <- scales::number(y, accuracy = accuracy)
  sem <- scales::number(ymax - y, accuracy = accuracy)
  paste("Mean maths scores:", mean, "+/-", sem)
}

gg_point <- ggplot(data=exam_data, 
                   aes(x = RACE),
) +
  stat_summary(aes(y = MATHS, 
                   tooltip = after_stat(  
                     tooltip(y, ymax))),  
    fun.data = "mean_se", 
    geom = GeomInteractiveCol,  
    fill = "light blue"
  ) +
  stat_summary(aes(y = MATHS),
    fun.data = mean_se,
    geom = "errorbar", width = 0.2, size = 0.2
  )

girafe(ggobj = gg_point,
       width_svg = 8,
       height_svg = 8*0.618)                                        
```

:::

## Displaying median instead of mean

::: {.panel-tabset}

## The plot

```{r}
#| echo: false
tooltip <- function(y, ymin, ymax, accuracy = .01) {
  median_val <- scales::number(y, accuracy = accuracy)
  paste("Median maths scores:", median_val)
}

gg_point <- ggplot(data = exam_data, 
                   aes(x = RACE)) +

  stat_summary(
    aes(y = MATHS,
        tooltip = after_stat(
          tooltip(y, ymin, ymax)
        )),
    fun = median,                  
    geom = GeomInteractiveCol,
    fill = "lightblue"
  ) +

  stat_summary(
    aes(y = MATHS),
    fun.data = median_hilow,       
    geom = "errorbar",
    width = 0.2,
    size = 0.2
  )

girafe(
  ggobj = gg_point,
  width_svg = 8,
  height_svg = 8 * 0.618
)                                    
```

## The code

```{r}
#| eval: false
tooltip <- function(y, ymin, ymax, accuracy = .01) {
  median_val <- scales::number(y, accuracy = accuracy)
  paste("Median maths scores:", median_val)
}

gg_point <- ggplot(data = exam_data, 
                   aes(x = RACE)) +

  stat_summary(
    aes(y = MATHS,
        tooltip = after_stat(
          tooltip(y, ymin, ymax)
        )),
    fun = median,                  
    geom = GeomInteractiveCol,
    fill = "lightblue"
  ) +

  stat_summary(
    aes(y = MATHS),
    fun.data = median_hilow,       
    geom = "errorbar",
    width = 0.2,
    size = 0.2
  )

girafe(
  ggobj = gg_point,
  width_svg = 8,
  height_svg = 8 * 0.618
)                                      
```

:::

## Displaying standard deviation instead of mean

## Displaying standard deviation

::: {.panel-tabset}

## The plot

```{r}
#| echo: false

mean_sd <- function(x) {
  m <- mean(x, na.rm = TRUE)
  s <- sd(x, na.rm = TRUE)
  data.frame(
    y = m,
    ymin = m - s,
    ymax = m + s
  )
}

tooltip <- function(y, ymin, ymax, accuracy = .01) {
  mean_val <- scales::number(y, accuracy = accuracy)
  sd_val <- scales::number(ymax - y, accuracy = accuracy)
  paste("Mean maths scores:", mean_val, "\nSD:", sd_val)
}

gg_point <- ggplot(data = exam_data, 
                   aes(x = RACE)) +

  stat_summary(
    aes(y = MATHS,
        tooltip = after_stat(
          tooltip(y, ymin, ymax)
        )),
    fun.data = mean_sd,                 
    geom = GeomInteractiveCol,
    fill = "lightblue"
  ) +

  stat_summary(
    aes(y = MATHS),
    fun.data = mean_sd,                 
    geom = "errorbar",
    width = 0.2,
    size = 0.2
  )

girafe(
  ggobj = gg_point,
  width_svg = 8,
  height_svg = 8 * 0.618
)
```

## The code

```{r}
#| eval: false

mean_sd <- function(x) {
  m <- mean(x, na.rm = TRUE)
  s <- sd(x, na.rm = TRUE)
  data.frame(
    y = m,
    ymin = m - s,
    ymax = m + s
  )
}

tooltip <- function(y, ymin, ymax, accuracy = .01) {
  mean_val <- scales::number(y, accuracy = accuracy)
  sd_val <- scales::number(ymax - y, accuracy = accuracy)
  paste("Mean maths scores:", mean_val, "\nSD:", sd_val)
}

gg_point <- ggplot(data = exam_data, 
                   aes(x = RACE)) +

  stat_summary(
    aes(y = MATHS,
        tooltip = after_stat(
          tooltip(y, ymin, ymax)
        )),
    fun.data = mean_sd,                 
    geom = GeomInteractiveCol,
    fill = "lightblue"
  ) +

  stat_summary(
    aes(y = MATHS),
    fun.data = mean_sd,                 
    geom = "errorbar",
    width = 0.2,
    size = 0.2
  )

girafe(
  ggobj = gg_point,
  width_svg = 8,
  height_svg = 8 * 0.618
)
```

:::

# Hover effect
::: {.panel-tabset}

## The plot

```{r}
#| echo: false
p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(           
    aes(data_id = CLASS),             
    stackgroups = TRUE,               
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618                      
)                                        

```

## The code
```{r}
#| eval: false
p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(           
    aes(data_id = CLASS),             
    stackgroups = TRUE,               
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618                      
)                                        

```
:::

# Styling hover effect
::: {.panel-tabset}

## The plot

```{r}
#| echo: false

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(data_id = CLASS),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(                        
    opts_hover(css = "fill: #202020;"),  
    opts_hover_inv(css = "opacity:0.2;") 
  )                                        
)                                        
```
## The code
```{r}
#| eval: false

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(data_id = CLASS),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(                        
    opts_hover(css = "fill: #202020;"),  
    opts_hover_inv(css = "opacity:0.2;") 
  )                                        
)                                        
```
:::

# Combining tooltip and hover effect
::: {.panel-tabset}

## The plot

```{r}
#| echo: false
p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(tooltip = CLASS, 
        data_id = CLASS),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(                        
    opts_hover(css = "fill: #202020;"),  
    opts_hover_inv(css = "opacity:0.2;") 
  )                                        
)                                        

```
## The code
```{r}
#| eval: false
p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(tooltip = CLASS, 
        data_id = CLASS),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(                        
    opts_hover(css = "fill: #202020;"),  
    opts_hover_inv(css = "opacity:0.2;") 
  )                                        
)                                        
```
:::


# Click effect with onclick
::: {.panel-tabset}

## The plot

```{r}
#| echo: false
exam_data$onclick <- sprintf("window.open(\"%s%s\")",
"https://www.moe.gov.sg/schoolfinder?journey=Primary%20school",
as.character(exam_data$ID))

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(onclick = onclick),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618)                                        

```

## The code
```{r}
#| eval: false
exam_data$onclick <- sprintf("window.open(\"%s%s\")",
"https://www.moe.gov.sg/schoolfinder?journey=Primary%20school",
as.character(exam_data$ID))

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(onclick = onclick),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618)                                        
```
:::


# Coordinated multiple views with ggiraph

::: {.panel-tabset}

## The plot

```{r}
#| echo: false

p1 <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(data_id = ID),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +  
  coord_cartesian(xlim=c(0,100)) + 
  scale_y_continuous(NULL,               
                     breaks = NULL)

p2 <- ggplot(data=exam_data, 
       aes(x = ENGLISH)) +
  geom_dotplot_interactive(              
    aes(data_id = ID),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") + 
  coord_cartesian(xlim=c(0,100)) + 
  scale_y_continuous(NULL,               
                     breaks = NULL)

girafe(code = print(p1 + p2), 
       width_svg = 6,
       height_svg = 3,
       options = list(
         opts_hover(css = "fill: #202020;"),
         opts_hover_inv(css = "opacity:0.2;")
         )
       ) 

```
## The code
```{r}
#| eval: false
p1 <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(data_id = ID),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +  
  coord_cartesian(xlim=c(0,100)) + 
  scale_y_continuous(NULL,               
                     breaks = NULL)

p2 <- ggplot(data=exam_data, 
       aes(x = ENGLISH)) +
  geom_dotplot_interactive(              
    aes(data_id = ID),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") + 
  coord_cartesian(xlim=c(0,100)) + 
  scale_y_continuous(NULL,               
                     breaks = NULL)

girafe(code = print(p1 + p2), 
       width_svg = 6,
       height_svg = 3,
       options = list(
         opts_hover(css = "fill: #202020;"),
         opts_hover_inv(css = "opacity:0.2;")
         )
       ) 

```
:::

# Interactive data visualisation - plotly method

::: {.panel-tabset}

## The plot

```{r}
#| echo: false
plot_ly(data = exam_data, 
             x = ~MATHS, 
             y = ~ENGLISH)

```
## The code
```{r}
#| eval: false
plot_ly(data = exam_data, 
             x = ~MATHS, 
             y = ~ENGLISH)

```
:::

## adding a visual element

::: {.panel-tabset}
## The plot

```{r}
#| echo: false
plot_ly(data = exam_data, 
        x = ~ENGLISH, 
        y = ~MATHS, 
        color = ~RACE)

```
## The code
```{r}
#| eval: false
plot_ly(data = exam_data, 
        x = ~ENGLISH, 
        y = ~MATHS, 
        color = ~RACE)

```
:::

## Creating an interactive scatterplot, ggplotly method

::: {.panel-tabset}
## The plot

```{r}
#| echo: false
p <- ggplot(data=exam_data, 
            aes(x = MATHS,
                y = ENGLISH)) +
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
ggplotly(p)

```
## The code
```{r}
#| eval: false
p <- ggplot(data=exam_data, 
            aes(x = MATHS,
                y = ENGLISH)) +
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
ggplotly(p)

```
:::

## Coordinated linked plot, using plotly
::: {.panel-tabset}
## The plot

```{r}
#| echo: false
d <- highlight_key(exam_data)
p1 <- ggplot(data=d, 
            aes(x = MATHS,
                y = ENGLISH)) +
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))

p2 <- ggplot(data=d, 
            aes(x = MATHS,
                y = SCIENCE)) +
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
subplot(ggplotly(p1),
        ggplotly(p2))

```

## The code
```{r}
#| eval: false
d <- highlight_key(exam_data)
p1 <- ggplot(data=d, 
            aes(x = MATHS,
                y = ENGLISH)) +
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))

p2 <- ggplot(data=d, 
            aes(x = MATHS,
                y = SCIENCE)) +
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
subplot(ggplotly(p1),
        ggplotly(p2))
```
:::

# Interactive data visualisation - crosstalk method

::: {.panel-tabset}
## The plot

```{r}
#| echo: false

DT::datatable(exam_data, class= "compact")

```

## The code
```{r}
#| eval: false
DT::datatable(exam_data, class= "compact")

```
:::

## Linked Brushing
::: {.panel-tabset}
## The plot

```{r}
#| echo: false

d <- highlight_key(exam_data) 
p <- ggplot(d, 
            aes(ENGLISH, 
                MATHS)) + 
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))

gg <- highlight(ggplotly(p),        
                "plotly_selected")  

crosstalk::bscols(gg,               
                  DT::datatable(d), 
                  widths = 5)        
```

## The code
```{r}
#| eval: false
d <- highlight_key(exam_data) 
p <- ggplot(d, 
            aes(ENGLISH, 
                MATHS)) + 
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))

gg <- highlight(ggplotly(p),        
                "plotly_selected")  

crosstalk::bscols(gg,               
                  DT::datatable(d), 
                  widths = 5)        
```
:::

# Alternatives

## Interactive boxplot
::: {.panel-tabset}
## The plot

```{r}
#| echo: false

plot_ly(
  data = exam_data,
  x = ~RACE,
  y = ~MATHS,
  color = ~RACE,
  type = "box"
)     
```

## The code
```{r}
#| eval: false

plot_ly(
  data = exam_data,
  x = ~RACE,
  y = ~MATHS,
  color = ~RACE,
  type = "box"
)       
```
:::

## Interactive scatterplot with tooltip
::: {.panel-tabset}
## The plot

```{r}
#| echo: false
plot_ly(
  data = exam_data,
  x = ~MATHS,
  y = ~ENGLISH,
  color = ~RACE,
  text = ~paste(
    "ID:", ID,
    "<br>Class:", CLASS,
    "<br>Maths:", MATHS,
    "<br>English:", ENGLISH
  ),
  hoverinfo = "text",
  type = "scatter",
  mode = "markers"
)   
```

## The code
```{r}
#| eval: false

plot_ly(
  data = exam_data,
  x = ~MATHS,
  y = ~ENGLISH,
  color = ~RACE,
  text = ~paste(
    "ID:", ID,
    "<br>Class:", CLASS,
    "<br>Maths:", MATHS,
    "<br>English:", ENGLISH
  ),
  hoverinfo = "text",
  type = "scatter",
  mode = "markers"
)      
```
:::

## Interactive violin plot
::: {.panel-tabset}
## The plot

```{r}
#| echo: false

plot_ly(
  data = exam_data,
  x = ~RACE,
  y = ~MATHS,
  color = ~RACE,
  type = "violin",
  box = list(visible = TRUE),
  meanline = list(visible = TRUE)
)    
```

## The code
```{r}
#| eval: false

plot_ly(
  data = exam_data,
  x = ~RACE,
  y = ~MATHS,
  color = ~RACE,
  type = "violin",
  box = list(visible = TRUE),
  meanline = list(visible = TRUE)
)       
```
:::

## Interactive histogram
::: {.panel-tabset}
## The plot

```{r}
#| echo: false
plot_ly(
  data = exam_data,
  x = ~MATHS,
  color = ~RACE,
  type = "histogram",
  opacity = 0.7
) %>%
  layout(
    barmode = "overlay",
    title = "Interactive Histogram of Maths Scores"
  )    
```

## The code
```{r}
#| eval: false

plot_ly(
  data = exam_data,
  x = ~MATHS,
  color = ~RACE,
  type = "histogram",
  opacity = 0.7
) %>%
  layout(
    barmode = "overlay",
    title = "Interactive Histogram of Maths Scores"
  )       
```
:::

## Interactive density heatmap
::: {.panel-tabset}
## The plot

```{r}
#| echo: false
plot_ly(
  data = exam_data,
  x = ~MATHS,
  y = ~ENGLISH,
  type = "histogram2d",
  colorscale = "YlOrRd"
) 
```

## The code
```{r}
#| eval: false
plot_ly(
  data = exam_data,
  x = ~MATHS,
  y = ~ENGLISH,
  type = "histogram2d",
  colorscale = "YlOrRd"
)       
```
:::

