---
title: "Hands-on Exercise 3"
author: "Persis Evangeline"
date: "April 30, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Getting started

```{r}
pacman::p_load(readxl, gifski, gapminder,
               plotly, gganimate, tidyverse)    
```

# Importing data

```{r}
col <- c("Country", "Continent")
globalPop <- read_xls("GlobalPopulation.xls",
                      sheet="Data") %>%
  mutate_at(col, as.factor) %>%
  mutate(Year = as.integer(Year))
```

# Animated data visualisations
Using gganimate, which includes animations, beyond what ggplot2 does.

::: {.panel-tabset}
## The plot

```{r}
#| echo: false
ggplot(globalPop, aes(x = Old, y = Young, 
                      size = Population, 
                      colour = Country)) +
  geom_point(alpha = 0.7, 
             show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  labs(title = 'Year: {frame_time}', 
       x = '% Aged', 
       y = '% Young') +
  transition_time(Year) +       
  ease_aes('linear')           
```

## The code
```{r}
#| eval: false
ggplot(globalPop, aes(x = Old, y = Young, 
                      size = Population, 
                      colour = Country)) +
  geom_point(alpha = 0.7, 
             show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  labs(title = 'Year: {frame_time}', 
       x = '% Aged', 
       y = '% Young') +
  transition_time(Year) +       
  ease_aes('linear')                 
```
:::

# Animated bubble plot

::: {.panel-tabset}
## The plot

```{r}
#| echo: false
bp <- globalPop %>%
  plot_ly(x = ~Old, 
          y = ~Young, 
          size = ~Population, 
          color = ~Continent,
          sizes = c(2, 100),
          frame = ~Year, 
          text = ~Country, 
          hoverinfo = "text",
          type = 'scatter',
          mode = 'markers'
          ) %>%
  layout(showlegend = FALSE)
bp     
```

## The code
```{r}
#| eval: false
bp <- globalPop %>%
  plot_ly(x = ~Old, 
          y = ~Young, 
          size = ~Population, 
          color = ~Continent,
          sizes = c(2, 100),
          frame = ~Year, 
          text = ~Country, 
          hoverinfo = "text",
          type = 'scatter',
          mode = 'markers'
          ) %>%
  layout(showlegend = FALSE)
bp                
```
:::

# Alternatives
## Animated line chart

::: {.panel-tabset}
## The plot

```{r}
#| echo: false
ggplot(globalPop, aes(x = Year, y = Population, colour = Country)) +
  geom_line(show.legend = FALSE) +
  labs(title = "Population Growth Over Time: {frame_along}") +
  transition_reveal(Year)   
```

## The code
```{r}
#| eval: false
ggplot(globalPop, aes(x = Year, y = Population, colour = Country)) +
  geom_line(show.legend = FALSE) +
  labs(title = "Population Growth Over Time: {frame_along}") +
  transition_reveal(Year)               
```
:::


## Animated bar chart

::: {.panel-tabset}
## The plot

```{r}
#| echo: false
globalPop_continent <- globalPop %>%
  group_by(Year, Continent) %>%
  summarise(Population = sum(Population), .groups = "drop")

ggplot(globalPop_continent,
       aes(x = Continent,
           y = Population,
           fill = Continent)) +
  geom_col(show.legend = FALSE) +
  labs(title = "Year: {frame_time}",
       x = "Continent",
       y = "Population") +
  transition_time(Year) +
  ease_aes("linear")
```

## The code
```{r}
#| eval: false
globalPop_continent <- globalPop %>%
  group_by(Year, Continent) %>%
  summarise(Population = sum(Population), .groups = "drop")

ggplot(globalPop_continent,
       aes(x = Continent,
           y = Population,
           fill = Continent)) +
  geom_col(show.legend = FALSE) +
  labs(title = "Year: {frame_time}",
       x = "Continent",
       y = "Population") +
  transition_time(Year) +
  ease_aes("linear")         
```
:::

## Animated area chart

::: {.panel-tabset}
## The plot

```{r}
#| echo: false
globalPop %>%
  group_by(Year, Continent) %>%
  summarise(Population = sum(Population), .groups = "drop") %>%
  ggplot(aes(x = Year, 
             y = Population, 
             fill = Continent,
             group = Continent)) +
  geom_area(stat = "identity", alpha = 0.8) +
  labs(title = "Global Population by Continent: {frame_along}") +
  transition_reveal(along = Year)
```

## The code
```{r}
#| eval: false
globalPop %>%
  group_by(Year, Continent) %>%
  summarise(Population = sum(Population), .groups = "drop") %>%
  ggplot(aes(x = Year, 
             y = Population, 
             fill = Continent,
             group = Continent)) +
  geom_area(stat = "identity", alpha = 0.8) +
  labs(title = "Global Population by Continent: {frame_along}") +
  transition_reveal(along = Year)             
```
:::

## Animated histogram

::: {.panel-tabset}
## The plot

```{r}
#| echo: false
ggplot(globalPop, aes(x = Old)) +
  geom_histogram(bins = 20, fill = "steelblue", color = "white") +
  labs(
    title = "Distribution of Ageing Population: {frame_time}",
    x = "% Aged",
    y = "Count"
  ) +
  transition_time(Year) +
  ease_aes("linear")
```

## The code
```{r}
#| eval: false
ggplot(globalPop, aes(x = Old)) +
  geom_histogram(bins = 20, fill = "steelblue", color = "white") +
  labs(
    title = "Distribution of Ageing Population: {frame_time}",
    x = "% Aged",
    y = "Count"
  ) +
  transition_time(Year) +
  ease_aes("linear")          
```
:::

## Animated scatter

::: {.panel-tabset}
## The plot

```{r}
#| echo: false
ggplot(globalPop,
       aes(x = Old,
           y = Young,
           colour = Country)) +
  geom_point(show.legend = FALSE) +
  facet_wrap(~Continent) +
  labs(
    title = "Age Structure by Continent: {frame_time}",
    x = "% Aged",
    y = "% Young"
  ) +
  transition_time(Year) +
  ease_aes("linear")
```

## The code
```{r}
#| eval: false
ggplot(globalPop,
       aes(x = Old,
           y = Young,
           colour = Country)) +
  geom_point(show.legend = FALSE) +
  facet_wrap(~Continent) +
  labs(
    title = "Age Structure by Continent: {frame_time}",
    x = "% Aged",
    y = "% Young"
  ) +
  transition_time(Year) +
  ease_aes("linear")         
```
:::




