---
title: "Hands-on Exercise 2"
author: "Persis Evangeline"
date: "April 23, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Getting started

There is a need to check if we have installed the relevant packages.

```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, hrbrthemes, 
               tidyverse)
```

Import and read the data set, which is the Exam_data.csv file

```{r}
exam_data <- read_csv("Exam_data.csv")
```

# Making different plots using ggplot.

First plot is to compare the English and Maths scores for Primary 3.

```{r}
ggplot(data=exam_data,
       aes(x=MATHS,
           y=ENGLISH)) +
  geom_point() + 
  geom_smooth(method=lm, size=0.5,
              vjust= -.5)+ 
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))+ 
  ggtitle("English scores versus Math scores for Primary 3")
```

Second plot is to compare the Science and Maths scores for Primary 3.
```{r}
ggplot(data=exam_data,
       aes(x=MATHS,
           y=SCIENCE)) +
  geom_point() + 
  geom_smooth(method=lm, size=0.5,
              vjust= -.5)+ 
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))+ 
  ggtitle("English scores versus Math scores for Primary 3")
```

I am going to plot the same type of graph to compare race and Maths scores. 
```{r}
ggplot(data=exam_data,
       aes(x=RACE,
           y=SCIENCE)) +
  geom_point() + 
  geom_smooth(method=lm, size=0.5,
              vjust= -.5)+ 
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))+ 
  ggtitle("English scores versus Math scores for Primary 3")
```
This plot appears strange as RACE is a categorical data type. Hence, the few values of RACE are compressed to the 
far left, making it difficult to interpret. 

A histogram or bar chart would be a better choice to plot a categorical and numerical into one plot. 
```{r}
ggplot(data=exam_data,
       aes(x=SCIENCE))+
  geom_histogram(bins=25,
                 boundary=100,
                 color="grey25",
                 fill="grey90")+
  ggtitle("Distribution of Science scores")+
  theme_economist()
```

This plot is much easier to interpret. It is important to take note of data types when deciding what plots,charts
and graphs to use.

I will make another plot to see the count of students by race.
However, the code below will give an error as geom_histogram() is for continuous numeric data, but RACE is a categorical/discrete variable.That is why ggplot says:stat_bin() requires a continuous x aesthetic.
```{r}
#ggplot(data=exam_data,
#       aes(x=RACE))+
#  geom_histogram(bins=25,
#                 boundary=100,
#                 color="grey25",
#                fill="grey90")+
#  ggtitle("Distribution of Science scores")+
#  theme_economist()
```

Instead, we can use geom_bar().
```{r}
ggplot(data = exam_data,
       aes(x = RACE)) +
  geom_bar(color = "grey25",
           fill = "grey90") +
  ggtitle("Distribution of Race") +
  theme_economist()
```

# Using ggrepel:
```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

# Working with ggtheme package:
```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum()
```

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum(axis_title_size = 18,
              base_size = 15,
              grid = "Y")
```

# Combining more than one graph into a single display:
Here we will assign the individual plots to a variable.

```{r}
p1 <- ggplot(data=exam_data, 
          aes(x=MATHS))+
      geom_histogram(bins=20,
                     boundary=100,
                     color="grey25",
                     fill="grey90")+
  coord_cartesian(xlim=c(0,100))+
  ggtitle("Distribution of Maths Score")
```


```{r}
p2 <- ggplot(data=exam_data,
             aes(x=ENGLISH))+
  geom_histogram(bins=20,
                 boundary=100,
                 color="blue2",
                 fill="white")+
  coord_cartesian(xlim=c(0,100))+
  ggtitle("Distribution of English scores")
```

```{r}
p3 <- ggplot(data=exam_data, 
             aes(x= MATHS, 
                 y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")

```
Now, combine the 2 plots into one.
```{r}
p1 + p2

```

Lets combine 3 into one. 
```{r}
p1+p2+p3

```
In the above plot, some portions of the title are cut out. So, we increase the width and height of the
area to plot.

```{r, fig.width=15, fig.height=10}
p1 + p2 + p3
```

There are other ways to display the charts into 1. 

“/” operator to stack two ggplot2 graphs,
“|” operator to place the plots beside each other,
“()” operator the define the sequence of the plotting.
```{r}
(p1 / p2) | p3
```

tagging/numbering:
```{r}
((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I')
```

Inserting the plot at specific locations/areas:
```{r}
p3 + inset_element(p2, 
                   left = 0.02, 
                   bottom = 0.7, 
                   right = 0.5, 
                   top = 1)

```

```{r}
patchwork <- (p1 / p2) | p3
patchwork & theme_economist()
```

# Alternatives
1. Label only selected points instead of every point
```{r}
nrow(subset(exam_data, MATHS > 80 | ENGLISH > 80))

ggplot(data = exam_data,
       aes(x = MATHS, y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm, linewidth = 0.5) +
  geom_label_repel(
    data = subset(exam_data, MATHS > 80 | ENGLISH > 80),
    aes(label = ID),
    fontface = "bold"
  ) +
  coord_cartesian(xlim = c(0, 100), ylim = c(0, 100)) +
  ggtitle("English scores versus Maths scores for P3")
```

2. Use geom_text_repel()
```{r}
ggplot(data = exam_data,
       aes(x = MATHS, y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm, linewidth = 0.5) +
  geom_text_repel(
    aes(label = ID),
    fontface = "bold",
    size = 3
  ) +
  coord_cartesian(xlim = c(0, 100), ylim = c(0, 100)) +
  ggtitle("English scores versus Maths scores for P3")
```

3. Map color to a grouping variable in the labelled scatterplot
```{r}
ggplot(data = exam_data,
       aes(x = MATHS, y = ENGLISH, colour = GENDER)) +
  geom_point() +
  geom_smooth(method = lm, linewidth = 0.5, se = FALSE) +
  geom_text_repel(
    data = subset(exam_data, MATHS > 85 | ENGLISH > 85),
    aes(label = ID),
    show.legend = FALSE
  ) +
  coord_cartesian(xlim = c(0, 100), ylim = c(0, 100)) +
  ggtitle("English V Maths scores by Gender")
```

4. Using a different theme instead of ggthemes. Use theme_wsj()
```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_histogram(
    bins = 20,
    boundary = 100,
    color = "grey25",
    fill = "grey90"
  ) +
  ggtitle("Distribution of Maths scores") +
  theme_fivethirtyeight()
```

5. Use theme()
```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_histogram(
    bins = 20,
    boundary = 100,
    color = "grey25",
    fill = "grey90"
  ) +
  ggtitle("Distribution of Maths scores") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 13),
    axis.text = element_text(size = 11)
  )
```

6. Combine 2 plots with wrap_plots()
```{r}
wrap_plots(p1, p2, ncol = 2)
```

7. Adding overall title and subtitle
```{r}
(p1 / p2) | p3 +
  plot_annotation(
    title = "Exam Performance",
    subtitle = "Distribution of Maths & English scores with relationship plot",
    tag_levels = "A"
  )
```