---
title: "Hands-on Exercise 2"
author: "Persis Evangeline"
date: "April 23, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

#Getting started

There is a need to check if we have installed the relevant packages.

```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, hrbrthemes, 
               tidyverse)
```

Import and read the data set, which is the Exam_data.csv file

```{r}
exam_data <- read_csv("Exam_data.csv")
```

Making different plots using ggplot.

First plot is to compare the English and Maths scores for Primary 3.

```{r}
ggplot(data=exam_data,
       aes(x=MATHS,
           y=ENGLISH)) +
  geom_point() + 
  geom_smooth(method=lm, size=0.5,
              vjust= -.5)+ 
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))+ 
  ggtitle("English scores versus Math scores for Primary 3")
```

Second plot is to compare the Science and Maths scores for Primary 3.
```{r}
ggplot(data=exam_data,
       aes(x=MATHS,
           y=SCIENCE)) +
  geom_point() + 
  geom_smooth(method=lm, size=0.5,
              vjust= -.5)+ 
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))+ 
  ggtitle("English scores versus Math scores for Primary 3")
```

I am going to plot the same type of graph to compare race and Maths scores. 
```{r}
ggplot(data=exam_data,
       aes(x=RACE,
           y=SCIENCE)) +
  geom_point() + 
  geom_smooth(method=lm, size=0.5,
              vjust= -.5)+ 
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))+ 
  ggtitle("English scores versus Math scores for Primary 3")
```
This plot appears strange as RACE is a categorical data type. Hence, the few values of RACE are compressed to the 
far left, making it difficult to interpret. 

A histogram or bar chart would be a better choice to plot a categorical and numerical into one plot. 
```{r}
ggplot(data=exam_data,
       aes(x=SCIENCE))+
  geom_histogram(bins=25,
                 boundary=100,
                 color="grey25",
                 fill="grey90")+
  ggtitle("Distribution of Science scores")+
  theme_economist()
```

This plot is much easier to interpret. It is important to take note of data types when deciding what plots,charts
and graphs to use.

I will make another plot to see the count of students by race.
However, the code below will give an error as geom_histogram() is for continuous numeric data, but RACE is a categorical/discrete variable.That is why ggplot says:stat_bin() requires a continuous x aesthetic.
```{r}
#ggplot(data=exam_data,
#       aes(x=RACE))+
#  geom_histogram(bins=25,
#                 boundary=100,
#                 color="grey25",
#                fill="grey90")+
#  ggtitle("Distribution of Science scores")+
#  theme_economist()
```

Instead, we can use geom_bar().
```{r}
ggplot(data = exam_data,
       aes(x = RACE)) +
  geom_bar(color = "grey25",
           fill = "grey90") +
  ggtitle("Distribution of Race") +
  theme_economist()
```

Combining more than one graph into a single display:
Here we will assign the individual plots to a variable.

```{r}
p1 <- ggplot(data=exam_data, 
          aes(x=MATHS))+
      geom_histogram(bins=20,
                     boundary=100,
                     color="grey25",
                     fill="grey90")+
  coord_cartesian(xlim=c(0,100))+
  ggtitle("Distribution of Maths Score")
```


```{r}
p2 <- ggplot(data=exam_data,
             aes(x=ENGLISH))+
  geom_histogram(bins=20,
                 boundary=100,
                 color="blue2",
                 fill="white")+
  coord_cartesian(xlim=c(0,100))+
  ggtitle("Distribution of English scores")
```

```{r}
p3 <- ggplot(data=exam_data, 
             aes(x= MATHS, 
                 y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")

```
Now, combine the plots into one.
```{r}
p1 + p2

```

Lets combine 3 into one. 
```{r}
p1+p2+p3

```
In the above plot, some portions of the title are cut out. So, we increase the width and height of the
area to plot.

```{r, fig.width=15, fig.height=10}
p1 + p2 + p3
```
