---
title: "Hands-on_Ex05_Heatmap_for_multivariate_data"
format: html
author: "Persis Evangeline"
date: "May 15, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

Heatmaps are data visualisation techniques that represent values using colours. Instead of displaying numbers directly, colours are used to show the magnitude or intensity of values within matrix or table.

Typically:
1. darker or stronger colours represent larger values
2. lighter or weaker colours represent smaller values

Heatmaps consist of rows, columns and cells. Heatmaps are useful for detecting patterns, identifying correlations, comparing multiple variables simultaneously or detecting clusters or groups. 
There are different types of heatmaps such as Correlation heatmaps, geographic heatmaps, activity heatmaps, matrix heatmaps.

# Installing and launching packages
```{r}
pacman::p_load(seriation, dendextend, heatmaply, tidyverse)

```

# Importing dataset

```{r}
wh <- read_csv("WHData-2018.csv")
```


# Preparing the data

```{r}
row.names(wh) <- wh$Country
```

# Transforming the dataframe into a matrix

```{r}
wh1 <- dplyr::select(wh, c(3, 7:12))
wh_matrix <- data.matrix(wh)
```


# Static heatmaps

::: panel-tabset
## The plot

```{r}
#| echo: false
wh_heatmap <- heatmap(wh_matrix,
                      Rowv=NA, Colv=NA)

```
## The code
```{r}
#| eval: false
wh_heatmap <- heatmap(wh_matrix,
                      Rowv=NA, Colv=NA)

```
:::

## Plot cluster heatmap

::: panel-tabset
## The plot

```{r}
#| echo: false
wh_heatmap <- heatmap(wh_matrix)

```
## The code
```{r}
#| eval: false
wh_heatmap <- heatmap(wh_matrix)

```
:::

## normalise the matrix column-wise

::: panel-tabset
## The plot

```{r}
#| echo: false
wh_heatmap <- heatmap(wh_matrix,
                      scale="column",
                      cexRow = 0.6, 
                      cexCol = 0.8,
                      margins = c(10, 4))

```
## The code
```{r}
#| eval: false
wh_heatmap <- heatmap(wh_matrix,
                      scale="column",
                      cexRow = 0.6, 
                      cexCol = 0.8,
                      margins = c(10, 4))

```
:::

# Interactive heatmap

::: panel-tabset
## The plot

```{r}
#| echo: false
heatmaply(mtcars)

```
## The code
```{r}
#| eval: false
heatmaply(mtcars)

```
:::

::: panel-tabset
## The plot

```{r}
#| echo: false
heatmaply(wh_matrix[, -c(1, 2, 4, 5)])
```
## The code
```{r}
#| eval: false
heatmaply(wh_matrix[, -c(1, 2, 4, 5)])
```
:::


# Data transformation
When analysing multivariate datasets, different variables often use different measurement scales and units. When the ranges of values differ greatly, variables with larger numerical values may dominate the analysis, especially in clustering, distance calculations, heatmaps, machine learning models. Transformation is important as it helps place variables with different ranges on a more comparable scale. There are three commonly used transformation methods, which are scale, normalise and percentile.  

## Scaling method

Scaling method is commonly used when variables are assumed to follow or approximately follow a normal distribution. The transformation is performed by subtracting the mean and then dividing by the standard deviation. 

::: panel-tabset
## The plot

```{r}
#| echo: false
heatmaply(wh_matrix[, -c(1, 2, 4, 5)])
```
## The code
```{r}
#| eval: false
heatmaply(wh_matrix[, -c(1, 2, 4, 5)])
```
:::

## Normalising method
Normalisation is a data transformation technique used to rescale values into a common range, typically 0 to 1. It is especially useful when variables come from different measurement scales, may not follow normal distribution or contain different ranges of values. Unlike scaling, normalisation does not standardise based on mean and standard deviation. Instead, it rescales values relative to the minimum and maximum values observed in the dataset. It preserves distribution shape. 

::: panel-tabset
## The plot

```{r}
#| echo: false
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]))
```
## The code
```{r}
#| eval: false
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]))
```
:::

## Percentising method

Percentising is a data transformation technique that converts values into their empirical percentiles. Instead of focusing on the raw numerical value, or the distance from the mean, percentising focuses on the relative position of a value within the dataset. Percentising is useful because variables become directly comparable, extreme scales are removed and interpretation becomes intuitive. Unlike scaling, no normal distribution assumption is needed. Unlike normalisation, it is less sensitive to extreme outliers.         

::: panel-tabset
## The plot

```{r}
#| echo: false
heatmaply(percentize(wh_matrix[, -c(1, 2, 4, 5)]))
```
## The code
```{r}
#| eval: false
heatmaply(percentize(wh_matrix[, -c(1, 2, 4, 5)]))
```
:::


# Clustering algorithm

In heatmaps, clustering algorithms are used to automatically group together similar variables, similar observations or both. With clustering, similar items are placed close together and create visually meaningful groupings. Main purpose is to identify similarity and structure in complex data.  

## Manual Approach

::: panel-tabset
## The plot

```{r}
#| echo: false
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "euclidean",
          hclust_method = "ward.D")
```
## The code
```{r}
#| eval: false
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "euclidean",
          hclust_method = "ward.D")
```
:::

## Statistical approach

::: panel-tabset
## The plot

```{r}
#| echo: false
wh_d <- dist(normalize(wh_matrix[, -c(1, 2, 4, 5)]), method = "euclidean")
dend_expend(wh_d)[[3]]

wh_clust <- hclust(wh_d, method = "average")
num_k <- find_k(wh_clust)
plot(num_k)
```
## The code
```{r}
#| eval: false
wh_d <- dist(normalize(wh_matrix[, -c(1, 2, 4, 5)]), method = "euclidean")
dend_expend(wh_d)[[3]]

wh_clust <- hclust(wh_d, method = "average")
num_k <- find_k(wh_clust)
plot(num_k)
```
:::

::: panel-tabset
## The plot

```{r}
#| echo: false
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "euclidean",
          hclust_method = "average",
          k_row = 3)
```
## The code
```{r}
#| eval: false
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "euclidean",
          hclust_method = "average",
          k_row = 3)
```
:::

# Seriation
Seriation means finding the "best" order to arrange rows and columns so similar items appear next to each other and patterns become visually clearer. Hierarchical clustering groups similar items together, but it does not fully determine the exact left-to-right or top-to-bottom order of the leaves in the dendogram. 
Seriation tries to reorder rows/columns while respecting the clustering structure, so nearby items are as similar as possible. The goal is to create smooth colour transitions, visibile blocks and cleaner patterns  in the heatmap.

Here, we use Optimal Leaf Ordering(OLO). OLO is a seriation algorithm. It starts with the hierarchical clustering tree, rotates branches left/right and searches for the ordering that minimises differences between adjacent items. 

::: panel-tabset
## The plot

```{r}
#| echo: false
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "OLO")
```
## The code
```{r}
#| eval: false
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "OLO")
```
:::

::: panel-tabset
## The plot

```{r}
#| echo: false
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "GW")
```
## The code
```{r}
#| eval: false
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "GW")
```
:::

::: panel-tabset
## The plot

```{r}
#| echo: false
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "mean")
```
## The code
```{r}
#| eval: false
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "mean")
```
:::

::: panel-tabset
## The plot

```{r}
#| echo: false
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "none")
```
## The code
```{r}
#| eval: false
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "none")
```
:::

# Working with colour palettes

::: panel-tabset
## The plot

```{r}
#| echo: false
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "none",
          colors = Blues)
```
## The code
```{r}
#| eval: false
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "none",
          colors = Blues)
```
:::


::: panel-tabset
## The plot

```{r}
#| echo: false
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          Colv=NA,
          seriate = "none",
          colors = Blues,
          k_row = 5,
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 5,
          main="World Happiness Score and Variables by Country, 2018 \nDataTransformation using Normalise Method",
          xlab = "World Happiness Indicators",
          ylab = "World Countries"
          )
```
## The code
```{r}
#| eval: false
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          Colv=NA,
          seriate = "none",
          colors = Blues,
          k_row = 5,
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 5,
          main="World Happiness Score and Variables by Country, 2018 \nDataTransformation using Normalise Method",
          xlab = "World Happiness Indicators",
          ylab = "World Countries"
          )
```
:::

# Alternatives

## Hierarchical clustering + OLO manually

::: panel-tabset
## The plot

```{r}
#| echo: false
library(seriation)

wh_norm <- normalize(wh_matrix[, -c(1, 2, 4, 5)])

row_dist <- dist(wh_norm)

row_order <- seriate(row_dist, method = "OLO")

heatmaply(
  wh_norm,
  Rowv = as.dendrogram(hclust(row_dist)),
  row_dend_left = TRUE
)
```
## The code
```{r}
#| eval: false
library(seriation)

wh_norm <- normalize(wh_matrix[, -c(1, 2, 4, 5)])

row_dist <- dist(wh_norm)

row_order <- seriate(row_dist, method = "OLO")

heatmaply(
  wh_norm,
  Rowv = as.dendrogram(hclust(row_dist)),
  row_dend_left = TRUE
)
```
:::

