---
title: "Hands-on_Ex05_visual_multivariate"
format: html
author: "Persis Evangeline"
date: "May 15, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Installing and launching R packages

```{r}

pacman::p_load(GGally, parallelPlot, tidyverse)

```

# Data preparation

```{r}
wh <- read_csv("WHData-2018.csv")

```

# Plotting static parallel coordinated plot

::: panel-tabset
## The plot

```{r}
#| echo: false
ggparcoord(data = wh, 
           columns = c(7:12))

```
## The code
```{r}
#| eval: false
ggparcoord(data = wh, 
           columns = c(7:12))

```
:::


# Plotting parallel coordinates with boxplot

::: panel-tabset
## The plot

```{r}
#| echo: false
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Parallel Coordinates Plot of World Happines Variables")

```
## The code
```{r}
#| eval: false
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Parallel Coordinates Plot of World Happines Variables")

```
:::

# Parallel coordinates with facet

::: panel-tabset
## The plot

```{r}
#| echo: false
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region)

```
## The code
```{r}
#| eval: false
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region)

```
:::

# Rotating x axis text label

::: panel-tabset
## The plot

```{r}
#| echo: false
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30))

```
## The code
```{r}
#| eval: false
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30))

```
:::

# Adjust the rotating x axix text label

::: panel-tabset
## The plot

```{r}
#| echo: false
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30, hjust=1))

```
## The code
```{r}
#| eval: false
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30, hjust=1))
```
:::


# Plotting interactive parallel coordinates plot:parallelPlot methods

## basic plot

::: panel-tabset
## The plot

```{r}
#| echo: false
wh <- wh %>%
  select("Happiness score", c(7:12))
parallelPlot(wh,
             width = 320,
             height = 250)

```
## The code
```{r}
#| eval: false
wh <- wh %>%
  select("Happiness score", c(7:12))
parallelPlot(wh,
             width = 320,
             height = 250)

```
:::

## rotate axis label

::: panel-tabset
## The plot

```{r}
#| echo: false
parallelPlot(wh,
             rotateTitle = TRUE)

```
## The code
```{r}
#| eval: false
parallelPlot(wh,
             rotateTitle = TRUE)

```
:::


## changing colour

::: panel-tabset
## The plot

```{r}
#| echo: false
parallelPlot(wh,
             continuousCS = "YlOrRd",
             rotateTitle = TRUE)
```
## The code
```{r}
#| eval: false
parallelPlot(wh,
             continuousCS = "YlOrRd",
             rotateTitle = TRUE)

```
:::

## parallel coordinates plot with histogram

::: panel-tabset
## The plot

```{r}
#| echo: false
histoVisibility <- rep(TRUE, ncol(wh))
parallelPlot(wh,
             rotateTitle = TRUE,
             histoVisibility = histoVisibility)
```
## The code
```{r}
#| eval: false
histoVisibility <- rep(TRUE, ncol(wh))
parallelPlot(wh,
             rotateTitle = TRUE,
             histoVisibility = histoVisibility)

```
:::




