---
title: "Hands-on_Ex05_Building_Ternary_PlotWR"
format: html
author: "Persis Evangeline"
date: "May 15, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---
A ternary plot is a special type of graph used to visualise compositional data made up of 3 parts that together sum to a constant.This constant is usually 1 or 100%. Some examples include population groups (young, economically active, elderly), soil composition(sand, clay, silt) and budget allocation( savings, expenses, investments). Each corner of the triangle represents 100% of 1 component and 0% of the other 2 components.   

# Getting Started

```{r}
pacman::p_load(plotly, ggtern, tidyverse)

```
# Importing Dataset

```{r}
pop_data <- read.csv("respopagsex2000to2018_tidy.csv")

```

# Preparing data

Utilising mutate() function  

```{r}
# Deriving the young, economy active and old measures

agpop_mutated <- pop_data %>% 
  mutate(`Year` = as.character(Year))%>%
  spread(AG, Population) %>%
  mutate(YOUNG = rowSums(.[4:8])) %>%
  mutate(ACTIVE = rowSums(.[9:16])) %>%
  mutate(OLD = rowSums(.[17:21])) %>%
  mutate(TOTAL = rowSums(.[22:24])) %>%
  filter(Year == 2018) %>%
  filter(TOTAL >0)
```

# Plotting Ternary diagram with R

::: panel-tabset
## The plot

```{r}
#| echo: false

#Building the static ternary plot
ggtern(data=agpop_mutated,aes(x=YOUNG,y=ACTIVE, z=OLD)) +
  geom_point()

```
## The code
```{r}
#| eval: false
#Building the static ternary plot
ggtern(data=agpop_mutated,aes(x=YOUNG,y=ACTIVE, z=OLD)) +
  geom_point()

```
:::


# Building the static ternary plot

::: panel-tabset
## The plot

```{r}
#| echo: false

ggtern(data=agpop_mutated, aes(x=YOUNG,y=ACTIVE, z=OLD)) +
  geom_point() +
  labs(title="Population structure, 2015") +
  theme_rgbw()

```
## The code
```{r}
#| eval: false
ggtern(data=agpop_mutated, aes(x=YOUNG,y=ACTIVE, z=OLD)) +
  geom_point() +
  labs(title="Population structure, 2015") +
  theme_rgbw()

```
:::

# Plotting an interactive ternary diagram

::: panel-tabset
## The plot

```{r}
#| echo: false

# reusable function for creating annotation object
label <- function(txt) {
  list(
    text = txt, 
    x = 0.1, y = 1,
    ax = 0, ay = 0,
    xref = "paper", yref = "paper", 
    align = "center",
    font = list(family = "serif", size = 15, color = "white"),
    bgcolor = "#b3b3b3", bordercolor = "black", borderwidth = 2
  )
}

# reusable function for axis formatting
axis <- function(txt) {
  list(
    title = txt, tickformat = ".0%", tickfont = list(size = 10)
  )
}

ternaryAxes <- list(
  aaxis = axis("Young"), 
  baxis = axis("Active"), 
  caxis = axis("Old")
)

# Initiating a plotly visualization 
plot_ly(
  agpop_mutated, 
  a = ~YOUNG, 
  b = ~ACTIVE, 
  c = ~OLD, 
  color = I("black"), 
  type = "scatterternary"
) %>%
  layout(
    annotations = label("Ternary Markers"), 
    ternary = ternaryAxes
  ) 

```
## The code
```{r}
#| eval: false
# reusable function for creating annotation object
label <- function(txt) {
  list(
    text = txt, 
    x = 0.1, y = 1,
    ax = 0, ay = 0,
    xref = "paper", yref = "paper", 
    align = "center",
    font = list(family = "serif", size = 15, color = "white"),
    bgcolor = "#b3b3b3", bordercolor = "black", borderwidth = 2
  )
}

# reusable function for axis formatting
axis <- function(txt) {
  list(
    title = txt, tickformat = ".0%", tickfont = list(size = 10)
  )
}

ternaryAxes <- list(
  aaxis = axis("Young"), 
  baxis = axis("Active"), 
  caxis = axis("Old")
)

# Initiating a plotly visualization 
plot_ly(
  agpop_mutated, 
  a = ~YOUNG, 
  b = ~ACTIVE, 
  c = ~OLD, 
  color = I("black"), 
  type = "scatterternary"
) %>%
  layout(
    annotations = label("Ternary Markers"), 
    ternary = ternaryAxes
  )

```
:::

# Alternatives

## Ternary plot with colour by planning area

::: panel-tabset
## The plot

```{r}
#| echo: false
#| fig-width: 10
#| fig-height: 8

library(ggtern)

ggtern(
  data = agpop_mutated,
  aes(
    x = YOUNG,
    y = ACTIVE,
    z = OLD,
    color = PA
  )
) +

  geom_point(size = 2, alpha = 0.8) +

  labs(
    title = "Population Structure by Planning Area (2018)",
    x = "Young",
    y = "Economically Active",
    z = "Old"
  ) +

  theme_bw() +

  theme(
    plot.title = element_text(
      hjust = 0.5,
      size = 16,
      face = "bold"
    ),

    legend.position = "bottom",

    legend.title = element_text(size = 10),
    legend.text = element_text(size = 7),

    tern.axis.title.T = element_text(size = 12),
    tern.axis.title.L = element_text(size = 12),
    tern.axis.title.R = element_text(size = 12),

    tern.axis.text.T = element_text(size = 8),
    tern.axis.text.L = element_text(size = 8),
    tern.axis.text.R = element_text(size = 8),

    plot.margin = margin(
      t = 20,
      r = 40,
      b = 20,
      l = 40
    )
  )
```
## The code
```{r}
#| eval: false
#| fig-width: 10
#| fig-height: 8

library(ggtern)

ggtern(
  data = agpop_mutated,
  aes(
    x = YOUNG,
    y = ACTIVE,
    z = OLD,
    color = PA
  )
) +

  geom_point(size = 2, alpha = 0.8) +

  labs(
    title = "Population Structure by Planning Area (2018)",
    x = "Young",
    y = "Economically Active",
    z = "Old"
  ) +

  theme_bw() +

  theme(
    plot.title = element_text(
      hjust = 0.5,
      size = 16,
      face = "bold"
    ),

    legend.position = "bottom",

    legend.title = element_text(size = 10),
    legend.text = element_text(size = 7),

    tern.axis.title.T = element_text(size = 12),
    tern.axis.title.L = element_text(size = 12),
    tern.axis.title.R = element_text(size = 12),

    tern.axis.text.T = element_text(size = 8),
    tern.axis.text.L = element_text(size = 8),
    tern.axis.text.R = element_text(size = 8),

    plot.margin = margin(
      t = 20,
      r = 40,
      b = 20,
      l = 40
    )
  )

```
:::

## Bubble Ternary Plot, point size by total population

::: panel-tabset
## The plot

```{r}
#| echo: false

ggtern(data = agpop_mutated,
       aes(x = YOUNG,
           y = ACTIVE,
           z = OLD,
           size = TOTAL)) +
  
  geom_point(alpha = 0.7, color = "steelblue") +
  
  labs(
    title = "Population Composition with Population Size",
    size = "Total Population"
  ) +
  
  theme_bw()

```
## The code
```{r}
#| eval: false
ggtern(data = agpop_mutated,
       aes(x = YOUNG,
           y = ACTIVE,
           z = OLD,
           size = TOTAL)) +
  
  geom_point(alpha = 0.7, color = "steelblue") +
  
  labs(
    title = "Population Composition with Population Size",
    size = "Total Population"
  ) +
  
  theme_bw()

```
:::

## Interactive Plotly Ternary Plot with Hover Labels

::: panel-tabset
## The plot

```{r}
#| echo: false

library(plotly)

plot_ly(
  data = agpop_mutated,

  a = ~YOUNG,
  b = ~ACTIVE,
  c = ~OLD,

  type = "scatterternary",
  mode = "markers",

  text = ~paste(
    "Planning Area:", PA,
    "<br>Young:", YOUNG,
    "<br>Active:", ACTIVE,
    "<br>Old:", OLD,
    "<br>Total:", TOTAL
  ),

  hoverinfo = "text",

  marker = list(
    size = 8,
    color = "darkred"
  )

) %>%

  layout(
    title = "Interactive Population Ternary Plot",

    margin = list(
      t = 100,
      l = 50,
      r = 50,
      b = 50
    ),

    ternary = list(
      aaxis = list(title = "Young"),
      baxis = list(title = "Active"),
      caxis = list(title = "Old")
    )
  )

```
## The code
```{r}
#| eval: false
library(plotly)

plot_ly(
  data = agpop_mutated,

  a = ~YOUNG,
  b = ~ACTIVE,
  c = ~OLD,

  type = "scatterternary",
  mode = "markers",

  text = ~paste(
    "Planning Area:", PA,
    "<br>Young:", YOUNG,
    "<br>Active:", ACTIVE,
    "<br>Old:", OLD,
    "<br>Total:", TOTAL
  ),

  hoverinfo = "text",

  marker = list(
    size = 8,
    color = "darkred"
  )

) %>%

  layout(
    title = "Interactive Population Ternary Plot",

    margin = list(
      t = 100,
      l = 50,
      r = 50,
      b = 50
    ),

    ternary = list(
      aaxis = list(title = "Young"),
      baxis = list(title = "Active"),
      caxis = list(title = "Old")
    )
  )

```
:::