---
title: "Hands-on_Ex05_Building_Correlation_Matrix"
format: html
author: "Persis Evangeline"
date: "May 15, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

A correlation matrix is a table that displays the correlation coefficients between multiple variables pairwise. It helps analysts quickly understand how variables move relative to one another. For example, if a dataset contains many different variables, a correlation matrix shows the relationship between every possible pair of variables. 

# Installing and launching R packages

```{r}
pacman::p_load(corrplot, ggstatsplot, tidyverse)
```

# Importing dataset

```{r}
wine <- read_csv("wine_quality.csv")
```

# Building correlation matrix, pairs method

The code below creates a scatterplot matrix, aka pairs plot. It automatically compares every variable against every other variable in the dataset. Each small box is a scatterplot between 2 variables.     

::: panel-tabset
## The plot

```{r}
#| echo: false
pairs(wine[,1:11])

```
## The code
```{r}
#| eval: false
pairs(wine[,1:11])

```
:::

::: panel-tabset
## The plot

```{r}
#| echo: false
pairs(wine[,2:12])
```
## The code
```{r}
#| eval: false
pairs(wine[,2:12])
```
:::

The difference in the 2 codes above is which columns are selected from the dataset.

# Removing redundant half

There is no need to show the entire correlation matrix as it is symmetric.So we can choose to show either the lower half or the upper half.

::: panel-tabset
## The plot

```{r}
#| echo: false
pairs(wine[,2:12], upper.panel = NULL)
```
## The code
```{r}
#| eval: false
pairs(wine[,2:12], upper.panel = NULL)
```
:::

::: panel-tabset
## The plot

```{r}
#| echo: false
pairs(wine[,2:12], lower.panel = NULL)
```
## The code
```{r}
#| eval: false
pairs(wine[,2:12], lower.panel = NULL)
```
:::


# Including with correlation coefficients

A correlation coefficient measures the direction and strength of the relationship between 2 variables. 
The range of the coefficient is: -1<=r<=1
+1, is a perfect positive relationship
-1, perfect negative relationship
0, no linear relationship

The significance of correlation coefficients are important. They help identify strong relationships, detect redundant variables, detect multicollinearity and understand real-world relationships.

Common guideline:

| Correlation | Interpretation |
| ----------- | -------------- |
| 0.00–0.19   | Very weak      |
| 0.20–0.39   | Weak           |
| 0.40–0.59   | Moderate       |
| 0.60–0.79   | Strong         |
| 0.80–1.00   | Very strong    |

::: panel-tabset
## The plot

```{r}
#| echo: false
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...) {
usr <- par("usr")
on.exit(par(usr))
par(usr = c(0, 1, 0, 1))
r <- abs(cor(x, y, use="complete.obs"))
txt <- format(c(r, 0.123456789), digits=digits)[1]
txt <- paste(prefix, txt, sep="")
if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
text(0.5, 0.5, txt, cex = cex.cor * (1 + r) / 2)
}

pairs(wine[,2:12], 
      upper.panel = panel.cor)
```
## The code
```{r}
#| eval: false
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...) {
usr <- par("usr")
on.exit(par(usr))
par(usr = c(0, 1, 0, 1))
r <- abs(cor(x, y, use="complete.obs"))
txt <- format(c(r, 0.123456789), digits=digits)[1]
txt <- paste(prefix, txt, sep="")
if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
text(0.5, 0.5, txt, cex = cex.cor * (1 + r) / 2)
}

pairs(wine[,2:12], 
      upper.panel = panel.cor)
```
:::

# Visualising Correlation Matrix, ggcormat()

::: panel-tabset
## The plot

```{r}
#| echo: false


ggstatsplot::ggcorrmat(
  data = wine, 
  cor.vars = 1:11,
  ggcorrplot.args = list(outline.color = "black", 
                         hc.order = TRUE,
                         tl.cex = 10),
  title    = "Correlogram for wine dataset",
  subtitle = "Four pairs are no significant at p < 0.05"
)

ggplot.component = list(
    theme(text=element_text(size=5),
      axis.text.x = element_text(size = 8),
      axis.text.y = element_text(size = 8)))
```
## The code
```{r}
#| eval: false

ggstatsplot::ggcorrmat(
  data = wine, 
  cor.vars = 1:11,
  ggcorrplot.args = list(outline.color = "black", 
                         hc.order = TRUE,
                         tl.cex = 10),
  title    = "Correlogram for wine dataset",
  subtitle = "Four pairs are no significant at p < 0.05"
)
ggplot.component = list(
    theme(text=element_text(size=5),
      axis.text.x = element_text(size = 8),
      axis.text.y = element_text(size = 8)))
```
:::

# Building multiple plots

::: panel-tabset
## The plot

```{r}
#| echo: false
grouped_ggcorrmat(
  data = wine,
  cor.vars = 1:11,
  grouping.var = type,
  type = "robust",
  p.adjust.method = "holm",
  plotgrid.args = list(ncol = 2),
  ggcorrplot.args = list(outline.color = "black", 
                         hc.order = TRUE,
                         tl.cex = 10),
  annotation.args = list(
    tag_levels = "a",
    title = "Correlogram for wine dataset",
    subtitle = "The measures are: alcohol, sulphates, fixed acidity, citric acid, chlorides, residual sugar, density, free sulfur dioxide and volatile acidity",
    caption = "Dataset: UCI Machine Learning Repository"
  )
)

```
## The code
```{r}
#| eval: false
grouped_ggcorrmat(
  data = wine,
  cor.vars = 1:11,
  grouping.var = type,
  type = "robust",
  p.adjust.method = "holm",
  plotgrid.args = list(ncol = 2),
  ggcorrplot.args = list(outline.color = "black", 
                         hc.order = TRUE,
                         tl.cex = 10),
  annotation.args = list(
    tag_levels = "a",
    title = "Correlogram for wine dataset",
    subtitle = "The measures are: alcohol, sulphates, fixed acidity, citric acid, chlorides, residual sugar, density, free sulfur dioxide and volatile acidity",
    caption = "Dataset: UCI Machine Learning Repository"
  )
)

```
:::

# Visualising correlation matrix using corrplot package

```{r}
wine.cor <- cor(wine[, 1:11])
```


::: panel-tabset
## The plot

```{r}
#| echo: false
corrplot(wine.cor)

```
## The code
```{r}
#| eval: false
corrplot(wine.cor)

```
:::

## Working with layout

::: panel-tabset
## The plot

```{r}
#| echo: false
corrplot(wine.cor, 
         method = "ellipse", 
         type="lower")

```
## The code
```{r}
#| eval: false
corrplot(wine.cor, 
         method = "ellipse", 
         type="lower")

```
:::

To further customise the layout of the corrgram:

::: panel-tabset
## The plot

```{r}
#| echo: false
corrplot(wine.cor, 
         method = "ellipse", 
         type="lower",
         diag = FALSE,
         tl.col = "black")
```
## The code
```{r}
#| eval: false
corrplot(wine.cor, 
         method = "ellipse", 
         type="lower",
         diag = FALSE,
         tl.col = "black")

```
:::


Other alternatives to customise the layout: 

1. Add correlation coefficients + rotate labels

::: panel-tabset
## The plot

```{r}
#| echo: false
corrplot(
  wine.cor,
  method = "ellipse",
  type = "lower",

  addCoef.col = "black",
  number.cex = 0.6,

  tl.col = "darkblue",
  tl.srt = 45,

  diag = FALSE
)
```
## The code
```{r}
#| eval: false
corrplot(
  wine.cor,
  method = "ellipse",
  type = "lower",

  addCoef.col = "black",
  number.cex = 0.6,

  tl.col = "darkblue",
  tl.srt = 45,

  diag = FALSE
)

```
:::

2. Cluster similar variables

::: panel-tabset
## The plot

```{r}
#| echo: false
corrplot(
  wine.cor,
  method = "color",

  order = "hclust",

  addrect = 3,

  tl.col = "black",
  tl.cex = 0.8,

  diag = FALSE
)
```
## The code
```{r}
#| eval: false
corrplot(
  wine.cor,
  method = "color",

  order = "hclust",

  addrect = 3,

  tl.col = "black",
  tl.cex = 0.8,

  diag = FALSE
)

```
:::

# Combining corrgram with the significance test
The purpose of the significance test in correlation analysis is to determine whether an observed correlation is likely to represent a real relationship in the population or if it occurred due to random chance in the sample data.

A correlation coefficient alone only tells us:
1. Strength of the relationship
2. Direction of the relationship

Significance testing is important as it helps us assess whether the observed relationship is credible. Based on the p value, we can interpret and decide if the observed relationship is credible.

::: panel-tabset
## The plot

```{r}
#| echo: false
wine.sig = cor.mtest(wine.cor, conf.level= .95)
corrplot(wine.cor,
         method = "number",
         type = "lower",
         diag = FALSE,
         tl.col = "black",
         tl.srt = 45,
         p.mat = wine.sig$p,
         sig.level = .05)
```
## The code
```{r}
#| eval: false
wine.sig = cor.mtest(wine.cor, conf.level= .95)
corrplot(wine.cor,
         method = "number",
         type = "lower",
         diag = FALSE,
         tl.col = "black",
         tl.srt = 45,
         p.mat = wine.sig$p,
         sig.level = .05)

```
:::

In the visualisation above, blue values indicate positive correlations. Red values indicate negative correlations. Larger absolute values indicate stronger relationships. 

# Reorder corrgram

::: panel-tabset
## The plot

```{r}
#| echo: false
corrplot.mixed(wine.cor, 
               lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               order="AOE",
               tl.col = "black")
```
## The code
```{r}
#| eval: false
corrplot.mixed(wine.cor, 
               lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               order="AOE",
               tl.col = "black")

```
:::

# Reordering a correlation matrix using hclust

::: panel-tabset
## The plot

```{r}
#| echo: false
corrplot(wine.cor, 
         method = "ellipse", 
         tl.pos = "lt",
         tl.col = "black",
         order="hclust",
         hclust.method = "ward.D",
         addrect = 3)
```
## The code
```{r}
#| eval: false
corrplot(wine.cor, 
         method = "ellipse", 
         tl.pos = "lt",
         tl.col = "black",
         order="hclust",
         hclust.method = "ward.D",
         addrect = 3)
```
:::

This visualisation shows a reordered correlation matrix using hierarchical clustering (hclust). 
Instead of displaying variables in their original oder, the variables are rearranged so that, 1. highly related variables are grouped together 2. making patterns and clusters easier to identify visually.
There are 3 clusters identified. Interpretation: blue ellipses means positive correlation, red ellipses means negative correlation, thinner ellipses means stronger correlation, circular shapes means weak/no correlation. Reordering the matrix improves interpretability by placing related variables closer together. Some inferences are are the strong negative relationship between alchohol and density or the strong positive relationship between free sulfur dioxide and total sulfur dioxide. 


