---
title: "Hands-on_Ex05_visual_multivariate"
format: html
author: "Persis Evangeline"
date: "May 15, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

A parallel coordinates plot is used to visualised many numerical variables at the same time. Instead of just 1 x-axis and 1 y-axis, it creates many vertical axes, one for each variable. In the world happiness dataset used in this example, the axes could be GDP, Corruption, Freedom etc and each country becomes one line that passes through all the axes.

Parallel coordinates plots help you compare multivariate patterns. Eg. which variables move together, whether clusters/groups exist, how differently countries behave from each other.

# Installing and launching R packages

```{r}

pacman::p_load(GGally, parallelPlot, tidyverse)

```

# Data preparation

```{r}
wh <- read_csv("WHData-2018.csv")

```

# Plotting static parallel coordinates plot

::: panel-tabset
## The plot

```{r}
#| echo: false
ggparcoord(data = wh, 
           columns = c(7:12))

```

## The code

```{r}
#| eval: false
ggparcoord(data = wh, 
           columns = c(7:12))

```
:::

# Plotting parallel coordinates with boxplot

The boxplot= TRUE means that a boxplot is drawn on each vertical axis. This helps summarise the distribution of each variable and makes the plot easier to interpret. The boxplots help by showing the centre of the data, spread/variability, skewness and possible outliers for each variable.

::: panel-tabset
## The plot

```{r}
#| echo: false
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Parallel Coordinates Plot of World Happines Variables")

```

## The code

```{r}
#| eval: false
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Parallel Coordinates Plot of World Happines Variables")

```
:::

# Parallel coordinates with facet

::: panel-tabset
## The plot

```{r}
#| echo: false
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region)

```

## The code

```{r}
#| eval: false
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region)

```
:::

# Rotating x axis text label

::: panel-tabset
## The plot

```{r}
#| echo: false
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30))

```

## The code

```{r}
#| eval: false
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30))

```
:::

# Adjust the rotating x axix text label

::: panel-tabset
## The plot

```{r}
#| echo: false
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30, hjust=1))

```

## The code

```{r}
#| eval: false
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30, hjust=1))
```
:::

# Plotting interactive parallel coordinates plot:parallelPlot methods

## basic plot

::: panel-tabset
## The plot

```{r}
#| echo: false
wh <- wh %>%
  select("Happiness score", c(7:12))
parallelPlot(wh,
             width = 320,
             height = 250)

```

## The code

```{r}
#| eval: false
wh <- wh %>%
  select("Happiness score", c(7:12))
parallelPlot(wh,
             width = 320,
             height = 250)

```
:::

## rotate axis label

::: panel-tabset
## The plot

```{r}
#| echo: false
parallelPlot(wh,
             rotateTitle = TRUE)

```

## The code

```{r}
#| eval: false
parallelPlot(wh,
             rotateTitle = TRUE)

```
:::

## changing colour

::: panel-tabset
## The plot

```{r}
#| echo: false
parallelPlot(wh,
             continuousCS = "YlOrRd",
             rotateTitle = TRUE)
```

## The code

```{r}
#| eval: false
parallelPlot(wh,
             continuousCS = "YlOrRd",
             rotateTitle = TRUE)

```
:::

## parallel coordinates plot with histogram

::: panel-tabset
## The plot

```{r}
#| echo: false
histoVisibility <- rep(TRUE, ncol(wh))
parallelPlot(wh,
             rotateTitle = TRUE,
             histoVisibility = histoVisibility)
```

## The code

```{r}
#| eval: false
histoVisibility <- rep(TRUE, ncol(wh))
parallelPlot(wh,
             rotateTitle = TRUE,
             histoVisibility = histoVisibility)

```
:::
