---
title: "Hands-on_Ex09C"
author: "Persis Evangeline"
date: "19 June, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

This section is about plotting analytical maps.

# Getting started

```{r}
pacman::p_load(tmap, tidyverse, sf)

```


# Importing data

```{r}
NGA_wp <- read_rds("data/NGA_wp.rds")

```


# Basic choropleth mapping

::: panel-tabset
## The plot

```{r}
#| echo: false

p1 <- tm_shape(NGA_wp) +
  tm_polygons(fill = "wp_functional",
             fill.scale = tm_scale_intervals(
               style = "equal",
               n = 10,
               values = "brewer.blues"),
             fill.legend = tm_legend(
               position = c("right", "bottom"))) +
  tm_borders(lwd = 0.1,
             fill_alpha = 1) +
  tm_title("Distribution of functional water point by LGAs")



p2 <- tm_shape(NGA_wp) + 
  tm_polygons(fill = "total_wp", 
              fill.scale = tm_scale_intervals(
                style = "equal",
                n = 10,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                position = c("right", "bottom"))) +
  tm_borders(lwd = 0.1, 
             fill_alpha = 1) + 
  tm_title("Distribution of total  water point by LGAs")

tmap_arrange(p2, p1, nrow = 1)
```

## The code

```{r}
#| eval: false

p1 <- tm_shape(NGA_wp) +
  tm_polygons(fill = "wp_functional",
             fill.scale = tm_scale_intervals(
               style = "equal",
               n = 10,
               values = "brewer.blues"),
             fill.legend = tm_legend(
               position = c("right", "bottom"))) +
  tm_borders(lwd = 0.1,
             fill_alpha = 1) +
  tm_title("Distribution of functional water point by LGAs")


p2 <- tm_shape(NGA_wp) + 
  tm_polygons(fill = "total_wp", 
              fill.scale = tm_scale_intervals(
                style = "equal",
                n = 10,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                position = c("right", "bottom"))) +
  tm_borders(lwd = 0.1, 
             fill_alpha = 1) + 
  tm_title("Distribution of total  water point by LGAs")

tmap_arrange(p2, p1, nrow = 1)

```
:::


# Choropleth map for rates

Rates should be mapped instead of counts because counts mainly reflect the spatial distribution of water points rather than the phenomenon being studied.

## Deriving Proportion of Functional Water Points and Non-Functional Water Points

We will tabulate the proportion of functional water points and the proportion of non-functional water points in each LGA. In the following code chunk, mutate() from dplyr package is used to derive two fields, namely pct_functional and pct_nonfunctional.

```{r}
NGA_wp <- NGA_wp %>%
  mutate(pct_functional = wp_functional/total_wp) %>%
  mutate(pct_nonfunctional = wp_nonfunctional/total_wp)

```


## Plotting map of rate

::: panel-tabset
## The plot

```{r}
#| echo: false

tm_shape(NGA_wp) +
  tm_polygons("pct_functional",
              fill.scale = tm_scale_intervals(
                style = "equal",
                n = 10,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                position = c("right", "bottom"))) + 
  tm_borders(lwd = 0.1,
             fill_alpha = 1) +
  tm_title("Rate map of functional water point by LGAs")
```

## The code

```{r}
#| eval: false

tm_shape(NGA_wp) +
  tm_polygons("pct_functional",
              fill.scale = tm_scale_intervals(
                style = "equal",
                n = 10,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                position = c("right", "bottom"))) + 
  tm_borders(lwd = 0.1,
             fill_alpha = 1) +
  tm_title("Rate map of functional water point by LGAs")

```
:::


# Extreme value maps

Extreme value maps are choropleth maps designed to highlight unusually low and high values, helping to identify spatial outliers.

## Percentile map

The percentile map is a special type of quantile map with six specific categories: 0-1%,1-10%, 10-50%,50-90%,90-99%, and 99-100%. The corresponding breakpoints can be derived by means of the base R quantile command, passing an explicit vector of cumulative probabilities as c(0,.01,.1,.5,.9,.99,1). Note that the begin and endpoint need to be included.

### Data preparation

#### Exclude records with NA by using the code chunk below.

```{r}
NGA_wp <- NGA_wp %>%
  drop_na()

```

#### Creating customised classification and extracting values

```{r}
percent <- c(0,.01,.1,.5,.9,.99,1)
var <- NGA_wp["pct_functional"] %>%
  st_set_geometry(NULL)
quantile(var[,1], percent)

```

#### Creating the get.var function

```{r}
get.var <- function(vname,df) {
  v <- df[vname] %>% 
    st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}

```

#### A percentile mapping function

```{r}
percentmap <- function(vnam, df, legtitle=NA, mtitle="Percentile Map"){
  percent <- c(0,.01,.1,.5,.9,.99,1)
  var <- get.var(vnam, df)
  bperc <- quantile(var, percent)
  tm_shape(df) +
  tm_polygons() +
  tm_shape(df) +
     tm_polygons(vnam,
             title=legtitle,
             breaks=bperc,
             palette="Blues",
          labels=c("< 1%", "1% - 10%", "10% - 50%", "50% - 90%", "90% - 99%", "> 99%"))  +
  tm_borders() +
  tm_layout(main.title = mtitle, 
            title.position = c("right","bottom"))
}

```

#### test the function

```{r}
percentmap("total_wp", NGA_wp)

```


## Box Map

A box map extends a quartile map by adding categories for lower and upper outliers, making extreme values easier to identify.

```{r}
ggplot(data = NGA_wp,
       aes(x = "",
           y = wp_nonfunctional)) +
  geom_boxplot()

```

### Creating the boxbreaks function

```{r}

boxbreaks <- function(v,mult=1.5) {
  qv <- unname(quantile(v))
  iqr <- qv[4] - qv[2]
  upfence <- qv[4] + mult * iqr
  lofence <- qv[2] - mult * iqr
  # initialize break points vector
  bb <- vector(mode="numeric",length=7)
  # logic for lower and upper fences
  if (lofence < qv[1]) {  # no lower outliers
    bb[1] <- lofence
    bb[2] <- floor(qv[1])
  } else {
    bb[2] <- lofence
    bb[1] <- qv[1]
  }
  if (upfence > qv[5]) { # no upper outliers
    bb[7] <- upfence
    bb[6] <- ceiling(qv[5])
  } else {
    bb[6] <- upfence
    bb[7] <- qv[5]
  }
  bb[3:5] <- qv[2:4]
  return(bb)
}

```

### Creating the get.var function

```{r}
get.var <- function(vname,df) {
  v <- df[vname] %>% st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}

```

### Test function

```{r}
var <- get.var("wp_nonfunctional", NGA_wp) 
boxbreaks(var)

```


### Boxmap Function

```{r}
boxmap <- function(vnam, df, 
                   legtitle=NA,
                   mtitle="Box Map",
                   mult=1.5){
  var <- get.var(vnam,df)
  bb <- boxbreaks(var)
  tm_shape(df) +
    tm_polygons() +
  tm_shape(df) +
     tm_fill(vnam,title=legtitle,
             breaks=bb,
             palette="Blues",
          labels = c("lower outlier", 
                     "< 25%", 
                     "25% - 50%", 
                     "50% - 75%",
                     "> 75%", 
                     "upper outlier"))  +
  tm_borders() +
  tm_layout(main.title = mtitle, 
            title.position = c("left",
                               "top"))
}

tmap_mode("plot")
boxmap("wp_nonfunctional", NGA_wp)

```

# Alternatives

## Compare functional and non-functional rate maps

::: panel-tabset
## The plot

```{r}
#| echo: false

NGA_wp <- NGA_wp %>%
  mutate(
    pct_functional = wp_functional / total_wp,
    pct_nonfunctional = wp_nonfunctional / total_wp
  )

functional_map <- tm_shape(NGA_wp) +
  tm_polygons(
    fill = "pct_functional",
    fill.scale = tm_scale_intervals(
      style = "quantile",
      n = 5,
      values = "brewer.blues"
    )
  ) +
  tm_borders(lwd = 0.1) +
  tm_title("Functional Water Point Rate")

nonfunctional_map <- tm_shape(NGA_wp) +
  tm_polygons(
    fill = "pct_nonfunctional",
    fill.scale = tm_scale_intervals(
      style = "quantile",
      n = 5,
      values = "brewer.reds"
    )
  ) +
  tm_borders(lwd = 0.1) +
  tm_title("Non-functional Water Point Rate")

tmap_arrange(functional_map, nonfunctional_map, nrow = 1)
```

## The code

```{r}
#| eval: false

NGA_wp <- NGA_wp %>%
  mutate(
    pct_functional = wp_functional / total_wp,
    pct_nonfunctional = wp_nonfunctional / total_wp
  )

functional_map <- tm_shape(NGA_wp) +
  tm_polygons(
    fill = "pct_functional",
    fill.scale = tm_scale_intervals(
      style = "quantile",
      n = 5,
      values = "brewer.blues"
    )
  ) +
  tm_borders(lwd = 0.1) +
  tm_title("Functional Water Point Rate")

nonfunctional_map <- tm_shape(NGA_wp) +
  tm_polygons(
    fill = "pct_nonfunctional",
    fill.scale = tm_scale_intervals(
      style = "quantile",
      n = 5,
      values = "brewer.reds"
    )
  ) +
  tm_borders(lwd = 0.1) +
  tm_title("Non-functional Water Point Rate")

tmap_arrange(functional_map, nonfunctional_map, nrow = 1)

```
:::

The functional water point rate is generally higher in the northern and northeastern LGAs, indicating better water point functionality in these regions.
Higher non-functional water point rates are concentrated in the southern and central LGAs, suggesting areas that may require greater maintenance and infrastructure intervention.


## Percentile map for non-functional water points

::: panel-tabset
## The plot

```{r}
#| echo: false

percent <- c(0, 0.01, 0.1, 0.5, 0.9, 0.99, 1)

var <- NGA_wp["pct_nonfunctional"] %>%
  st_set_geometry(NULL)

breaks_pct <- quantile(var[, 1], percent, na.rm = TRUE)

breaks_pct <- unique(breaks_pct)

tm_shape(NGA_wp) +
  tm_polygons(
    fill = "pct_nonfunctional",
    fill.scale = tm_scale_intervals(
      breaks = breaks_pct,
      values = "brewer.reds"
    ),
    fill.legend = tm_legend(
      title = "Percentile Class",
      position = c("right", "bottom")
    )
  ) +
  tm_borders(lwd = 0.1) +
  tm_title("Percentile Map of Non-functional Water Point Rate")
```

## The code

```{r}
percent <- c(0, 0.01, 0.1, 0.5, 0.9, 0.99, 1)

var <- NGA_wp["pct_nonfunctional"] %>%
  st_set_geometry(NULL)

breaks_pct <- quantile(var[, 1], percent, na.rm = TRUE)

breaks_pct <- unique(breaks_pct)

tm_shape(NGA_wp) +
  tm_polygons(
    fill = "pct_nonfunctional",
    fill.scale = tm_scale_intervals(
      breaks = breaks_pct,
      values = "brewer.reds"
    ),
    fill.legend = tm_legend(
      title = "Percentile Class",
      position = c("right", "bottom")
    )
  ) +
  tm_borders(lwd = 0.1) +
  tm_title("Percentile Map of Non-functional Water Point Rate")

```
:::

Most LGAs fall within the middle percentile classes, while only a small number of LGAs are classified in the highest percentile group with exceptionally high non-functional water point rates.
The darkest red areas represent spatial outliers with very high proportions of non-functional water points, highlighting locations that may require priority maintenance and resource allocation.























