---
title: "Hands-on_Ex09B"
author: "Persis Evangeline"
date: "19 June, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Getting started

Proportional symbol maps use symbol size to represent the magnitude of point-based data, such as counts or frequencies.
They can be graduated (classed) or proportional (unclassed), with proportional symbols scaled directly to the data values.
This exercise uses tmap to create a proportional symbol map of the number of wins at Singapore Pools outlets.

```{r}
pacman::p_load(sf, tmap, tidyverse)


```

# Geospatial data wrangling

```{r}
sgpools <- read_csv("data/SGPools_svy21.csv")

```

```{r}
list(sgpools) 


```

## Creating a sf data frame from an aspatial data frame

The code chunk below converts sgpools data frame into a simple feature data frame by using st_as_sf() of sf packages


```{r}
sgpools_sf <- st_as_sf(sgpools, 
                       coords = c("XCOORD", "YCOORD"),
                       crs= 3414)

```


Learning pointers:

1. The coords argument requires you to provide the column name of the x-coordinates first then followed by the column name of the y-coordinates.

2. The crs argument required you to provide the coordinates system in epsg format. EPSG: 3414 is Singapore SVY21 Projected Coordinate System. You can search for other country’s epsg code by refering to epsg.io.

```{r}

list(sgpools_sf)

```

The output shows that sgppols_sf is in point feature class. It’s epsg ID is 3414. The bbox provides information of the extend of the geospatial data.



# Drawing Proportional Symbol Map


```{r}
tmap_mode("view")

```


## interactive point symbol map

::: panel-tabset
## The plot

```{r}
#| echo: false

tm_shape(sgpools_sf) + 
  tm_bubbles(fill = "red",
           size = 1,
           col = "black",
           lwd = 1)
```

## The code

```{r}
#| eval: false

tm_shape(sgpools_sf) + 
  tm_bubbles(fill = "red",
           size = 1,
           col = "black",
           lwd = 1)

```
:::

## make it proportional


## interactive point symbol map

::: panel-tabset
## The plot

```{r}
#| echo: false

tm_shape(sgpools_sf) + 
  tm_bubbles(fill = "red",
             size = "Gp1Gp2 Winnings",
             col = "black",
             lwd = 1)
```

## The code

```{r}
#| eval: false

tm_shape(sgpools_sf) + 
  tm_bubbles(fill = "red",
             size = "Gp1Gp2 Winnings",
             col = "black",
             lwd = 1)

```
:::

## change colour

::: panel-tabset
## The plot

```{r}
#| echo: false

tm_shape(sgpools_sf) + 
  tm_bubbles(fill = "OUTLET TYPE", 
             size = "Gp1Gp2 Winnings",
             col = "black",
             lwd = 1)
```

## The code

```{r}
#| eval: false

tm_shape(sgpools_sf) + 
  tm_bubbles(fill = "OUTLET TYPE", 
             size = "Gp1Gp2 Winnings",
             col = "black",
             lwd = 1)

```
:::

## Faceted plot

::: panel-tabset
## The plot

```{r}
#| echo: false

tm_shape(sgpools_sf) + 
  tm_bubbles(fill = "OUTLET TYPE", 
             size = "Gp1Gp2 Winnings",
             col = "black",
             lwd = 1) + 
  tm_facets(by= "OUTLET TYPE",
            nrow = 1,
            sync = TRUE)
```

## The code

```{r}
#| eval: false
tm_shape(sgpools_sf) + 
  tm_bubbles(fill = "OUTLET TYPE", 
             size = "Gp1Gp2 Winnings",
             col = "black",
             lwd = 1) + 
  tm_facets(by= "OUTLET TYPE",
            nrow = 1,
            sync = TRUE)

```
:::



```{r}
tmap_mode("plot")

```

# Alternatives

## opebStreetMap Basemap

::: panel-tabset
## The plot

```{r}
#| echo: false

tmap_mode("view")

tm_shape(sgpools_sf) +
  tm_basemap("OpenStreetMap") +
  tm_bubbles(
    fill = "red",
    size = "Gp1Gp2 Winnings",
    col = "black",
    lwd = 1
  )
```

## The code

```{r}
#| eval: false
tmap_mode("view")

tm_shape(sgpools_sf) +
  tm_basemap("OpenStreetMap") +
  tm_bubbles(
    fill = "red",
    size = "Gp1Gp2 Winnings",
    col = "black",
    lwd = 1
  )

```
:::

## Use different colours by outlet type

::: panel-tabset
## The plot

```{r}
#| echo: false

tm_shape(sgpools_sf) +
  tm_bubbles(
    fill = "OUTLET TYPE",
    size = "Gp1Gp2 Winnings",
    palette = "Set2",
    col = "grey30",
    lwd = 0.8
  )
```

## The code

```{r}
#| eval: false
tm_shape(sgpools_sf) +
  tm_bubbles(
    fill = "OUTLET TYPE",
    size = "Gp1Gp2 Winnings",
    palette = "Set2",
    col = "grey30",
    lwd = 0.8
  )

```
:::

## Add transparency to overlapping bubbles

::: panel-tabset
## The plot

```{r}
#| echo: false

tm_shape(sgpools_sf) +
  tm_bubbles(
    fill = "blue",
    size = "Gp1Gp2 Winnings",
    fill_alpha = 0.5,
    col = "black",
    lwd = 0.5
  )
```

## The code

```{r}
#| eval: false
tm_shape(sgpools_sf) +
  tm_bubbles(
    fill = "blue",
    size = "Gp1Gp2 Winnings",
    fill_alpha = 0.5,
    col = "black",
    lwd = 0.5
  )

```
:::

## Static proportional symbol map

::: panel-tabset
## The plot

```{r}
#| echo: false

tmap_mode("plot")

tm_shape(sgpools_sf) +
  tm_bubbles(
    fill = "orange",
    size = "Gp1Gp2 Winnings",
    col = "black",
    lwd = 0.8
  ) +
  tm_layout(
    title = "Singapore Pools Outlets by Winnings",
    frame = FALSE
  )
```

## The code

```{r}
#| eval: false
tmap_mode("plot")

tm_shape(sgpools_sf) +
  tm_bubbles(
    fill = "orange",
    size = "Gp1Gp2 Winnings",
    col = "black",
    lwd = 0.8
  ) +
  tm_layout(
    title = "Singapore Pools Outlets by Winnings",
    frame = FALSE
  )

```
:::

## Bubble map with labels

::: panel-tabset
## The plot

```{r}
#| echo: false

tm_shape(sgpools_sf) +
  tm_bubbles(
    fill = "OUTLET TYPE",
    size = "Gp1Gp2 Winnings",
    col = "black",
    lwd = 0.8
  ) +
  tm_text(
    text = "NAME",
    size = 0.6
  )
```

## The code

```{r}
#| eval: false
tm_shape(sgpools_sf) +
  tm_bubbles(
    fill = "OUTLET TYPE",
    size = "Gp1Gp2 Winnings",
    col = "black",
    lwd = 0.8
  ) +
  tm_text(
    text = "NAME",
    size = 0.6
  )

```
:::

As the labels make the figure look too overcrowded, lets modify it. 

### Show labels only when zoomed in

::: panel-tabset
## The plot

```{r}
#| echo: false

tmap_mode("view")

tm_shape(sgpools_sf) +
  tm_bubbles(
    fill = "OUTLET TYPE",
    size = "Gp1Gp2 Winnings",
    col = "black",
    lwd = 0.8,
    popup.vars = c(
      "Outlet" = "NAME",
      "Winnings" = "Gp1Gp2 Winnings"
    )
  )
```

## The code

```{r}
#| eval: false
tmap_mode("view")

tm_shape(sgpools_sf) +
  tm_bubbles(
    fill = "OUTLET TYPE",
    size = "Gp1Gp2 Winnings",
    col = "black",
    lwd = 0.8,
    popup.vars = c(
      "Outlet" = "NAME",
      "Winnings" = "Gp1Gp2 Winnings"
    )
  )

```
:::

















