---
title: "Hands-on_Ex09"
author: "Persis Evangeline"
date: "19 June, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Getting started

A choropleth map uses colour shading to represent the spatial distribution of data values across geographic areas.
The main R package being used is tmap. Along with that, 4 other packages will be used which are readr, tidyr, dplyr and sf. 

```{r}
pacman::p_load(sf, tmap, tidyverse)
```


Two data set will be used to create the choropleth map. They are:

Master Plan 2014 Subzone Boundary (Web) (i.e. MP14_SUBZONE_WEB_PL) in ESRI shapefile format. It can be downloaded at data.gov.sg This is a geospatial data. It consists of the geographical boundary of Singapore at the planning subzone level. The data is based on URA Master Plan 2014.

Singapore Residents by Planning Area / Subzone, Age Group, Sex and Type of Dwelling, June 2011-2020 in csv format (i.e. respopagesextod2011to2020.csv). This is an aspatial data fie. It can be downloaded at Department of Statistics, Singapore Although it does not contain any coordinates values, but it’s PA and SZ fields can be used as unique identifiers to geocode to MP14_SUBZONE_WEB_PL shapefile.

# Importing data into R

```{r}
mpsz <- st_read(dsn = "data", 
                layer = "MP14_SUBZONE_WEB_PL")


```


## examine contents of data

::: panel-tabset
## The plot

```{r}
#| echo: false

mpsz
```

## The code

```{r}
#| eval: false

mpsz

```
:::

# Importing attribute data into R

```{r}

popdata <- read_csv("data/respopagesextod2011to2020.csv")
```

# Data preparation

Before a thematic map can be prepared, you are required to prepare a data table with year 2020 values. 
YOUNG: age group 0 to 4 until age groyup 20 to 24,
ECONOMY ACTIVE: age group 25-29 until age group 60-64,
AGED: age group 65 and above,
TOTAL: all age group, and
DEPENDENCY: the ratio between young and aged against economy active group


## Data wrangling

```{r}

popdata2020 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup() %>%
  pivot_wider(names_from=AG, 
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[12])) %>%
mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])+
rowSums(.[13:15]))%>%
mutate(`AGED`=rowSums(.[16:21])) %>%
mutate(`TOTAL`=rowSums(.[3:21])) %>%  
mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, 
       `ECONOMY ACTIVE`, `AGED`, 
       `TOTAL`, `DEPENDENCY`)

```


## Joining attribute data and geospatial data

```{r}

popdata2020 <- popdata2020 %>%
  mutate(across(c(PA, SZ), toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)
```


Next, left_join() of dplyr is used to join the geographical data and attribute table using planning subzone name e.g. SUBZONE_N and SZ as the common identifier.

```{r}

mpsz_pop2020 <- left_join(mpsz, popdata2020,
                          by = c("SUBZONE_N" = "SZ"))
```

```{r}
dir.create("data/rds", recursive = TRUE, showWarnings = FALSE)

write_rds(mpsz_pop2020, "data/rds/mpszpop2020.rds")

```


# Chloropleth mapping geospatial data using tmap

There are 2 approaches which can be used to prepare thematic map using tmap, they are:

1. Plotting a thematic map quickly by using qtm().

2. Plotting highly customisable thematic map by using tmap elements.


## Plotting a chloropleth map quickly using qtm()

::: panel-tabset
## The plot

```{r}
#| echo: false

tmap_mode("plot")
qtm(mpsz_pop2020, 
    fill = "DEPENDENCY")
```

## The code

```{r}
#| eval: false

tmap_mode("plot")
qtm(mpsz_pop2020, 
    fill = "DEPENDENCY")

```
:::

Learning points:

1. tmap_mode() with “plot” option is used to produce a static map. For interactive mode, “view” option should be used.

2. fill argument is used to map the attribute (i.e. DEPENDENCY)


## Creating a choropleth map by using tmap’s elements

::: panel-tabset
## The plot

```{r}
#| echo: false

tm_shape(mpsz_pop2020)+
  tm_polygons(fill = "DEPENDENCY", 
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "Dependency ratio")) +
  tm_title("Distribution of Dependency Ratio by planning subzone") +
  tm_layout(frame = TRUE) +
  tm_borders(fill_alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

## The code

```{r}
#| eval: false

tm_shape(mpsz_pop2020)+
  tm_polygons(fill = "DEPENDENCY", 
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "Dependency ratio")) +
  tm_title("Distribution of Dependency Ratio by planning subzone") +
  tm_layout(frame = TRUE) +
  tm_borders(fill_alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))

```
:::


### Drawing a base map

The basic building block of tmap is tm_shape() followed by one or more layer elemments such as tm_fill() and tm_polygons().

In the code chunk below, tm_shape() is used to define the input data (i.e mpsz_pop2020) and tm_polygons() is used to draw the planning subzone polygons

::: panel-tabset
## The plot

```{r}
#| echo: false

tm_shape(mpsz_pop2020) +
  tm_polygons()
```

## The code

```{r}
#| eval: false

tm_shape(mpsz_pop2020) +
  tm_polygons()

```
:::


### Drawing a choropleth map using tm_polygons()

To draw a choropleth map showing the geographical distribution of a selected variable by planning subzone, we just need to assign the target variable such as Dependency to tm_polygons().


::: panel-tabset
## The plot

```{r}
#| echo: false

tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY")
```

## The code

```{r}
#| eval: false

tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY")
```
:::

Learning pointers:

1. The default interval binning used to draw the choropleth map is called “pretty”.

2. The default colour scheme used is YlOrRd of ColorBrewer.

3. By default, Missing value will be shaded in grey.


###  Drawing a choropleth map using tm_fill() and tm_border()


Actually, tm_polygons() is a wraper of tm_fill() and tm_border(). tm_fill() shades the polygons by using the default colour scheme and tm_borders() adds the borders of the shapefile onto the choropleth map.

::: panel-tabset
## The plot

```{r}
#| echo: false

tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY")
```

## The code

```{r}
#| eval: false

tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY")
```
:::


Notice that the planning subzones are shared according to the respective dependecy values

To add the boundary of the planning subzones, tm_borders will be used as shown in the code chunk below.


::: panel-tabset
## The plot

```{r}
#| echo: false
tm_shape(mpsz_pop2020)+
  tm_polygons(fill = "DEPENDENCY") +
  tm_borders(lwd = 0.01,  
             fill_alpha = 0.1)
```

## The code

```{r}
#| eval: false
tm_shape(mpsz_pop2020)+
  tm_polygons(fill = "DEPENDENCY") +
  tm_borders(lwd = 0.01,  
             fill_alpha = 0.1)
```
:::


Notice that light-gray border lines have been added on the choropleth map.

The alpha argument is used to define transparency number between 0 (totally transparent) and 1 (not transparent). By default, the alpha value of the col is used (normally 1).


## Data classification methods of tmap

Choropleth maps use data classification to group many observations into a smaller number of value ranges or classes for easier interpretation.

tmap supports 10 classification methods (e.g., pretty, quantile, equal, jenks), which can be specified using the style argument in tm_fill() or tm_polygons().



::: panel-tabset
## The plot

```{r}
#| echo: false
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
      fill.scale = tm_scale_intervals(
        style = "jenks",
        n = 5)) +
  tm_borders(fill_alpha = 0.5)
```

## The code

```{r}
#| eval: false
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
      fill.scale = tm_scale_intervals(
        style = "jenks",
        n = 5)) +
  tm_borders(fill_alpha = 0.5)
```
:::

### Plotting choropleth maps with built-in classification methods


::: panel-tabset
## The plot

```{r}
#| echo: false
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
      fill.scale = tm_scale_intervals(
        style = "jenks",
        n = 5)) +
  tm_borders(fill_alpha = 0.5)
```

## The code

```{r}
#| eval: false
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
      fill.scale = tm_scale_intervals(
        style = "jenks",
        n = 5)) +
  tm_borders(fill_alpha = 0.5)
```
:::


::: panel-tabset
## The plot

```{r}
#| echo: false
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
      fill.scale = tm_scale_intervals(
        style = "equal",
        n = 5)) +
  tm_borders(fill_alpha = 0.5)
```

## The code

```{r}
#| eval: false
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
      fill.scale = tm_scale_intervals(
        style = "equal",
        n = 5)) +
  tm_borders(fill_alpha = 0.5)
```
:::

### DIY

Using what you had learned, prepare choropleth maps by using different classification methods supported by tmap and compare their differences.

::: panel-tabset
## The plot

```{r}
#| echo: false
methods <- c("pretty", "equal", "quantile", "jenks")

tmap_arrange(
  tm_shape(mpsz_pop2020) +
    tm_polygons("DEPENDENCY",
                fill.scale = tm_scale_intervals(style = "pretty", n = 5)) +
    tm_borders(fill_alpha = 0.5) +
    tm_title("Pretty"),

  tm_shape(mpsz_pop2020) +
    tm_polygons("DEPENDENCY",
                fill.scale = tm_scale_intervals(style = "equal", n = 5)) +
    tm_borders(fill_alpha = 0.5) +
    tm_title("Equal"),

  tm_shape(mpsz_pop2020) +
    tm_polygons("DEPENDENCY",
                fill.scale = tm_scale_intervals(style = "quantile", n = 5)) +
    tm_borders(fill_alpha = 0.5) +
    tm_title("Quantile"),

  tm_shape(mpsz_pop2020) +
    tm_polygons("DEPENDENCY",
                fill.scale = tm_scale_intervals(style = "jenks", n = 5)) +
    tm_borders(fill_alpha = 0.5) +
    tm_title("Jenks"),

  ncol = 2
)
```

## The code

```{r}
#| eval: false
methods <- c("pretty", "equal", "quantile", "jenks")

tmap_arrange(
  tm_shape(mpsz_pop2020) +
    tm_polygons("DEPENDENCY",
                fill.scale = tm_scale_intervals(style = "pretty", n = 5)) +
    tm_borders(fill_alpha = 0.5) +
    tm_title("Pretty"),

  tm_shape(mpsz_pop2020) +
    tm_polygons("DEPENDENCY",
                fill.scale = tm_scale_intervals(style = "equal", n = 5)) +
    tm_borders(fill_alpha = 0.5) +
    tm_title("Equal"),

  tm_shape(mpsz_pop2020) +
    tm_polygons("DEPENDENCY",
                fill.scale = tm_scale_intervals(style = "quantile", n = 5)) +
    tm_borders(fill_alpha = 0.5) +
    tm_title("Quantile"),

  tm_shape(mpsz_pop2020) +
    tm_polygons("DEPENDENCY",
                fill.scale = tm_scale_intervals(style = "jenks", n = 5)) +
    tm_borders(fill_alpha = 0.5) +
    tm_title("Jenks"),

  ncol = 2
)
```
:::



### DIY

Preparing choropleth maps by using similar classification method but with different numbers of classes (i.e. 2, 6, 10, 20). Compare the output maps, what observation can you draw?

::: panel-tabset
## The plot

```{r}
#| echo: false
tmap_arrange(
  tm_shape(mpsz_pop2020) +
    tm_polygons("DEPENDENCY",
                fill.scale = tm_scale_intervals(style = "jenks", n = 2)) +
    tm_borders(fill_alpha = 0.5) +
    tm_title("Jenks: 2 classes"),

  tm_shape(mpsz_pop2020) +
    tm_polygons("DEPENDENCY",
                fill.scale = tm_scale_intervals(style = "jenks", n = 6)) +
    tm_borders(fill_alpha = 0.5) +
    tm_title("Jenks: 6 classes"),

  tm_shape(mpsz_pop2020) +
    tm_polygons("DEPENDENCY",
                fill.scale = tm_scale_intervals(style = "jenks", n = 10)) +
    tm_borders(fill_alpha = 0.5) +
    tm_title("Jenks: 10 classes"),

  tm_shape(mpsz_pop2020) +
    tm_polygons("DEPENDENCY",
                fill.scale = tm_scale_intervals(style = "jenks", n = 20)) +
    tm_borders(fill_alpha = 0.5) +
    tm_title("Jenks: 20 classes"),

  ncol = 2
)
```

## The code

```{r}
#| eval: false
tmap_arrange(
  tm_shape(mpsz_pop2020) +
    tm_polygons("DEPENDENCY",
                fill.scale = tm_scale_intervals(style = "jenks", n = 2)) +
    tm_borders(fill_alpha = 0.5) +
    tm_title("Jenks: 2 classes"),

  tm_shape(mpsz_pop2020) +
    tm_polygons("DEPENDENCY",
                fill.scale = tm_scale_intervals(style = "jenks", n = 6)) +
    tm_borders(fill_alpha = 0.5) +
    tm_title("Jenks: 6 classes"),

  tm_shape(mpsz_pop2020) +
    tm_polygons("DEPENDENCY",
                fill.scale = tm_scale_intervals(style = "jenks", n = 10)) +
    tm_borders(fill_alpha = 0.5) +
    tm_title("Jenks: 10 classes"),

  tm_shape(mpsz_pop2020) +
    tm_polygons("DEPENDENCY",
                fill.scale = tm_scale_intervals(style = "jenks", n = 20)) +
    tm_borders(fill_alpha = 0.5) +
    tm_title("Jenks: 20 classes"),

  ncol = 2
)
```
:::




### Plotting choropleth map with custome break

tmap allows custom class intervals by specifying breakpoints manually using the breaks argument in tm_fill() or tm_polygons(). To create n classes, n+1 break values must be provided, including the minimum and maximum values. Before defining custom breaks, it is good practice to examine the descriptive statistics of the variable to understand its distribution.

```{r}
summary(mpsz_pop2020$DEPENDENCY)


```

With reference to the results above, we set break point at 0.60, 0.70, 0.80, and 0.90. In addition, we also need to include a minimum and maximum, which we set at 0 and 100. Our breaks vector is thus c(0, 0.60, 0.70, 0.80, 0.90, 1.00)


::: panel-tabset
## The plot

```{r}
#| echo: false
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
          breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00)) +
  tm_borders(fill_alpha = 0.5)
```

## The code

```{r}
#| eval: false
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
          breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00)) +
  tm_borders(fill_alpha = 0.5)


```
:::


## Colour Scheme

tmap supports colour ramps either defined by the user or a set of predefined colour ramps from the RColorBrewer package.

### Using colourBrewer palette

::: panel-tabset
## The plot

```{r}
#| echo: false
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
      fill.scale = tm_scale_intervals(
        style = "quantile",
        n = 5,
        values = "brewer.greens")) +
  tm_borders(fill_alpha = 0.5)
```

## The code

```{r}
#| eval: false
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
      fill.scale = tm_scale_intervals(
        style = "quantile",
        n = 5,
        values = "brewer.greens")) +
  tm_borders(fill_alpha = 0.5)


```
:::

To reverse the colour shaing, add a "-" prefix. 

::: panel-tabset
## The plot

```{r}
#| echo: false
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
      fill.scale = tm_scale_intervals(
        style = "quantile",
        n = 5,
        values = "-brewer.greens")) +
  tm_borders(fill_alpha = 0.5)
```

## The code

```{r}
#| eval: false
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
      fill.scale = tm_scale_intervals(
        style = "quantile",
        n = 5,
        values = "-brewer.greens")) +
  tm_borders(fill_alpha = 0.5)


```
:::

## Map layouts

Map layout refers to the arrangement of map elements such as the map itself, title, scale bar, compass, margins, and aspect ratio into a cohesive visualization. These elements, together with colour schemes and classification methods, influence the overall appearance and readability of the map.


### Map legend

In tmap, several tm_legend() options are provided to change the placement, format and appearance of the legend.

::: panel-tabset
## The plot

```{r}
#| echo: false
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
      fill.scale = tm_scale_intervals(
        style = "jenks",
        n = 5,
        values = "brewer.greens"),
      fill.legend = tm_legend(
        title = "Dependency ratio")) +
  tm_borders(fill_alpha = 0.5) +
  tm_title("Distribution of Dependency Ratio by planning subzone \n(Jenks classification)")
```

## The code

```{r}
#| eval: false
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
      fill.scale = tm_scale_intervals(
        style = "jenks",
        n = 5,
        values = "brewer.greens"),
      fill.legend = tm_legend(
        title = "Dependency ratio")) +
  tm_borders(fill_alpha = 0.5) +
  tm_title("Distribution of Dependency Ratio by planning subzone \n(Jenks classification)")
```
:::


### Map style

tmap allows a wide variety of layout settings to be changed. They can be called by using tmap_style().

::: panel-tabset
## The plot

```{r}
#| echo: false
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("classic")
```

## The code

```{r}
#| eval: false
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("classic")
```
:::

### Cartographic furniture

::: panel-tabset
## The plot

```{r}
#| echo: false
tm_shape(mpsz_pop2020)+
  tm_polygons(fill = "DEPENDENCY", 
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "Dependency ratio")) +
  tm_title("Distribution of Dependency Ratio by planning subzone") +
  tm_layout(frame = TRUE) +
  tm_borders(fill_alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

## The code

```{r}
#| eval: false
tm_shape(mpsz_pop2020)+
  tm_polygons(fill = "DEPENDENCY", 
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "Dependency ratio")) +
  tm_title("Distribution of Dependency Ratio by planning subzone") +
  tm_layout(frame = TRUE) +
  tm_borders(fill_alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```
:::

## Drawing Small Multiple Choropleth Maps

Small multiple (facet) maps display multiple maps side by side or stacked to compare spatial patterns across another variable, such as time. In tmap, they can be created using multiple aesthetic values, tm_facets(), or tmap_arrange().


### By assigning multiple values to at least one of the aesthetic arguments


::: panel-tabset
## The plot

```{r}
#| echo: false
tm_shape(mpsz_pop2020)+
  tm_fill(c("YOUNG", "AGED"),
          style = "equal", 
          palette = "Blues") +
  tm_layout(legend.position = c("right", "bottom")) +
  tm_borders(alpha = 0.5) +
  tmap_style("white")
```

## The code

```{r}
#| eval: false
tm_shape(mpsz_pop2020)+
  tm_fill(c("YOUNG", "AGED"),
          style = "equal", 
          palette = "Blues") +
  tm_layout(legend.position = c("right", "bottom")) +
  tm_borders(alpha = 0.5) +
  tmap_style("white")
```
:::


In this example, small multiple choropleth maps are created by assigning multiple values to at least one of the aesthetic arguments.


::: panel-tabset
## The plot

```{r}
#| echo: false
tm_shape(mpsz_pop2020)+ 
  tm_polygons(c("DEPENDENCY","AGED"),
          style = c("equal", "quantile"), 
          palette = list("Blues","Greens")) +
  tm_layout(legend.position = c("right", "bottom"))
```

## The code

```{r}
#| eval: false
tm_shape(mpsz_pop2020)+ 
  tm_polygons(c("DEPENDENCY","AGED"),
          style = c("equal", "quantile"), 
          palette = list("Blues","Greens")) +
  tm_layout(legend.position = c("right", "bottom"))
```
:::


### By defining a group-by variable in tm_facets()

::: panel-tabset
## The plot

```{r}
#| echo: false
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          thres.poly = 0) + 
  tm_facets(by="REGION_N", 
            free.coords=TRUE) +
  tm_layout(legend.show = FALSE,
            title.position = c("center", "center"), 
            title.size = 20) +
  tm_borders(alpha = 0.5)
```

## The code

```{r}
#| eval: false
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          thres.poly = 0) + 
  tm_facets(by="REGION_N", 
            free.coords=TRUE) +
  tm_layout(legend.show = FALSE,
            title.position = c("center", "center"), 
            title.size = 20) +
  tm_borders(alpha = 0.5)
```
:::


### By creating multiple stand-alone maps with tmap_arrange()

::: panel-tabset
## The plot

```{r}
#| echo: false
youngmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("YOUNG", 
              style = "quantile", 
              palette = "Blues")

agedmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Blues")

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)
```

## The code

```{r}
#| eval: false
youngmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("YOUNG", 
              style = "quantile", 
              palette = "Blues")

agedmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Blues")

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)
```
:::

## Mapping Spatial Object Meeting a Selection Criterion

Instead of creating small multiple choropleth map, you can also use selection funtion to map spatial objects meeting the selection criterion.


::: panel-tabset
## The plot

```{r}
#| echo: false
tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N=="CENTRAL REGION", ])+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(legend.outside = TRUE,
            legend.height = 0.45, 
            legend.width = 5.0,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```

## The code

```{r}
#| eval: false
tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N=="CENTRAL REGION", ])+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(legend.outside = TRUE,
            legend.height = 0.45, 
            legend.width = 5.0,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```
:::


# Alternatives


## Chloropleth map

::: panel-tabset
## The plot

```{r}
#| echo: false
tm_shape(mpsz_pop2020) +
  tm_polygons(
    fill = "DEPENDENCY",
    fill.scale = tm_scale_intervals(
      style = "quantile",
      n = 5,
      values = "brewer.blues"
    ),
    fill.legend = tm_legend(title = "Dependency Ratio")
  ) +
  tm_borders(fill_alpha = 0.4) +
  tm_title("Dependency Ratio by Planning Subzone")
```

## The code

```{r}
#| eval: false
tm_shape(mpsz_pop2020) +
  tm_polygons(
    fill = "DEPENDENCY",
    fill.scale = tm_scale_intervals(
      style = "quantile",
      n = 5,
      values = "brewer.blues"
    ),
    fill.legend = tm_legend(title = "Dependency Ratio")
  ) +
  tm_borders(fill_alpha = 0.4) +
  tm_title("Dependency Ratio by Planning Subzone")
```
:::

The dependency ratio is highest in several central and southern planning subzones, indicating a larger proportion of dependents (young and elderly) relative to the working-age population in these areas.


## Jenks classification map

::: panel-tabset
## The plot

```{r}
#| echo: false
tm_shape(mpsz_pop2020) +
  tm_polygons(
    fill = "DEPENDENCY",
    fill.scale = tm_scale_intervals(
      style = "jenks",
      n = 5,
      values = "brewer.greens"
    ),
    fill.legend = tm_legend(title = "Dependency Ratio")
  ) +
  tm_borders(lwd = 0.2, fill_alpha = 0.5) +
  tm_title("Dependency Ratio using Jenks Classification")
```

## The code

```{r}
#| eval: false
tm_shape(mpsz_pop2020) +
  tm_polygons(
    fill = "DEPENDENCY",
    fill.scale = tm_scale_intervals(
      style = "jenks",
      n = 5,
      values = "brewer.greens"
    ),
    fill.legend = tm_legend(title = "Dependency Ratio")
  ) +
  tm_borders(lwd = 0.2, fill_alpha = 0.5) +
  tm_title("Dependency Ratio using Jenks Classification")
```
:::

The Jenks classification map reveals that most planning subzones have moderate dependency ratios, while only a few distinct subzones exhibit exceptionally high dependency burdens compared to the rest of Singapore.


## Custom chloropleth map

::: panel-tabset
## The plot

```{r}
#| echo: false
tm_shape(mpsz_pop2020) +
  tm_polygons(
    fill = "DEPENDENCY",
    fill.scale = tm_scale_intervals(
      breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00),
      values = "brewer.oranges"
    ),
    fill.legend = tm_legend(title = "Dependency Ratio")
  ) +
  tm_borders(fill_alpha = 0.4) +
  tm_title("Dependency Ratio with Custom Breaks")
```

## The code

```{r}
#| eval: false
tm_shape(mpsz_pop2020) +
  tm_polygons(
    fill = "DEPENDENCY",
    fill.scale = tm_scale_intervals(
      breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00),
      values = "brewer.oranges"
    ),
    fill.legend = tm_legend(title = "Dependency Ratio")
  ) +
  tm_borders(fill_alpha = 0.4) +
  tm_title("Dependency Ratio with Custom Breaks")
```
:::

The custom-breaks map shows that many planning subzones fall within the 0.70–0.89 dependency ratio range, with clusters of very high dependency ratios (0.90–1.00) concentrated in central and southern Singapore.


## Comparing population groups

::: panel-tabset
## The plot

```{r}
#| echo: false
tm_shape(mpsz_pop2020) +
  tm_polygons(
    fill = c("YOUNG", "ECONOMY ACTIVE", "AGED"),
    fill.scale = tm_scale_intervals(
      style = "quantile",
      n = 5,
      values = "brewer.blues"
    )
  ) +
  tm_borders(fill_alpha = 0.4) +
  tm_title("Population Groups by Planning Subzone")
```

## The code

```{r}
#| eval: false
tm_shape(mpsz_pop2020) +
  tm_polygons(
    fill = c("YOUNG", "ECONOMY ACTIVE", "AGED"),
    fill.scale = tm_scale_intervals(
      style = "quantile",
      n = 5,
      values = "brewer.blues"
    )
  ) +
  tm_borders(fill_alpha = 0.4) +
  tm_title("Population Groups by Planning Subzone")
```
:::

The maps show that young, economically active, and aged populations are all concentrated in major residential planning subzones, with the economically active group consistently forming the largest population segment across Singapore.


## Regional facet chloropleth map

::: panel-tabset
## The plot

```{r}
#| echo: false
tm_shape(mpsz_pop2020) +
  tm_polygons(
    fill = "DEPENDENCY",
    fill.scale = tm_scale_intervals(
      style = "quantile",
      n = 5,
      values = "brewer.purples"
    ),
    fill.legend = tm_legend(title = "Dependency Ratio")
  ) +
  tm_facets(by = "REGION_N", free.coords = TRUE) +
  tm_borders(fill_alpha = 0.4) +
  tm_layout(legend.show = FALSE) +
  tm_title("Dependency Ratio by Region")
```

## The code

```{r}
#| eval: false
tm_shape(mpsz_pop2020) +
  tm_polygons(
    fill = "DEPENDENCY",
    fill.scale = tm_scale_intervals(
      style = "quantile",
      n = 5,
      values = "brewer.purples"
    ),
    fill.legend = tm_legend(title = "Dependency Ratio")
  ) +
  tm_facets(by = "REGION_N", free.coords = TRUE) +
  tm_borders(fill_alpha = 0.4) +
  tm_layout(legend.show = FALSE) +
  tm_title("Dependency Ratio by Region")
```
:::

The Central Region exhibits the greatest variation in dependency ratios across subzones, while the other regions generally show more moderate and spatially clustered dependency patterns.





