---
title: "Take Home Exercise 2"
format: html
author: "Persis Evangeline"
date: "June 2, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Overview

## Background
We investigate an anomalous SaidIT post made by John Windward on 17 May 2046 at 4:21 am. The post appears to have been published in an unrelated forum, suggesting malfunctioning behaviour. The objective is to reconstruct the sequence of events leading to the post, determine the origin and meaning of its contents, identify if similar incidents have occurred before and recommend an intervention to reduce the likelihood of its recurrence. 

## Questions to address

1. How was the anomalous SaidIT post created?

2. What is the origin and meaning of the post contents?

3. Have similar incidents occurred before?

4. What intervention point would most effectively prevent future occurrences?


## Methodology

This analysis will include the following below:


Data preparation & wrangling<br>
↓<br>
Event Log Analysis<br>
↓<br>
Network Construction<br>
↓<br>
Baseline System Behaviour<br>
↓<br>
Anomalous Chain Reconstruction<br>
↓<br>
Content Origin Analysis<br>
↓<br>
Historical Comparison<br>
↓<br>
Intervention Analysis

# Setup

```{r}
pacman::p_load(
  tidyverse,
  jsonlite,
  lubridate,

  igraph,
  tidygraph,
  ggraph,

  DT,
  patchwork,
  ggrepel,

  stringdist,
  tidytext,

  knitr,
  kableExtra
)

```


## Loading data

### Event log

```{r}
mc2_raw <- fromJSON(
  "MC2 data/VAST_Challenge_2026_MC2/MC2_data.json",
  simplifyVector = FALSE
)

```

### Organisation chart

```{r}
org_raw <- fromJSON(
  "MC2 data/VAST_Challenge_2026_MC2/org_chart.json"
)
```


# Data wrangling

## Preparing Event data

```{r}
events <- tibble(
  event = mc2_raw$events
) |>
  unnest_wider(event)

glimpse(events)

events <- events |>
  mutate(
    datetime = as.POSIXct(
      when,
      origin = "1970-01-01",
      tz = "UTC"
    ),
    date = as.Date(datetime),
    hour = hour(datetime),
    n_parties = map_int(parties, length)
  ) |>
  arrange(datetime)

events |>
  summarise(
    total_events = n(),
    earliest_event = min(datetime),
    latest_event = max(datetime)
  )

events |>
  count(short_name, sort = TRUE)


events |> count(n_parties)

```


## Preparing actor information 

```{r}
event_parties <- events |>
  select(id,
         datetime,
         short_name,
         parties) |>
  unnest_longer(parties,
                values_to = "party")

event_parties <- event_parties |>
  mutate(
    actor_type = case_when(
      str_detect(party, "^person:") ~ "Person",
      str_detect(party, "^Agent/") ~ "Agent",
      str_detect(party, "^world:") ~ "System",
      str_detect(party, "^team:") ~ "Team",
      str_detect(party, "^company:") ~ "Company",
      TRUE ~ "Other"
    )
  )

event_parties |>
  count(actor_type,
        sort = TRUE)

event_parties |>
  count(party,
        sort = TRUE) |>
  slice_head(n = 20)


```


## Preparing organisation data

```{r}
org_nodes <- as_tibble(org_raw$nodes)

glimpse(org_nodes)

org_nodes |>
  count(type)

org_edges <- as_tibble(org_raw$edges)

glimpse(org_edges)

org_edges |>
  count(relation)



```

## Viewing the 4 tables made earlier


::: panel-tabset
## The plot

```{r}
#| echo: false

events |>
  slice_head(n = 5) |>
  kable(caption = "First 5 records of events")

event_parties |>
  slice_head(n = 5) |>
  kable(caption = "First 5 records of event_parties")

org_nodes |>
  slice_head(n = 5) |>
  kable(caption = "First 5 records of org_nodes")

org_edges |>
  slice_head(n = 5) |>
  kable(caption = "First 5 records of org_edges")
```

## The code

```{r}
#| eval: false
events |>
  slice_head(n = 5) |>
  kable(caption = "First 5 records of events")

event_parties |>
  slice_head(n = 5) |>
  kable(caption = "First 5 records of event_parties")

org_nodes |>
  slice_head(n = 5) |>
  kable(caption = "First 5 records of org_nodes")

org_edges |>
  slice_head(n = 5) |>
  kable(caption = "First 5 records of org_edges")
```
:::


## Network construction

Not all events have clear from and to information. Therefore, only events with explicit or clear inferable interactions are included into the directed network edges. 

### Create event edges

::: panel-tabset
## The plot

```{r}
#| echo: false

get_detail <- function(x, path) {
  out <- purrr::pluck(x, !!!path, .default = NA_character_)
  
  if (length(out) == 0 || is.null(out) || all(is.na(out))) {
    NA_character_
  } else {
    as.character(out[1])
  }
}

get_party <- function(x, position) {
  if (length(x) >= position) {
    as.character(x[position])
  } else {
    NA_character_
  }
}

get_contacts <- function(x) {
  contacts <- purrr::pluck(x, "contacts", .default = NA_character_)
  
  if (length(contacts) == 0 || all(is.na(contacts))) {
    NA_character_
  } else {
    paste(unlist(contacts), collapse = "; ")
  }
}

event_edges <- events |>
  mutate(
    from = case_when(
      short_name == "sent" ~ map_chr(details, ~ get_detail(.x, c("from"))),
      
      short_name == "propose_meeting" ~ map_chr(details, ~ get_detail(.x, c("a2a", "from"))),
      
      short_name == "queue_subordinate_task" ~ map_chr(parties, ~ get_party(.x, 1)),
      
      short_name %in% c("read_file", "list_files", "access_files") ~ map_chr(parties, ~ get_party(.x, 1)),
      
      short_name == "suggest_contacts" ~ map_chr(parties, ~ get_party(.x, 1)),
      
      TRUE ~ NA_character_
    ),
    
    to = case_when(
      short_name == "sent" ~ map_chr(details, ~ get_detail(.x, c("to"))),
      
      short_name == "propose_meeting" ~ map_chr(details, ~ get_detail(.x, c("a2a", "to"))),
      
      short_name == "queue_subordinate_task" ~ map_chr(parties, ~ get_party(.x, 2)),
      
      short_name %in% c("read_file", "list_files") ~ map_chr(details, ~ get_detail(.x, c("target"))),
      
      short_name == "access_files" ~ map_chr(details, ~ get_detail(.x, c("file"))),
      
      short_name == "suggest_contacts" ~ map_chr(details, get_contacts),
      
      TRUE ~ NA_character_
    ),
    
    edge_type = short_name
  ) |>
  filter(
    !is.na(from),
    !is.na(to),
    from != "",
    to != "",
    to != "NA"
  ) |>
  select(
    event_id = id,
    datetime,
    short_name,
    from,
    to,
    edge_type
  )

event_edges |>
  slice_head(n = 10) |>
  kable(caption = "First 10 records of event_edges")

```

## The code

```{r}
#| eval: false
get_detail <- function(x, path) {
  out <- purrr::pluck(x, !!!path, .default = NA_character_)
  
  if (length(out) == 0 || is.null(out) || all(is.na(out))) {
    NA_character_
  } else {
    as.character(out[1])
  }
}

get_party <- function(x, position) {
  if (length(x) >= position) {
    as.character(x[position])
  } else {
    NA_character_
  }
}

get_contacts <- function(x) {
  contacts <- purrr::pluck(x, "contacts", .default = NA_character_)
  
  if (length(contacts) == 0 || all(is.na(contacts))) {
    NA_character_
  } else {
    paste(unlist(contacts), collapse = "; ")
  }
}

event_edges <- events |>
  mutate(
    from = case_when(
      short_name == "sent" ~ map_chr(details, ~ get_detail(.x, c("from"))),
      
      short_name == "propose_meeting" ~ map_chr(details, ~ get_detail(.x, c("a2a", "from"))),
      
      short_name == "queue_subordinate_task" ~ map_chr(parties, ~ get_party(.x, 1)),
      
      short_name %in% c("read_file", "list_files", "access_files") ~ map_chr(parties, ~ get_party(.x, 1)),
      
      short_name == "suggest_contacts" ~ map_chr(parties, ~ get_party(.x, 1)),
      
      TRUE ~ NA_character_
    ),
    
    to = case_when(
      short_name == "sent" ~ map_chr(details, ~ get_detail(.x, c("to"))),
      
      short_name == "propose_meeting" ~ map_chr(details, ~ get_detail(.x, c("a2a", "to"))),
      
      short_name == "queue_subordinate_task" ~ map_chr(parties, ~ get_party(.x, 2)),
      
      short_name %in% c("read_file", "list_files") ~ map_chr(details, ~ get_detail(.x, c("target"))),
      
      short_name == "access_files" ~ map_chr(details, ~ get_detail(.x, c("file"))),
      
      short_name == "suggest_contacts" ~ map_chr(details, get_contacts),
      
      TRUE ~ NA_character_
    ),
    
    edge_type = short_name
  ) |>
  filter(
    !is.na(from),
    !is.na(to),
    from != "",
    to != "",
    to != "NA"
  ) |>
  select(
    event_id = id,
    datetime,
    short_name,
    from,
    to,
    edge_type
  )

event_edges |>
  slice_head(n = 10) |>
  kable(caption = "First 10 records of event_edges")

```
:::

### Check edge types


::: panel-tabset
## The plot

```{r}
#| echo: false

event_edges |>
  count(edge_type, sort = TRUE) |>
  kable(caption = "Number of edges created by event type")
```

## The code

```{r}
#| eval: false
event_edges |>
  count(edge_type, sort = TRUE) |>
  kable(caption = "Number of edges created by event type")
```
:::



### Creating event nodes


::: panel-tabset
## The plot

```{r}
#| echo: false

event_nodes <- tibble(
  name = unique(c(event_edges$from, event_edges$to))
) |>
  mutate(
    node_type = case_when(
      str_detect(name, "^person:") ~ "Person",
      str_detect(name, "^Agent/") ~ "Agent",
      str_detect(name, "^agent:") ~ "Agent",
      str_detect(name, "^world:") ~ "System",
      str_detect(name, "^team:") ~ "Team",
      str_detect(name, "^company:") ~ "Company",
      str_detect(name, "\\.json$") ~ "File",
      name == "/" ~ "File System",
      TRUE ~ "Other"
    ),
    
    label = name |>
      str_remove("^person:") |>
      str_remove("^Agent/person:") |>
      str_remove("^agent:person:") |>
      str_remove("^world:") |>
      str_replace_all("_", " ") |>
      str_replace_all(":", " ") |>
      str_to_title()
  )

event_nodes |>
  slice_head(n = 10) |>
  kable(caption = "First 10 records of event_nodes")
```

## The code

```{r}
#| eval: false
event_nodes <- tibble(
  name = unique(c(event_edges$from, event_edges$to))
) |>
  mutate(
    node_type = case_when(
      str_detect(name, "^person:") ~ "Person",
      str_detect(name, "^Agent/") ~ "Agent",
      str_detect(name, "^agent:") ~ "Agent",
      str_detect(name, "^world:") ~ "System",
      str_detect(name, "^team:") ~ "Team",
      str_detect(name, "^company:") ~ "Company",
      str_detect(name, "\\.json$") ~ "File",
      name == "/" ~ "File System",
      TRUE ~ "Other"
    ),
    
    label = name |>
      str_remove("^person:") |>
      str_remove("^Agent/person:") |>
      str_remove("^agent:person:") |>
      str_remove("^world:") |>
      str_replace_all("_", " ") |>
      str_replace_all(":", " ") |>
      str_to_title()
  )

event_nodes |>
  slice_head(n = 10) |>
  kable(caption = "First 10 records of event_nodes")
```
:::


### Check node types

::: panel-tabset
## The plot

```{r}
#| echo: false

event_nodes |>
  count(node_type, sort = TRUE) |>
  kable(caption = "Number of nodes by node type")
```

## The code

```{r}
#| eval: false
event_nodes |>
  count(node_type, sort = TRUE) |>
  kable(caption = "Number of nodes by node type")
```
:::


### Creating event graph

::: panel-tabset
## The plot

```{r}
#| echo: false

event_graph <- tbl_graph(
  nodes = event_nodes,
  edges = event_edges,
  directed = TRUE
)

event_graph
```

## The code

```{r}
#| eval: false
event_graph <- tbl_graph(
  nodes = event_nodes,
  edges = event_edges,
  directed = TRUE
)

event_graph
```
:::


### Graph summary

::: panel-tabset
## The plot

```{r}
#| echo: false

tibble(
  total_nodes = gorder(event_graph),
  total_edges = gsize(event_graph),
  directed = is_directed(event_graph)
) |>
  kable(caption = "Summary of constructed event interaction graph")
```

## The code

```{r}
#| eval: false
tibble(
  total_nodes = gorder(event_graph),
  total_edges = gsize(event_graph),
  directed = is_directed(event_graph)
) |>
  kable(caption = "Summary of constructed event interaction graph")
```
:::

## Creating organisation network

::: panel-tabset
## The plot

```{r}
#| echo: false
org_graph <- tbl_graph(
  nodes = org_nodes,
  edges = org_edges,
  directed = TRUE
)

org_graph

```

## The code

```{r}
#| eval: false
org_graph <- tbl_graph(
  nodes = org_nodes,
  edges = org_edges,
  directed = TRUE
)

org_graph

```
:::

::: panel-tabset
## The plot

```{r}
#| echo: false
tibble(
  total_org_nodes = gorder(org_graph),
  total_org_edges = gsize(org_graph),
  directed = is_directed(org_graph)
) |>
  kable(caption = "Summary of constructed organisation graph")

```

## The code

```{r}
#| eval: false
tibble(
  total_org_nodes = gorder(org_graph),
  total_org_edges = gsize(org_graph),
  directed = is_directed(org_graph)
) |>
  kable(caption = "Summary of constructed organisation graph")

```
:::

## Integrating organisation structure

This is to link employees appearing in the event logs to their organisational context from the organisational chart.

::: panel-tabset
## The plot

```{r}
#| echo: false

event_people <- event_parties |>
  filter(actor_type == "Person") |>
  distinct(party)

people_org_lookup <- org_nodes |>
  filter(type == "person") |>
  select(
    party = id,
    person_label = label,
    title
  )

event_people_org <- event_people |>
  left_join(
    people_org_lookup,
    by = "party"
  )

event_people_org |>
  slice_head(n = 10) |>
  kable(caption = "First 10 event actors linked with organisation chart")

event_people_org |>
  filter(is.na(person_label)) |>
  kable(caption = "People appearing in event logs but not found in organisation chart")

```

## The code

```{r}
#| eval: false

event_people <- event_parties |>
  filter(actor_type == "Person") |>
  distinct(party)

people_org_lookup <- org_nodes |>
  filter(type == "person") |>
  select(
    party = id,
    person_label = label,
    title
  )

event_people_org <- event_people |>
  left_join(
    people_org_lookup,
    by = "party"
  )

event_people_org |>
  slice_head(n = 10) |>
  kable(caption = "First 10 event actors linked with organisation chart")

event_people_org |>
  filter(is.na(person_label)) |>
  kable(caption = "People appearing in event logs but not found in organisation chart")

```
:::



# System overview

The purpose is to provide context to understand where anomalous event occurred. There is a need to understand what "normal" activity looks like before investigating the anomaly.

## Event log analysis

Summarising what we learnt from the event log. 

::: panel-tabset
## The plot

```{r}
#| echo: false

events |>
  summarise(
    total_events = n(),
    earliest_event = min(datetime),
    latest_event = max(datetime)
  )

```

## The code

```{r}
#| eval: false

events |>
  summarise(
    total_events = n(),
    earliest_event = min(datetime),
    latest_event = max(datetime)
  )

```
:::

::: panel-tabset
## The plot

```{r}
#| echo: false

events |>
  count(short_name, sort = TRUE)

```

## The code

```{r}
#| eval: false

events |>
  count(short_name, sort = TRUE)

```
:::

::: panel-tabset
## The plot

```{r}
#| echo: false

events |>
  count(n_parties)

```

## The code

```{r}
#| eval: false
events |>
  count(n_parties)

```
:::

::: panel-tabset
## The plot

```{r}
#| echo: false

event_parties |>
  count(actor_type)

```

## The code

```{r}
#| eval: false
event_parties |>
  count(actor_type)

```
:::

## Event log summary table:

::: panel-tabset
## The plot

```{r}
#| echo: false

event_summary <- tibble(
  Metric = c(
    "Total Events",
    "Unique Event Types",
    "Earliest Event",
    "Latest Event"
  ),
  Value = c(
    nrow(events),
    n_distinct(events$short_name),
    min(events$datetime),
    max(events$datetime)
  )
)

event_summary |>
  kable(caption = "Summary of Event Log")
```

## The code

```{r}
#| eval: false
event_summary <- tibble(
  Metric = c(
    "Total Events",
    "Unique Event Types",
    "Earliest Event",
    "Latest Event"
  ),
  Value = c(
    nrow(events),
    n_distinct(events$short_name),
    min(events$datetime),
    max(events$datetime)
  )
)

event_summary |>
  kable(caption = "Summary of Event Log")

```
:::

## Event participant composition:

::: panel-tabset
## The plot

```{r}
#| echo: false

event_parties |>
  count(actor_type) |>
  ggplot(
    aes(
      x = reorder(actor_type, n),
      y = n
    )
  ) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Composition of Actors Appearing in Event Logs",
    x = "Actor Type",
    y = "Number of Appearances"
  ) +
  theme_minimal()

```

## The code

```{r}
#| eval: false
event_parties |>
  count(actor_type) |>
  ggplot(
    aes(
      x = reorder(actor_type, n),
      y = n
    )
  ) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Composition of Actors Appearing in Event Logs",
    x = "Actor Type",
    y = "Number of Appearances"
  ) +
  theme_minimal()

```
:::

## Events by number of parties:

::: panel-tabset
## The plot

```{r}
#| echo: false

events |>
  count(n_parties) |>
  ggplot(
    aes(
      x = factor(n_parties),
      y = n
    )
  ) +
  geom_col() +
  labs(
    title = "Number of Participants per Event",
    x = "Number of Parties",
    y = "Number of Events"
  ) +
  theme_minimal()

```

## The code

```{r}
#| eval: false
events |>
  count(n_parties) |>
  ggplot(
    aes(
      x = factor(n_parties),
      y = n
    )
  ) +
  geom_col() +
  labs(
    title = "Number of Participants per Event",
    x = "Number of Parties",
    y = "Number of Events"
  ) +
  theme_minimal()

```
:::


## Event type distribution visual
Here we aim to find what activities dominate the system.

::: panel-tabset
## The plot

```{r}
#| echo: false

event_type_counts <- events |>
  count(short_name, sort = TRUE)

event_type_counts |>
  ggplot(
    aes(
      x = reorder(short_name, n),
      y = n
    )
  ) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Distribution of Event Types in Tenant Thread Activity Logs",
    subtitle = "This shows which activities dominate the multi-agent system.",
    x = "Event Type",
    y = "Number of Events"
  ) +
  theme_minimal()

```

## The code

```{r}
#| eval: false

event_type_counts <- events |>
  count(short_name, sort = TRUE)

event_type_counts |>
  ggplot(
    aes(
      x = reorder(short_name, n),
      y = n
    )
  ) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Distribution of Event Types in Tenant Thread Activity Logs",
    subtitle = "This shows which activities dominate the multi-agent system.",
    x = "Event Type",
    y = "Number of Events"
  ) +
  theme_minimal()

```
:::


## Daily system activity visual

::: panel-tabset
## The plot

```{r}
#| echo: false
events |>
  summarise(
    min_date = min(date),
    max_date = max(date),
    n_days = n_distinct(date)
  )


```

## The code

```{r}
#| eval: false

events |>
  summarise(
    min_date = min(date),
    max_date = max(date),
    n_days = n_distinct(date)
  )

```
:::

This shows there are 70 records, 70 distinct days, therefore we can do a daily activity line chart.

::: panel-tabset
## The plot

```{r}
#| echo: false
daily_activity <- events |>
  count(date)

daily_activity |>
  ggplot(
    aes(
      x = date,
      y = n
    )
  ) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(
    title = "Daily System Activity Over Time",
    subtitle = "Number of events recorded per day",
    x = "Date",
    y = "Number of Events"
  ) +
  theme_minimal()


```

## The code

```{r}
#| eval: false

daily_activity <- events |>
  count(date)

daily_activity |>
  ggplot(
    aes(
      x = date,
      y = n
    )
  ) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(
    title = "Daily System Activity Over Time",
    subtitle = "Number of events recorded per day",
    x = "Date",
    y = "Number of Events"
  ) +
  theme_minimal()

```
:::

## Most common interaction paths

Here, it would be necessary to discover which interactions occur most frequently within the multi-agent system.  

::: panel-tabset
## The plot

```{r}
#| echo: false

common_paths <- event_edges |>
  count(from, to, sort = TRUE) |>
  slice_head(n = 15) |>
  mutate(
    interaction_path = paste(from, "→", to)
  )

common_paths |>
  ggplot(
    aes(
      x = reorder(interaction_path, n),
      y = n
    )
  ) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Most Common Interaction Paths",
    subtitle = "Top 15 interaction paths observed in the system",
    x = "Interaction Path",
    y = "Frequency"
  ) +
  theme_minimal()

```

## The code

```{r}
#| eval: false

common_paths <- event_edges |>
  count(from, to, sort = TRUE) |>
  slice_head(n = 15) |>
  mutate(
    interaction_path = paste(from, "→", to)
  )

common_paths |>
  ggplot(
    aes(
      x = reorder(interaction_path, n),
      y = n
    )
  ) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Most Common Interaction Paths",
    subtitle = "Top 15 interaction paths observed in the system",
    x = "Interaction Path",
    y = "Frequency"
  ) +
  theme_minimal()

```
:::


## system interaction network visual

::: panel-tabset
## The plot

```{r}
#| echo: false

set.seed(123)

ggraph(event_graph,
       layout = "fr") +

  geom_edge_link(
    alpha = 0.15,
    colour = "grey70"
  ) +

  geom_node_point(
    aes(colour = node_type),
    size = 3
  ) +

  labs(
    title = "System Interaction Network",
    subtitle = "Interactions between people, agents, systems and files"
  ) +

  theme_graph()
```

## The code

```{r}
#| eval: false

set.seed(123)

ggraph(event_graph,
       layout = "fr") +

  geom_edge_link(
    alpha = 0.15,
    colour = "grey70"
  ) +

  geom_node_point(
    aes(colour = node_type),
    size = 3
  ) +

  labs(
    title = "System Interaction Network",
    subtitle = "Interactions between people, agents, systems and files"
  ) +

  theme_graph()

```
:::


## Organisation structure visual

The purpose is to show where John Winward sits in the company, which gives organisational context. This will help give context and insight when looking into this event chain containing the anomaly post by John.

::: panel-tabset
## The plot

```{r}
#| echo: false

john_org_nodes <- c(
  "company:tenant_thread",
  "department:customer_support",
  "team:phone_center",
  "team:billing",
  "team:concierge",
  "team:public_relations",
  "person:john_windward"
)

customer_support_nodes <- org_nodes |>
  filter(id %in% john_org_nodes) |>
  mutate(
    highlight = id == "person:john_windward"
  )

customer_support_edges <- org_edges |>
  filter(
    source %in% john_org_nodes,
    target %in% john_org_nodes
  )

customer_support_graph <- tbl_graph(
  nodes = customer_support_nodes,
  edges = customer_support_edges,
  directed = TRUE
)

set.seed(123)

ggraph(customer_support_graph, layout = "tree") +
  geom_edge_link(
    arrow = arrow(length = unit(2, "mm")),
    end_cap = circle(2, "mm"),
    colour = "grey70"
  ) +
  geom_node_point(
    aes(
      colour = type,
      size = highlight
    )
  ) +
  geom_node_text(
    aes(label = label),
    repel = TRUE,
    size = 3
  ) +
  scale_size_manual(
    values = c("FALSE" = 3, "TRUE" = 6),
    guide = "none"
  ) +
  labs(
    title = "Organisation Structure Around John Windward",
    subtitle = "John Windward is highlighted within the Customer Support department",
    colour = "Node Type"
  ) +
  theme_graph()
```

## The code

```{r}
#| eval: false

john_org_nodes <- c(
  "company:tenant_thread",
  "department:customer_support",
  "team:phone_center",
  "team:billing",
  "team:concierge",
  "team:public_relations",
  "person:john_windward"
)

customer_support_nodes <- org_nodes |>
  filter(id %in% john_org_nodes) |>
  mutate(
    highlight = id == "person:john_windward"
  )

customer_support_edges <- org_edges |>
  filter(
    source %in% john_org_nodes,
    target %in% john_org_nodes
  )

customer_support_graph <- tbl_graph(
  nodes = customer_support_nodes,
  edges = customer_support_edges,
  directed = TRUE
)

set.seed(123)

ggraph(customer_support_graph, layout = "tree") +
  geom_edge_link(
    arrow = arrow(length = unit(2, "mm")),
    end_cap = circle(2, "mm"),
    colour = "grey70"
  ) +
  geom_node_point(
    aes(
      colour = type,
      size = highlight
    )
  ) +
  geom_node_text(
    aes(label = label),
    repel = TRUE,
    size = 3
  ) +
  scale_size_manual(
    values = c("FALSE" = 3, "TRUE" = 6),
    guide = "none"
  ) +
  labs(
    title = "Organisation Structure Around John Windward",
    subtitle = "John Windward is highlighted within the Customer Support department",
    colour = "Node Type"
  ) +
  theme_graph()

```
:::


# Extracting the anomalous chain

The purpose is to reconstruct the exact sequence of events leading to the anomalous post.

## Locating  the target event

## event window extraction

## Incident timeline visual

## Event chain network

## key actors involved

# Content origin analysis

The purpose is to determine what the post means and where its contents originated.

## Extract text sources

## Content Similarity Analysis visual

## Evidence supporting content origin

## interpretation of post meaning

# Historical behaviour analysis

The purpose is to determine where the incident is isolated. Also, to determine whether similar incidents occurred previously.

## Identifying similar events

## Historical incident timeline

## Comparative analysis

Using the comparison table, compare between historical cases and the May 17 incident.

## Historical event chain comparison

# Intervention analysis

## Centrality analysis

## Critical node analysis visual

## Candidate intervention points

## Recommended intervention

# Mini challenge answers

1. How was the anomalous SaidIT post made?

2. What does the post mean?

3. Are there prior occurrences?

4. What intervention should be implemented?

# Conclusion



